### 2020-02-01 main.tex
Initial Overleaf Import

 \documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4

%\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
                                                          % if you need a4paper

\usepackage{FG2019}
\usepackage{multirow}
\newcommand{\minitab}[2][c]{\begin{tabular}{#1}#2\end{tabular}}
\usepackage[acronym]{glossaries}
\usepackage{acronym}
\usepackage{bigstrut}
% symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amssymb,amsthm}
\usepackage{amsmath}% assumes amsmath package installed
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
% \usepackage{hyperref}       % hyperlinks
\usepackage[colorlinks = true,
            linkcolor = blue,
            urlcolor  = blue,
            citecolor = blue,
            anchorcolor = blue]{hyperref}

\usepackage{url}            % simple URL typesetting
\usepackage{mathptmx} % assumes new font selection scheme installed
\usepackage{times} % assumes new font selection scheme installed
\usepackage{lipsum}
% \usepackage{slashbox}
\usepackage{booktabs}       % professional-quality tables
\usepackage{enumitem}
\usepackage [english]{babel}
\usepackage [autostyle, english = american]{csquotes}
\MakeOuterQuote{"}
\usepackage[font=itshape]{quoting}
\usepackage{lipsum}
% \usepackage{subfig}
%% add
\usepackage{tikz}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{color, colortbl}
\usepackage{comment}
\usepackage{tabularx, booktabs, ragged2e}
\usepackage{arydshln}
\usepackage{todonotes}
\usepackage{ctable}
\usepackage{multirow}
\usepackage{makecell}

% \usepackage[small]{caption}
% \usepackage{subcaption}
% \usepackage{subfigure} 
\usepackage{sidecap}
\usepackage{wrapfig}
\usepackage[subrefformat=parens,labelformat=parens]{subcaption} %% provides subfigure
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{float}% If comment this, figure moves
\usepackage{geometry}

\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage[]{algorithm2e}
\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont

\graphicspath{{images/}} %Setting the graphicspath
\usepackage{bibunits}

\defaultbibliography{bias}
\defaultbibliographystyle{ieee}

\usepackage{mathtools}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}

\usepackage{array}
\newcolumntype{x}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}

% \import{sections/}{experimental.tex}
% \import{sections/}{analysis.tex}

% \newcommand{\argmax}{\mathop{\mathrm{argmax}}\limits} 
\newcommand{\argmax}{\mathop{\mathrm{argmax}}\limits} 
\newcommand{\argmin}{\mathop{\mathrm{argmin}}\limits}

\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\newcommand{\xmark}{\ding{56}}%
\newcommand{\checkc}{\ding{51}}%

\newcommand{\com}[1]{\textcolor{red}{#1}}
\newcommand{\ie}{\textit{i}.\textit{e}., }
\newcommand{\eg}{\textit{e}.\textit{g}., }
\newcommand*{\etc}{etc.\@\xspace}

\newcommand{\yann}[1]{\todo[color=orange!40]{\rotatebox{90}{YANN: #1}}}
\newcommand{\samson}[1]{\todo[color=blue!40]{\rotatebox{90}{SAMSON: #1}}}
\newcommand{\joe}[1]{\todo[color=purple!40]{\rotatebox{90}{JOE: #1}}}
\newcommand{\bob}[1]{\todo[color=green!40]{\rotatebox{90}{BOB: #1}}}
\newcommand{\gena}[1]{\todo[color=orange!40]{\rotatebox{90}{GENA: #1}}}

\newacronym{m}{M}{\textit{Male}}
\newacronym{f}{F}{\textit{Female}}
\newacronym{a}{A}{\textit{Asian}}
\newacronym{b}{B}{\textit{Black}}
\newacronym{i}{I}{\textit{Indian}}
\newacronym{w}{W}{\textit{White}}
\newacronym{af}{AF}{\textit{Asian}-\textit{Female}}
\newacronym{am}{AM}{\textit{Asian}-\textit{Male}}
\newacronym{bf}{BF}{\textit{Black}-\textit{Female}}
\newacronym{bm}{BM}{\textit{Black}-\textit{Male}}
\newacronym{if}{IF}{\textit{Indian}-\textit{Female}}
\newacronym{im}{IM}{\textit{Indian}-\textit{Male}}
\newacronym{wf}{WF}{\textit{White}-\textit{Female}}
\newacronym{wm}{WM}{\textit{White}-\textit{Male}}

\newacronym{ml}{ML}{machine learning}
\newacronym{fr}{FR}{facial recognition}
\newacronym{fv}{FV}{facial verification}

\newacronym{cnn}{CNN}{convolutional neural network}
\newacronym{nn}{NN}{neural network}
\newacronym{mtcnn}{MTCNN}{\emph{multi-task \gls{cnn}}}

\newacronym{gan}{GAN}{generative adversarial network}
\newacronym{se}{SE}{\emph{Squeeze-and-Excitation}}
\newacronym{d}{$D$}{discriminator}
\newacronym{g}{$G$}{generator}
\newacronym{dbvae}{DB-VAE}{Debiasing Variational Autoencoder}

\newacronym{lut}{LUT}{Look-Up-Table}
\newacronym{soa}{SOTA}{state-of-the-art}

\newacronym{fiw}{FIW}{Families In the Wild}
\newacronym{lfw}{LFW}{Labeled Faces in the Wild}
\newacronym{bfw}{BFW}{Balanced Faces In the Wild}
\newacronym{rfw}{RFW}{Racial Faces in-the-Wild:}
\newacronym{dp}{DemogPairs}{Demographic Pairs}
\newacronym{itwcc}{ITWCC}{Wild Child Celebrity}

\newacronym{fd}{FD}{Face Discrimination}
\newacronym{bb}{BB}{bounding box}

\newacronym{sdm}{SDM}{signal detection model}
\newacronym{roc}{ROC}{receiver operating characteristic}
\newacronym{nmse}{NMSE}{Normalized Mean Square Error}
\newacronym{det}{DET}{Detection Error Trade-off}
\newacronym{tp}{TP}{true-positive}
\newacronym{tpir}{TPIR}{true-positive identification rate}
\newacronym{frir}{FRIR}{false-reject identification rate}
\newacronym{fpir}{FRIR}{false-positive identification rate}

\newacronym{fn}{FN}{false-negative}
\newacronym{frr}{FRR}{false-reject rate}
\newacronym{fnr}{FNR}{false-negative rate}

\newacronym{fpr}{FPR}{false-positive rate}
\newacronym{tpr}{TPR}{true-positive rate}

\newacronym{tar}{TAR}{True Acceptance Rate}
\newacronym{far}{FAR}{False Acceptance Rate}
\newacronym{eer}{EER}{Equal Error Rate}
\newacronym{det}{DET}{Detection Error Trade-off}

\newacronym{cs}{CS}{Cosine Similiarity}

\newacronym{lime}{LIME}{Local Interpretable Model-Agnostic Explanations}
\newacronym{nas}{NAS}{Neural Architecture Search}
\newacronym{gapf}{GAPF}{Generative Adversarial Privacy and Fairness}

                                                          % paper
%\setlist[itemize]{leftmargin=*}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\mc}[2]{\multicolumn{#1}{c}{#2}}
\definecolor{Gray}{gray}{0.85}
\definecolor{LightCyan}{rgb}{0.88,1,1}

\newcolumntype{a}{>{\columncolor{Gray}}c}
\newcolumntype{b}{>{\columncolor{white}}c}
\renewcommand{\arraystretch}{1.4}
\newcommand{\xmark}{\ding{53}}%

\newcommand{\highlightb}[1]{%
\colorbox{blue!30}{$\displaystyle#1$}}
\newcommand{\vo}{\vec{o}\@ifnextchar{^}{\,}{}}

\newcommand{\vx}{\vec{x}\@ifnextchar{^}{\,}{}}

\FGfinalcopy % *** Uncomment this line for the final submission
\let\vec\mathbf

% \IEEEoverridecommandlockouts                              % This command is only
                                                          % needed if you want to
                                                          % use the \thanks command
% \overrideIEEEmargins
% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files

\def\FGPaperID{41} % *** Enter the FG 2019 Paper ID here

\title{\LARGE \bf\vspace{-10mm}
Face Recognition: Too Bias, or Not Too Bias?
}

%use this in case of a single affiliation
%\author{\parbox{16cm}{\centering
%    {\large Huibert Kwakernaak}\\
%    {\normalsize
%    Faculty of Electrical Engineering, Mathematics and Computer Science, University of Twente, Enschede, The Netherlands\\}}
%    \thanks{This work was not supported by any organization.}% <-this % stops a space
%}

%use this in case of several affiliations
\author{\parbox{16cm}{\centering
    {\large Joseph P Robinson$^1$, Gennady Livitz$^2$, Yann Henon$^2$, Can Qin$^1$,\\ Yun Fu$^1$, and Samson Timoner$^2$}\\
    {\normalsize
    \hspace{-.4in}$^{1}$Northeastern University\hspace{.7in} $^{2}$ISM Connect}}
    % $^1$ Northeastern University, Boston, MA, USA\\
    % $^2$ ISM Connect, Cambridge, MA, USA}}
    \thanks{}% <-this % stops a space
}

% \author{Joseph P Robinson$^1$, Yuncheng Li$^{2}$, Ning Zhang$^{2}$, Yun Fu$^{1}$, and Sergey Tulyakov$^{2}$\\
% \hspace{-.4in}$^{1}$Northeastern University\hspace{.7in} $^{2}$Snap Inc. \\\vspace{2mm}
% } %\\

\begin{document}
\begin{bibunit}

\def\angle{0}
\def\radius{3}
\def\cyclelist{{"orange","blue","red","green","magenta","cyan"}}
\newcount\cyclecount \cyclecount=-1
\newcount\ind \ind=-1

\ifFGfinal
\thispagestyle{empty}
\pagestyle{empty}
\else
\author{Anonymous FG 2019 submission\\ Paper ID \FGPaperID \\}
\pagestyle{plain}
\fi
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We reveal critical insights into problems of bias in state-of-the-art \gls{fr} systems using a novel \gls{bfw} dataset: data balanced for gender and ethnic groups. Classic signal detection theory revealed trends in the underlying score distribution across subgroups. Specifically, we show variations in the optimal scoring threshold varies for face-pairs across different subgroups. Thus, the conventional approach of learning a  global threshold for all pairs resulting in performance gaps among subgroups. By learning subgroup-specific thresholds, we not only mitigate problems in performance gaps but also show a notable boost in the overall performance. Furthermore, we do a human evaluation to measure the bias in humans, which supports the hypothesis that such a bias exists in human perception. 
\end{abstract}

\input{sections/introduction.tex}

\input{sections/relatedworks.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{sections/section3.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{sections/analysis.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t!]
    \caption{\small{\textbf{Quantitative of human assessment.} Different human subgroups listed per row. Each column is the subgroup labeled. Note that each people are best within their subgroup, and second-best within the same subgroup but different gender. CF shows the least variation, but with the lowest accuracy. CM shows the best accuracy, but second to \gls{wm} in deviation from the mean. Thus, scores for males vary more than females.}}
    \label{tab:humsn-eval-results} 
\begin{center}
\footnotesize
\scalebox{0.9}{
\begin{tabular}{l c c c c c}
        &  CF  &      CM     &   WF     &   WM     &  Avg\\\midrule
        CF &  \textbf{0.529}&  0.480&0.438&0.447 &0.474\pm0.041 \\
        CM & 0.456 & \textbf{0.504}  & 0.444 &0.362 &0.441\pm0.059 \\
        WF & 0.447 &0.438 &  \textbf{0.573}& 0.480 & 0.485\pm0.062 \\
        WM & 0.301&0.474 &  0.453 & \textbf{0.561} & 0.447\pm0.108 \\\midrule
        Avg &  0.433& 0.474&0.477 &0.463 &0.462\pm0.020\\
 \end{tabular}}
 \end{center}
 \vspace{-5mm}
\end{table}

\glsresetall
\section{CONCLUSION}
We introduce a new data set \gls{bfw} with eight subgroups balanced across gender and ethnicity. With this, and upon highlighting the challenges and shortcomings of grouping subjects as a single subset, we provide evidence that forming subgroups is meaningful, as the \gls{fr} algorithm rarely makes mistakes across subgroups. We trained Arc-Face net on MSCeleb, expecting that the results would suffer from bias because of the imbalanced train-set. Once established that the results do suffer from problems of bias, we observed that the same threshold across ethnic and gender subgroups leads to differences in the \gls{fpr} up to a factor of two, which is seemingly the cause of the frenzy about bias in \gls{fr} in main-stream media. Furthermore, we ameliorate these differences with a per-subgroup threshold, leveling out \gls{fpr}, and achieving a higher \gls{tpr}. We hypothesized that most humans grown amongst more than their own demographic and, therefore, effectively learn from imbalanced datasets-- a human evaluation validated that humans are biased, as most recognized their personal demographic best. The focused research findings presented here, along with the public database included, are extendable in vast ways. Thus, we see this as the slither to a much larger problem of bias in ML.

\putbib
\end{bibunit}

% \newpage\newpage
% \col
\begin{bibunit}
\input{supplemental.tex}
\end{bibunit}
% \end{comment}

\end{document}

### 2020-02-01 notes.tex
Initial Overleaf Import

\begingroup
\let\clearpage\relax 
\onecolumn 
\newpage

\section*{TODO}
\subsection{Joe}
\begin{enumerate}
    \item  WRITING
    \begin{itemize}
        \item More complete lit review-- complete draft 1 of Related Works (1 week)
    \end{itemize} 
    \item DATA
    \begin{itemize}
        \item Confirm no overlap between Deomonpairs and training data + write about this/ resources in more depth in experimental setup & implementation (3/4 week)
        \item Verify data used to train arcface being used, put in report, along other with spec of network (footnote to its repo)
        \item Various figures showing daataset stats, like distribution of ethnicity, subjects per, faces per subject, etc., Make sure to produce in way updating is quick and easy, as plan to speak more about the goals for data. Figures and tables in report, with next versions looking and feeling (and sized) as intended for this draft/ at this point in time (2 days). In summary: Provide absolute data stats and format as table in paper. Update Fig 4 and Table 1. Discuss data goal (more on this under @Samson's TODO).
    \end{itemize}
    \item EXPERIMENTS
    \begin{itemize}
        \item Varying threshold
        \begin{itemize}
            \item Make DET for all pairs combines, and then see improvements with varying threshold (i.e., single list with all pairs, first run with single threshold found on held out set, and then another run with variable threshold as found to be best for each subgroup
        \end{itemize}
        \begin{itemize}
            \item Use ethnicity classifier to determine best threshold via LUT built on held-out validation data
            \item use NN list of subgroup to predict the subgroup in a semi-supervised manner (ie say use just the labels we have on our set and use on larger scale experiment with samples without attribute labels
        \end{itemize}
        \item FAR\@Rank 1 + 5 table for various thresholds and all (Need more empirical ways to evaluate and baseline) 
        \item Find evidence (interrupt model)— why Asian females fail— is there a common pattern in those that pass (checkout LIME)
    \item ANALYSIS
    \begin{itemize}
        \item How is the quality of images they fail against, i.e., low?
        \item  Alignment part— how do we do aligning Asian females? Face detection?
        \item Skin color of Asians in training data (i.e., check average face)

    \end{itemize}
    \end{itemize}
\end{enumerate}
\subsection{Samson}
\begin{enumerate}
    \item Sanity test with mxnet (Samson),  26 Friday 2019
    \item Patrick at NIST (Samson: Tomorrow set a time)
    \item Run repeat test with mx-net version of ArcFace (resnet 100)
    \item Reach out to IBM, for Diversity in Faces
    \item Work with Joe to determine goals for data, i.e., specifications of what we would like to include and means of obtaining more if needed.
\end{enumerate}
\subsection{Gena}
\begin{enumerate}
    \item TBD
\end{enumerate}

\section*{IDEAS}
\noindent\texbf{(1) Fig. 1 - Improvements for Signal Detection Model}\\
-- make Fig 1 larger, magnify overlapping region and highlight as different color (i.e., maybe even using a simple clip art and dress it up with PP (or equivalent) to create attractive way to zoom... then, perhaps show sample pairs along the few, or, say, as (b) to the right of histograms, with one section for each graph--- fig : (a) visualization in Fig 1 with zoomed in effect |<horizontal dashed line>| (b) rows should pair samples (different type per row in some way highlighting where in the histogram each sample is found)

\noindent\texbf{(2) Fig. ~\ref{fig:detcurves}}\\
-- add lines depicting threshold for a specific rate (ie horizontal to put more emphasis on different in threshold).

-- Make 3 sperate PDFs and use latex subgraphs tighten up view, while adding subcaption references (i.e., (a) (b) and (c)

\begin{figure}[h]
\setlength{\unitlength}{0.14in} % selecting unit length
\centering % used for centering Figure
\begin{picture}(32,15) % picture environment with the size (dimensions)
 % 32 length units wide, and 15 units high.
\put(3,4){\framebox(6,3){$H_{B}(q)$}}
\put(13,4){\framebox(6,3){$N[\cdot]$}}
\put(23,4){\framebox(6,3){$H_{C}(q)$}}
\put(0,5.5){\vector(1,0){3}}\put(9,5.5){\vector(1,0){4}}
\put(19,5.5){\vector(1,0){4}}\put(29,5.5){\vector(1,0){3}}
\put(-1,6.5) {$u(k)$}\put(30,6.5) {$y(k)$} \put(9.5,6.5)
{$x_{B}(k)$}\put(19.5,6.5) {$x_{C}(k)$}
\end{picture}
\caption{An LNL Block Oriented Model Structure} % title of the Figure
\label{fig:lnlblock} % label to refer figure in text
\end{figure}

\section*{MY VIEW ON ADDING MORE AUTHORS AND COLLABORATIONS}
Some prefer rather fewer authors (surprisingly, more than I had imagined), but it is an important topic to be transparent about, so here are my feelings in the matter: I am 100\% for more authors if it means a better paper-- if there are ways to improve a paper, and someone has means to deliver on this, and then they deliver or are trusted to deliver, then offer them a part of this effort. This often inspires, which can directly translate to motivate. With that gives us an extra pair of hands to write, eyes and mind to revise, and added skills to enhance visuals (surely unique skill set should compliment deliverable, assuming the deliverable in itself is significant and is all expected of them (ie proposed model, building data collection pipeline and managing data collection + provide stats/ montages for database contributions, novel model idea and code, etc.,)). However, one noteworthy item and 110\% effort on improving paper and working on action items could always pay off provided good candidate. Thus, such a goal with promise as this paper has could inspire, motivate, and be appreciated by colleague. Thus, I am usually more about including then the opposite. Though, of course, they must contribute (which is something to speak about up front with complete transparency)... In summary: It is subjective, but the more the merrier 90\% of the time in my book (10\% only as a grad student when you need those 1 or 2 papers with just you and advisor, but that is not case of course + those are in the past at this point :) )>... that is my feelings on authorship in case you were wondering lol ;) Thus, if others you think would be interested please let me know and arrange to chat about it.... Also, notice ~80\% of the best papers have several, several authors, while the others are break-through ideas.

\endgroup

### 2020-02-01 sections/analysis.tex
Initial Overleaf Import

\begin{figure}[t!]
	\centering    
	\includegraphics[trim=.6in 1.4in 2.1in 1.65in,clip,width=.8\linewidth]{images/confusion.pdf}
		\caption{\small{\textbf{Confusion matrix.} Percent error (Rank 1) for all faces of \gls{bfw} queried against all others. Notice errors concentrate within a subgroup, consistent with the \gls{sdm} in Fig.~\ref{fig:detection-model} (\ie \gls{af} show worst performance, and is mostly confused with faces of the same demographic). This plot is evidence that while race/ethnicity may be challenging to define, the subgroups are meaningful.}}
		\label{fig:confusion} 
\end{figure}

\section{Results and Analysis}
% The \gls{soa} Arcface~\cite{deng2019arcface} was used throughout to control the model across all experiments.

We experimented with multiple \gls{cnn}-based models to support our claims of variation in optimal threshold across subgroups is not model-specific, but model-agnostic. Specifically, we analyze three \glspl{cnn} variants: we encode faces using VGG-16~\cite{simonyan2014very},  50-layer residual network (\ie ResNet-50)~\cite{he2016deep}, and \gls{senet} (\ie SENet-50)~\cite{hu2018squeeze}, with results of only the latter (\ie the best performing \gls{senet}) model used throughout the main paper. Results for the other two, alongside \gls{senet} for comparison, are in the Supplemental Material.

\glsunset{am}\glsunset{af}\glsunset{bm}\glsunset{bf}\glsunset{im}\glsunset{if}\glsunset{wm}\glsunset{wf}

% \vspace{-2pt}
\subsection{Score Analysis}
Fig.~\ref{fig:detection-model} shows score distributions for faces of the same identity (\ie \emph{Genuine}) and different (\ie \emph{Imposter}), with \gls{sdm}s split by subgroup. Observe that scores of imposters  approximately distribute about a peak of zero independent of subgroup, with slightly different standard deviation. Observe that the broadness in score distribution of the \emph{genuine} pairs varies across subgroups. Asian Fem
Fig.~\ref{fig:confusion} shows the confusion matrix of the subgroups. A vast majority of errors occurs in intra-subgroup. It is interesting to note that while the definition of  each group  based on ethnicity and race may not be crisply defined, the confusion matrix indicates that in practice the \gls{cnn} finds that the groups are effectively separate. The categories are, therefore, meaningful in the context of \gls{fr}.

\subsection{\gls{det} analysis}
\glsunset{m}
\glsunset{f}

\gls{det} curves averaged across 5-folds show per-subgroup trade-offs (Fig.~\ref{fig:detcurves}). Note that \gls{m} perform better than \gls{f}, precisely as one would expect from the tails of score-distributions for \emph{genuine} pairs (Fig.~\ref{fig:detection-model}). \Gls{af} and \gls{if} perform the worst.

\begin{table}[t!]
\glsunset{tar}
\glsunset{far}
\caption{\small{\textbf{\gls{tar} at intervals of \gls{far}}. For each subgroup, and overall average \gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom). Higher is better. The proposed shows improvement in all cases.}}\label{tab:ethnicy-far} 
\include{tables/ethnicity_far}
\glsreset{tar}
\glsreset{far}
\end{table}

The gender-based \gls{det} curve shows that there are different performances for \gls{m} and \gls{f} at a fixed threshold (dashed-line). Similar effects exist for the other curves as well (lines omitted to declutter). Many \gls{fr} applications, systems are operated at the highest \gls{fpr} allowed, so the line of constant threshold indicates that a single threshold produced at different operating points (\ie \gls{fpr}) \gls{m} and \gls{f}, which is undesirable.  The difference in \gls{fpr} is approximately a factor of 2-- quite large indeed. If this kind of difference is common in industry, one would expect a factor of 3 more false positives to be reported based on one subgroup than another, which has a strong potential to be the cause of media articles reporting bias in 
\gls{fr}~\cite{england2019,snow2018}.

\subsection{Verification threshold} \label{subsec:analysis:verification}
We seek to reduce the bias between subgroups. Such that an operating point (\ie \gls{fpr}) is constant across subgroups. To accomplish that, we used a per subgroup threshold. In \gls{fv}, we consider one image as the query and all others as test. For this, the ethnicity of the query image is assumed. We can then examine the \gls{det} curves and pick the best threshold per subgroup for a certain \gls{fpr}.

We evaluated \gls{tar} for specific \gls{far} values. As described in Section~\ref{subsec:pf}, the verification experiments were 5-fold, with no overlap in subject ID between folds. Results reported are averaged across folds in all cases and are shown in Table~\ref{tab:ethnicy-far}. For each subgroup, the \gls{tar} of using a global-threshold is reported (upper row), as well as using the optimal per subgroup threshold (lower row).

Even for lower \gls{far}, there are notable improvements, often of the order of 1\%, which can be challenging to achieve when \gls{far} is near $\geq$90\%. More importantly, each subgroup has the desired \gls{fpr}, so that substantial differences in \gls{fpr} will remain unfounded. We further experimented using ethnicity estimators and using the ethnicity of both query and test image, which yielded similar results to those reported here (results not shown).

\begin{figure}[t!] 
	\centering    
	\includegraphics[trim=0in 0.0in 0in 0in,clip, width=\linewidth] {images/human_eval.pdf}
		\caption{\small{\textbf{Qualitative results of human assessment.} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}, which scores plotted next to each. Humans tend to be more successful at recognizing their own subgroup, with few exceptions (bottom). }}
		% (\eg bottom row). } Result summary in Table~\ref{tab:human-eval}.
		\label{fig:human-eval} 
		\vspace{-5mm}
\end{figure}

% \vspace{-5pt}
\subsection{Human evaluation}
 Quantitative and qualitative results are in Table~\ref{tab:humsn-eval-results} and Fig.~\ref{fig:human-eval}, respectfully. One might expect that the most exposure to others would be within the same subgroup, and, therefore, would be best at labeling their own. Secondarily, they would be best at labeling images of the same ethnicity, but opposite gender. Our findings concur. Each subgroup is best at labeling their type, and then second best at labeling the same ethnicity but opposite sex. Interestingly, each group of images is best tagged by the corresponding subgroup, with the second-to-best having the same ethnicity and opposite gender. On average, subgroups are comparable at labeling images.

### 2020-02-01 sections/introduction.tex
Initial Overleaf Import

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\glsresetall
\section{Introduction}
As more of society becomes integrated with \gls{ml}, bias, fairness, and the formalization of \gls{ml} standards are topics of high interest ~\cite{10.1007/978-3-030-13469-3_68, anne2018women, wang2018racial}. One effect of the growing dependency on technology is the increasing concern regarding biased and unfair algorithms.  For instance, \gls{fr} systems can be untrustworthy and racist in some cases~\cite{england2019,snow2018}.

%The trust of people in machinery built to make decisions that range from casual-to-critical matters still takes time to build  .

\begin{figure}[t!]
\centering
         \includegraphics[trim=0in 0.2in 0in 0in,clip,width=.9\linewidth]{images/montage.pdf}  
\caption{\textbf{\gls{bfw}.} The average face of its different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender. The others represent the ethnicity and gender of the race and gender, respectfully. Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}
% \emph{Bottom-left} average per gender.  \emph{Top-left}: average face of \gls{bfw}. Notice the distinguishable variation in face features within and across subgroups. Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}
% \caption{\small{\textbf{Data of BFW.} Subgroups depicted by average faces. \emph{Top-left}: average of BFW. \emph{Top-right}: averages for races,  Asian (A), Black (B), Indian (I), White (W). \emph{Bottom-left}: averages for genders, Female (F) Male (M). \emph{Bottom-Right}: combined, \gls{af}, \gls{am}, \gls{bf}, \gls{bm}, \gls{if}, \gls{im}, \gls{wf}, \gls{wm}. 
% \emph{Top-left}: average face of \gls{bfw}.
% Notice the distinguishable variation in face features within and between subgroups.}}
\label{fig:avg-faces}
% \vspace{-3pt}
\end{figure}

Typically, \glspl{cnn} are trained on faces identified by a detection system. Specifically, for \gls{fr}, the goal is to project faces to an N-dimensional space with minimal distances between samples of the same identity and maximize the gap separating those that are different. Thus, the overarching goal is to discriminate between subjects with face encodings. We can deploy such a \gls{cnn} to encode faces to compare via a similarity score: genuine pair if the score is high enough; and rejected as an imposter.

A fixed threshold acts as a decision boundary in similarity score space. Thus, the face features in the same class must satisfy a criterion in the form of a single value~\cite{deng2019arcface, liu2017sphereface, wang2018additive, wang2018cosface}. However, a single, global threshold is a crude measure that leads to \gls{fr} errors. Furthermore, the held-out set used to determine the global threshold tends to share the same distribution with test data, which makes the results skewed in favor of specific demographics that make up a majority in both sets. That skew, the difference in the performance of an algorithm of certain ethnic groups, is our definition of bias. A key question is: \emph{is \gls{fr} too biased, or not?}

\begin{table*}[t!]
\scriptsize
    % \glsunset{bfw}
    % \glsunset{rfw}
    % \glsunset{dp}
    \centering
       \caption{\textbf{The proposed \gls{bfw} compared to related datasets.} \gls{bfw} is exactly balanced across ID, gender, and ethnicity (Table~\ref{tab:ethnic-splits}). Compared with \gls{dp}, \gls{bfw} provides more samples per subject and subgroups per set. Also, \gls{bfw} uses a single resource, VGG2. \gls{rfw}, on the other hand, supports a different task (\ie domain adaptation). Furthermore, \gls{rfw} focuses on race-distribution across of pairs, while not considering the distribution of identities. }
    \begin{tabular}{rccccccc}%x{10mm}x{3mm}x{10mm}x{8mm}}%\toprule

    \multicolumn{2}{c}{Database} & \multicolumn{3}{c}{Number of}& \multicolumn{3}{c}{Balanced Labels}\\
    \cmidrule(lr){1-2}	\cmidrule(lr){3-5} \cmidrule(lr){6-8}
    Name & Source Data & Faces &  IDs & Subgroups & ID & Ethnicity & Gender\\\midrule
    \gls{dp}~\cite{demogPairs}     & CASIA-W, VGG\&VGG2 & 10,800& 600 & 6 &\checkc& \checkc &\checkc \\
    \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &\approx80,000&\approx12,000& 4 & \xmark & \checkc &\xmark \\
    \gls{bfw} (ours) & VGG2 & 20,000 & 800 &8 & \checkc & \checkc &\checkc \\\bottomrule
    \end{tabular}

    \label{tab:compared}
    \vspace{-5mm}
\end{table*}

The adverse effects of a global threshold are two-fold: \textbf{(1)} the mapping produced by a \gls{cnn} is nonuniform. Therefore, distances between pairs of faces in different demographics vary in distribution of similarity scores (Fig~\ref{fig:detection-model}); \textbf{(2)} the evaluation set is imbalanced as well. Particular demographics making up a majority of the population will carry most weight on reported performance ratings. Reported results skew away from common traits to the underrepresented subgroups. Alas, demographics like gender, ethnicity, race, and age are underrepresented in most public datasets.

Making matters more challenging is that race and ethnicity are loosely defined.  For example, the US Census Bureau allows an individual to self-identify race (\href{https://www.census.gov/mso/www/training/pdf/race-ethnicity-onepager.pdf}{https://www.census.gov}). We define it as a group of people having facial characteristics similar to those found in a region. The result is various types of biases in FR systems in favor of or against particular demographics remain a question.

To address \textbf{(2)}, the lack of a balanced data, we introduce a new benchmark for \gls{fr}, called \gls{bfw} (Table~\ref{tab:compared}, ~\ref{tab:ethnic-splits}). \gls{bfw} serves as a platform to fairly evaluate \gls{fr} systems and enable demographic-specific ratings to be reported. We use \gls{bfw} to gain a deeper understanding of the extent of bias present in facial embeddings extracted from a \gls{soa} \gls{cnn} model. We then suggest a mechanism to counter the biased feature space to mitigate problems of bias with more balanced performance ratings for different demographics, while improving the overall accuracy. Specifically, we propose using an adaptive threshold that varies depending on the characteristics of detected facial attributes (\ie gender and ethnicity, Fig~\ref{fig:avg-faces}). We show an increase in accuracy with a balanced performance for different subgroups of people. Similarly, we show the positive effect of adjusting the similarity threshold based on the facial features of matched faces. Thus, selective use of similarity thresholds in current \gls{soa} \gls{fr} systems provides more intuition in \gls{fr}-based research, while providing a method easily adoptable in practice.

% Fig.~\ref{fig:avg-faces} depicts the average face of each subgroup.

\begin{table*}[!t]
    \centering
    \caption{\small{\textbf{Database stats and nomenclature, optimal thresholds ($t_o$), and accuracy scores.} \textit{Header:} Subgroup definitions. \textit{Top-row:} Statistics of \gls{bfw}. \textit{Middle-row:} Number of pairs for each partition. \textit{Bottom-row:} Accuracy when applying a global threshold $t_g$, the optimal threshold $t_o$, and accuracy with $t_o$ per subgroup. Columns grouped by race and then further split by gender. Out of millions of pairs, accuracy is inconsistent across subgroups. Furthermore, $F$ tend to perform inferior to that of $M$ (\ie up to 8\%).}}\label{tab:ethnic-splits}
    \include{tables/ethnic_splits}
     \vspace{-5mm}

\end{table*}

The contributions in this paper are as follows:\footnote{For data, code, more visit \scriptsize{\href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}}.}
\begin{enumerate}
    \item We built a balanced dataset as a proxy to measure verification performance per subgroup.
    \item We analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup significantly boost and balances performances.
    \item We showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
    \item We conducted human-evaluations to demonstrate bias in human perception. 
\end{enumerate}

### 2020-02-01 sections/relatedworks.tex
Initial Overleaf Import

\section{Background Information}
\subsection{Bias in \gls{ml}}
Although the community is excited about the progress and the commercial value of \gls{ml}, the trust between the society and techniques will take time to build due to the bias in algorithms. The exact definitions and implications of bias tend to vary between sources, as do its sources and types. One common theme is the tendency of bias to hinder performance ratings in ways that skew in favor of a particular subpopulation. At a high-level, bias can be introduced by humans~\cite{windmann1998subconscious}, data and label types~\cite{tommasi2017deeper}, \gls{ml} models~\cite{amini2019uncovering, kim2019learning}, and during evaluation~\cite{stock2018convnets}. For instance, a vehicle-detection model might miss cars if training set was mostly trucks. In practice, many \gls{ml} systems learn on biased datasets, which could harm our society.

% \vspace{-5pt}
\subsection{Bias in \gls{fr}}
Problems of bias in \gls{fr} have been driven by different motivations and have thereforem been solved by different means. To name a few: problems of data augmentation~\cite{yin2019feature}, one-shot learning~\cite{ding2018one}, demographic parity and fairness with priority on privacy~\cite{huang2018generative}, domain adaptation~\cite{wang2018racial}, differences in face-based attributes across demographics~\cite{wang2018they}, and even exploratory data analysis~\cite{muthukumar2019}. Yin et al. \cite{yin2019feature} proposed to augment the feature space of underrepresented classes using other classes with a diverse collection of samples to encourage distributions of underrepresented classes to more closely resemble that of the others. Similarly,~\cite{ding2018one} formulated the imbalanced class as one-shot learning and trained a \gls{gan} to generate face features as a means of augmenting class with as few as one sample. \cite{huang2018generative} proposed \gls{gapf} to create fair representations of the data in a quantifiable way, allowing for the finding of a decorrelation scheme from the data without access to its statistics. Wang et al. \cite{wang2018they} defined subgroups at a finer-level (\ie Chinese, Japanese, Korean) to determine the familiarity of faces across these subgroups. Genders have also been used as subgroups for work in bias, whether for efforts of analysis and understanding of gender-based face encodings~\cite{muthukumar2019}. Most recently,~\cite{wang2018racial} proposed adapting domains to bridge the gap between races by knowledge transfer, which was supported by a novel labeled data collection, \gls{rfw}. The release of \gls{rfw} came after the completion of \gls{bfw}. Although similar in terms of demographics (\ie Asian, Black, Indian, and White), \gls{rfw} uses faces from MSCeleb~\cite{guo2016ms} based on the assumption CASIA-Face~\cite{yi2014learning} and VGG2~\cite{Cao18} was used for training. In contrast, our \gls{bfw} samples from VGG2 and assumes MSCeleb was the training set. As shown in \cite{wang2018racial}, MSCeleb is highly imbalanced, primarily consisting of images of white individuals. For this, we expect bias from data, for this is the training set. Furthermore, our experiments allow for the parsing of results based on gender demographics (\ie although \gls{rfw} does separate males and females in each subgroup, no consideration for splits in just gender or just races are incorporated into the experimental design). Thus, \gls{rfw} and \gls{bfw} are complementary, with both adding metadata on demographics for subjects with faces in renowned, large-scale \gls{fr} datasets.

Most similar to our work are the recent efforts in~\cite{das2018, demogPairs, lopez2019dataset, srinivas2019face}. A common factor of these efforts are claims of insufficient data supply to support studies on bias in \gls{fr}, and the introduction of new metadata on demographics for image sets parsed out of existing collections. Specifically, \cite{demogPairs} curated a set of faces based on racial demographics (\ie Asian, Black, and White) called \gls{dp}, while~\cite{srinivas2019face} honed in on adults versus children called gls{itwcc}. Like the proposed \gls{bfw}, both were built by sampling existing databases, but with the addition of tags for the respective subgroups of interest. Besides, the additional data of \gls{bfw} (\ie added an additional subgroup \gls{i}, along with additional subjects with more faces for all subgroups), we also further split subgroups by gender. Furthermore, we focus on the problem of facial verification and the different levels of sensitivity in cosine similarity scores per subgroup.

% \vspace{-5pt}
\subsection{Human bias in \gls{ml}}
Bias is not unique to \gls{ml}, as humans also susceptible to a perceived bias across demographics: it exists across different races, genders, and even ages~\cite{10.1007/978-3-030-13469-3_68, bar2006, meissner2001, nicholls2018}. \cite{wang2018they} showed machines surpass human performance in classifying faces as Japanese, Chinese, or Korean by nearly 150\%. Precisely, humans barely pass random with 
38.89\% accuracy (\ie $\frac{1}{3}$ is random).

We expect human bias to skew results in favor of their genders and races. For this, we measure the human perception of faces across demographics in a controlled experiment. In the end, the results concur~\cite{wang2018they}-- we too observed an overall average below random (\ie $<$50\%). Furthermore, we provide details on settings and counts on the number of submissions per demographics while analyzing per demographic.

### 2020-02-01 sections/section3.tex
Initial Overleaf Import

\glsunset{if}\glsunset{im}\glsunset{af}\glsunset{am}\glsunset{bf}\glsunset{bm}\glsunset{wf}\glsunset{wm}
\section{The BFW Benchmark and Dataset}

We now discuss the \gls{bfw} dataset, and protocols to evaluate \gls{ml}-based \gls{fr}. We  conclude this section by reviewing the human evaluation conducted as part of this work, to detect bias in humans.

\subsection{The data}
Problems of bias in \gls{fr} motivated the design of \gls{bfw}. The data evenly represent various subgroups partitioned by demographics. Inspired by \gls{dp}~\cite{demogPairs}, the specification of \gls{bfw} follows in suit. \gls{bfw} includes additional subgroups (\ie \gls{if} and \gls{im}), an increased in the number subjects per subgroup, with many more pairs (Table~\ref{tab:ethnic-splits}).

\vspace{1mm}
\noindent\textbf{Compiling subject list.} 
 All subjects were sampled from VGG2~\cite{Cao18}. Thus, unlike others that depend on multiple sources, \gls{bfw} has less potential conflicts in train/test overlap for existing models. We used pre-trained ethnicity~\cite{ambekar2009name} and gender~\cite{levi2015age} classifiers to find candidates for the different subgroups.

\vspace{1mm}
\noindent\textbf{Detecting faces.} Faces were detected using \gls{mtcnn}~\cite{zhang2016joint}.\footnote{\href{https://github.com/polarisZhao/mtcnn-pytorch}{https://github.com/polarisZhao/mtcnn-pytorch}} Then, assigned into one of two sets. Faces within detected bounding box (BB) regions extended out 130\% in each direction, with zero-padding as the boundary condition made-up one set. The second set were faces aligned and cropped for Arcface~\cite{deng2019arcface} (see the next step). Also, coordinates of the BB and the five landmarks per MTCNN were stored as part of the static, raw data. Samples with multiple face detections had the BB area times the confidence score of the MTCNN to determine the instance most likely to be the face, with others set aside and labeled \textit{miss-detection}.

\begin{figure}[!t] 
	\centering    
 \glsunset{fpr}
  \glsunset{fnr}
	\includegraphics[trim=7mm 0.1in 5mm 0in,clip,width=\linewidth]{images/detcurve-improved.pdf}
		\caption{\textbf{\gls{det} curves.} \emph{Top-left}: per gender. \emph{Top-right}: per ethnicity. \emph{Bottom}: per subgroup (\ie combined). A dashed lined shows a difference by about 2$\times$ in \gls{fpr} for the same threshold $\theta_{const}$. \gls{fnr} is the number of match errors, so closer a curve is to the bottom better.}
		\glsreset{det}\glsreset{fpr}
\label{fig:detcurves} 
\vspace{-5mm}
\end{figure}

\vspace{1mm}
\noindent\textbf{Validating labels.} 
Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Arcface~\cite{deng2019arcface}. A matrix of cosine similarity scores was then generated for each subject, and removed samples (\ie rows) with median scores below threshold $\theta=0.2$. Mathematically, the $j^{th}$ subject with $N_j$ faces is removed if the ordinal rank of its score $n = \frac{P\times N}{100}\geq\theta$, with the percentile $P=50$, $\theta=0.2$ set manually, and an ordered list of scores. This allowed us to quickly prune \gls{fp} face detections. Following~\cite{robinson2016families, robinson2018visual}, we built a JAVA tool to visually validate the remaining faces. The faces were first ordered from most-to-least confidence, with confidence set as the average score, and then displayed as image icons on top toggling buttons arranged as a grid in a sliding pane window. Labeling then consisted of going subject-by-subject and flagging faces of \emph{imposters}. 
% Each subject underwent two passes by different annotators.

%  by flagging \emph{imposters}

\vspace{1mm}
\noindent\textbf{Sampling faces and creating folds.} We created lists of pairs in five-folds with subjects split evenly per subject and without overlap across folds. Furthermore, balance in the number of faces per was obtained by sampling twenty-five faces at random from each. Next, we generated a list of all face pairs per subject, resulting in $\sum_{l=1}^{L}\sum_{k=1}^{K_d} {N_k \choose 2}$ positive pairs, where the number of faces of all $K_l$ subjects $N_k=25$  for each of the $L$ subgroups (Table~\ref{tab:ethnic-splits}). Next, we assigned subjects to a fold. To preserve balance across folds, we sorted subjects by the number of pairs and then started assigning to alternating folds from the one with the most samples. Note, this left no overlap in identity between folds. Later, a negative set from samples within the same subgroup randomly matched until the count met that of the positive. Finally, we doubled the total count with negative pairs from across subgroups but in the same fold.
\begin{figure}[t!] 
	\glsunset{fpr}
	\centering
	\centering
	\includegraphics[trim=0.3in 0.1in 0.2in 0.2in,clip,angle=-0,origin=c,width=1\linewidth]{images/violinplots.pdf}
		\caption{\textbf{\Gls{sdm} across subgroups.} The scores for \emph{imposters} have medians about 0.3 but with variation in upper percentiles, while \emph{genuine} pairs vary in both (\eg \gls{af} has more area of overlap). A varying threshold varying across different subgroups yields a constant \gls{fpr}.} \label{fig:detection-model} 
\end{figure}

\vspace{-5pt}
\subsection{Problem formulation}\label{subsec:pf} 
% \gls{lfw}~\cite{LFWTech}, a commonly used benchmark for ~\gls{fr}, reserves specific train-test face-pair lists. 
\Gls{fv} is the special case of the two-class (\ie boolean) classification. Hence, pairs are labeled as the ``same'' or ``different'' \textit{genuine} pairs (\ie \textit{match}) or \textit{imposter} (\ie \textit{mismatch}), respectfully. This formulation (\ie \gls{fv}) is highly practical for applications like access control, re-identification, and surveillance. Typically, training a separate model for each unique subjects is infeasible. Firstly, the computational costs compound as the number of subjects increase.  Secondly, such a scheme would require model retraining each time a new person is added. Instead, we train models to encode facial images in a feature space that captures the uniqueness of a face, to then determine the outcome based on the output of a scoring (or distance) function. Formally put:
\begin{equation}\label{eg:matcher}
    f_{boolean}(\vec{x}_i, \vec{x}_j) = d(\vec{x}_i, \vec{x}_j) \leq \theta
\end{equation}

where $f_{boolean}$ is the \textit{matcher} of the feature vector $\vec{x}$ for the $i^{th}$ and $j^{th}$ sample~\cite{LFWTech}.

Cosine similarity is used as the \emph{matcher} in Eq~\ref{eg:matcher} the closeness of $i^{th}$ and $j^{th}$ features, \ie
$
s_l= \frac{f_i\cdot f_j}{||f_i||_2||f_j||_2}
$ is the closeness of the $l^{th}$ pair.

\subsection{Human Assessment\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}}\label{subsec:human-assessment}
We evaluated human on face pairs focusing on two racial groups: Chinese and Caucasians. To focus on the experiment, we honed-in on two groups, white Americans (W) and Chinese from China (C). The purpose was to the minimize variability by only analyzing the subsets of the broader groups of whites and Asians.

Samples were collected by recruiting subjects from multiple sources (\eg social media, email lists, and family/friends)-- a total of 120 participants were sampled at random from all the submissions that were (1) complete and (2) from a W or C participant. Specifically, there were 60 W and 60 C, both with \gls{m} and \gls{f} split evenly. A total of 50 face pairs of non-famous ``look-alikes'' were collected from the internet, with 20 ({\emph WA}) and 20 ({\emph C}) pairs (male and female split evenly). The other 10 pairs are of others (\eg Hispanic/ Latino, Japanese, African). Survey was created, distributed, and recorded via \href{https://paperform.co}{PaperForm}.

### 2020-02-01 supplemental.tex
Initial Overleaf Import

% 
% \newpage
\onecolumn
\renewcommand{\thesection}{\alph{section}}
\glsresetall
\setcounter{section}{0}

\section{Other \gls{cnn} models.}
Variations in optimal threshold are shown shown to generalize to other models (Fig.~\ref{fig:sdm-appendix-a}). Like in Fig.~\ref{fig:detcurves}, the \gls{det} curves for three \gls{cnn}-based models, each trained on VGG2 with softmax but with different backbones.\footnote{Used pre-trained models and public Github, \href{https://github.com/rcmalli/keras-vggface}{https://github.com/rcmalli/keras-vggface}} Notice similar trends across subgroups and models, which is consistent with  ArcFace as well. For example, the plots generated with ArcFace and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst).

\begin{figure}[h!]
    \centering
    \includegraphics[width=.9\linewidth, trim={0mm 0mm 0mm 0mm},clip]{images/SDM.pdf}\\
    \caption{\textbf{\gls{det} curves for different CNN models}. \gls{fnr} (\%) (vertical) vs \gls{fnp} (log-scale)  (horizontal) for VGG2~\cite{Cao18} models with different backbones (vgg16, Resnet50~\cite{he2016deep}, SEnet50~\cite{hu2018squeeze}). Lower is better. For each plot, \gls{wm} is the best performing curve, \gls{af} is the worst. The the ordering of the curves is roughly the same for each backbone. Very similar results are found for ArcFace in Fig.~\ref{fig:detcurves}.
    }\label{fig:sdm-appendix-a}
\end{figure}

Sample faces per subgroup of the \gls{bfw} dataset are shown in Fig.~\ref{fig:montage:app}.
\begin{figure}[h!]
    \centering
    \includegraphics[width=.8\linewidth]{images/facemontage.pdf}
    \caption{\textbf{Sample of \gls{bfw}}. Each row depicts a different gender, \gls{f} (top) and \gls{m} (bottom). Columns are grouped by ethnicity (\ie \gls{a}, \gls{b}, \gls{i}, and \gls{w}, respectfully).}
    \label{fig:montage:app}
\end{figure}

### 2020-02-01 tables/datastats.tex
Initial Overleaf Import

% \begin{table}[!t]
    % \centering
    % % \small{
    % \caption{Data statistics. The abbreviations for subgroups used throughout report.}
    \begin{tabular}{p{2.25cm} m{1cm}>{\centering} p{1.2cm}  m{1cm}>{\centering}} % Columns will widen to hold the content in them by default, to specify a precise width, use p{width}, e.g. p{5cm}
        % \toprule 
        % & \multicolumn{5}{c}{\textbf{Subgroups}} \\
        % \cmidrule(l){2-6} % Horizontal line spanning less than the full width of the table - you can add (r) or (l) just before the opening curly bracket to shorten the rule on the left or right side
       Subgroup &\# Faces &\# Subjects &\# Pairs\\ % Column names row
        \midrule
        \gls{af}&1,997 & 100   &52,980 \\ % row 1
        \gls{am}&2,002 & 100   &50,470  \\ % row 2
        \gls{bf}& 1,763& 100   &56,190  \\ % row 3
        \gls{bm}& 1,803&   100  &55,805 \\ % row 4
        \gls{if}& 1,902 &  100  &38,265 \\ % row 5
        \gls{im}& 1,890&  100   &41,560 \\ % row 6
        \gls{wf}&1,869 &  100   &59,705 \\ % row 7
        \gls{wm}&1,969 &  100   &58,715 \\ % row 819150 pairs for asian_females

        \midrule
        % \midrule
        Total & 15,195 &  799    & 413,690 \\ % Summary/total row
        % \bottomrule
    \end{tabular}
    % }
    % \label{tbl:datastats} 
% \end{table}

### 2020-02-01 tables/gender_splits.tex
Initial Overleaf Import

\begin{tabular}{l c c c c} % Columns will widen to hold the content in them by default, to specify a precise width, use p{width}, e.g. p{5cm}
    \toprule 
    & \multicolumn{2}{c}{\textbf{Females}} & \multicolumn{2}{c}{\textbf{Males}}        \\
    \cmidrule(l){2-3} \cmidrule(l){4-5} % Horizontal line spanning less than the full width of the table - you can add (r) or (l) just before the opening curly bracket to shorten the rule on the left or right side
     & A & B & A & B\\ % Column names row
    \midrule
    No. Positive & 28,354 & 28,354 & 29,479 & 29,479\\ % row 1
    No. Negative &39,970 & 129,389 & 53,535& 170,047 \\ % row 2
    Total & 68,324 & 157,743 &  83,014&199,526 \\ % row 3
    \midrule
    Threshold & 0.191 & 0.317 & 0.187 & 0.300 \\ % row 4
    Top Acc. & 0.9175 & 0.9287 & 0.9180 &0.9369 \\ % row 5
    % \bottomrule
\end{tabular}

### 2020-02-01 tmp.tex
Initial Overleaf Import

\begin{tabular}{lllr}\\\toprule\\ &   &    &     p1 \\ fold & label & type &        \\\midrule\\ 1 & 0 & bb &  34729 \\ &   & fd &  16605 \\ &   & fs &  27597 \\ &   & gfgd &    604 \\ &   & gfgs &    736 \\ &   & gmgd &    621 \\ &   & gmgs &    726 \\ &   & md &  13372 \\ &   & ms &  14996 \\ &   & sibs &  13344 \\ &   & ss &  34728 \\ & 1 & bb &  29275 \\ &   & fd &  13596 \\ &   & fs &  28108 \\ &   & gfgd &    916 \\ &   & gfgs &    854 \\ &   & gmgd &    711 \\ &   & gmgs &    565 \\ &   & md &  12924 \\ &   & ms &  32597 \\ &   & sibs &   8765 \\ &   & ss &  29275 \n2 & 0 & bb &  18880 \\ &   & fd &   7508 \\ &   & fs &   8461 \\ &   & gfgd &    401 \\ &   & gfgs &    403 \\ &   & gmgd &    829 \\ &   & gmgs &    848 \\ &   & md &   6915 \\ &   & ms &  10638 \\ &   & sibs &   6681 \\ &   & ss &  18882 \\ & 1 & bb &  28476 \\ &   & fd &   7888 \\ &   & fs &  12338 \\ &   & gfgd &   1181 \\ &   & gfgs &    309 \\ &   & gmgd &    960 \\ &   & gmgs &    407 \\ &   & md &   8474 \\ &   & ms &   9022 \\ &   & sibs &  10342 \\ &   & ss &  28476 \\3 & 0 & bb &  23981 \\ &   & fd &   9744 \\ &   & fs &  12237 \\ &   & gfgd &   1456 \\ &   & gfgs &   1014 \\ &   & gmgd &    391 \\ &   & gmgs &    386 \\ &   & md &   9314 \\ &   & ms &  15235 \\ &   & sibs &   7143 \\ &   & ss &  23997 \\ & 1 & bb &  19363 \\ &   & fd &   9369 \\ &   & fs &  11039 \\ &   & gfgd &    829 \\ &   & gfgs &    772 \\ &   & gmgd &    583 \\ &   & gmgs &    609 \\ &   & md &   8850 \\ &   & ms &   8937 \\ &   & sibs &   5264 \\ &   & ss &  19363 \\ 4 & 0 & bb &  14016 \\ &   & fd &   7262 \\ &   & fs &   9371 \\ &   & gfgd &    913 \\ &   & gfgs &    766 \\ &   & gmgd &    948 \\ &   & gmgs &    593 \\ &   & md &   8089 \\ &   & ms &  11026 \\ &   & sibs &   5149 \\ &   & ss &  14013 \\ & 1 & bb &  15838 \\ &   & fd &   7601 \\ &   & fs &   7202 \\ &   & gfgd &    337 \\ &   & gfgs &    662 \\ &   & gmgd &    269 \\ &   & gmgs &    552 \\ &   & md &   8367 \\ &   & ms &   6565 \\ &   & sibs &   7182 \\ &   & ss &  15838 \\ 5 & 0 & bb &  18671 \\ &   & fd &   8733 \\ &   & fs &  11421 \\ &   & gfgd &    690 \\ &   & gfgs &    431 \\ &   & gmgd &    664 \\ &   & gmgs &    381 \\ &   & md &   8970 \\ &   & ms &  12984 \\ &   & sibs &   7477 \\ &   & ss &  18661 \\ & 1 & bb &  17310 \\ &   & fd &  11385 \\ &   & fs &  10393 \\ &   & gfgd &    803 \\ &   & gfgs &    753 \\ &   & gmgd &    930 \\ &   & gmgs &    803 \\ &   & md &   8058 \\ &   & ms &   7768 \\ &   & sibs &   8238 \\ &   & ss &  17310 \\\bottomrule
\end{tabular}

### 2020-02-01 todo.tex
Initial Overleaf Import

--- 2020-02-01 notes.tex
@@
 \newpage
+\newpage\newpage
 
@@
     \item Reach out to IBM, for Diversity in Faces
-    \item Work with Joe to determine goals for data, i.e., specifications of what we would like to include and means of obtaining more if needed.
+    \item Work with Joe (and Gena if applicable, Nikil also if so) to determine goals for data, i.e., specifications of what we would like to include and means of obtaining more if needed.
+
 \end{enumerate}
@@
 \begin{enumerate}
-    \item TBD
+    \item Arrange to have quick powwow to get caught up with project, understand and help refine/improve goals and themes for this research track/ paper.
+Also, speak on what type of role you would like to play/ action items if any.
 \end{enumerate}
@@
 
-\begin{figure}[h]
-\setlength{\unitlength}{0.14in} % selecting unit length
-\centering % used for centering Figure
-\begin{picture}(32,15) % picture environment with the size (dimensions)
- % 32 length units wide, and 15 units high.
-\put(3,4){\framebox(6,3){$H_{B}(q)$}}
-\put(13,4){\framebox(6,3){$N[\cdot]$}}
-\put(23,4){\framebox(6,3){$H_{C}(q)$}}
-\put(0,5.5){\vector(1,0){3}}\put(9,5.5){\vector(1,0){4}}
-\put(19,5.5){\vector(1,0){4}}\put(29,5.5){\vector(1,0){3}}
-\put(-1,6.5) {$u(k)$}\put(30,6.5) {$y(k)$} \put(9.5,6.5)
-{$x_{B}(k)$}\put(19.5,6.5) {$x_{C}(k)$}
-\end{picture}
-\caption{An LNL Block Oriented Model Structure} % title of the Figure
-\label{fig:lnlblock} % label to refer figure in text
-\end{figure}
+% \begin{figure}[h]
+% \setlength{\unitlength}{0.14in} % selecting unit length
+% \centering % used for centering Figure
+% \begin{picture}(32,15) % picture environment with the size (dimensions)
+%  % 32 length units wide, and 15 units high.
+% \put(3,4){\framebox(6,3){$H_{B}(q)$}}
+% \put(13,4){\framebox(6,3){$N[\cdot]$}}
+% \put(23,4){\framebox(6,3){$H_{C}(q)$}}
+% \put(0,5.5){\vector(1,0){3}}\put(9,5.5){\vector(1,0){4}}
+% \put(19,5.5){\vector(1,0){4}}\put(29,5.5){\vector(1,0){3}}
+% \put(-1,6.5) {$u(k)$}\put(30,6.5) {$y(k)$} \put(9.5,6.5)
+% {$x_{B}(k)$}\put(19.5,6.5) {$x_{C}(k)$}
+% \end{picture}
+% \caption{An LNL Block Oriented Model Structure} % title of the Figure
+% \label{fig:lnlblock} % label to refer figure in text
+% \end{figure}
 
@@
 \endgroup
+\twocolumn

### 2020-02-01 sections/analysis.tex
version complete

@@
 	\centering    
-	\includegraphics[trim=.6in 1.4in 2.1in 1.65in,clip,width=.8\linewidth]{images/confusion.pdf}
+	\includegraphics[width=\linewidth]{images/confusion.pdf}
 		\caption{\small{\textbf{Confusion matrix.} Percent error (Rank 1) for all faces of \gls{bfw} queried against all others.
@@
 		\label{fig:confusion} 
+		\vspace{-5mm}
 \end{figure}
@@
 \section{Results and Analysis}
-% The \gls{soa} Arcface~\cite{deng2019arcface} was used throughout to control the model across all experiments.
+An {\em off-the-shelf} \gls{cnn} was used throughout to control the model across all experiments.
+For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{https://github.com/clcarwin/sphereface_pytorch}{https://github.com/clcarwin/sphereface_pytorch}} Following~\cite{liu2017sphereface}, upon applying an affine transformation to align faces according to pre-defined eye locations, each face got fed through the network twice, the original and horizontally flipped.
+The two features were fused by concatenation.
 
-We experimented with multiple \gls{cnn}-based models to support our claims of variation in optimal threshold across subgroups is not model-specific, but model-agnostic.
-Specifically, we analyze three \glspl{cnn} variants: we encode faces using VGG-16~\cite{simonyan2014very},  50-layer residual network (\ie ResNet-50)~\cite{he2016deep}, and \gls{senet} (\ie SENet-50)~\cite{hu2018squeeze}, with results of only the latter (\ie the best performing \gls{senet}) model used throughout the main paper.
+Additional \gls{cnn}-based models demonstrate the same phenomena: proportional to the overall model performance, exact in which the ordering subgroups in sensitivity in scores space (Appendix~\ref{app:sec:other:models}).
+% Specifically, we analyze three \glspl{cnn} variants: we encode faces using VGG-16~\cite{simonyan2014very},  50-layer residual network (\ie ResNet-50)~\cite{he2016deep}, and \gls{senet} (\ie SENet-50)~\cite{hu2018squeeze}, with results of only the latter (\ie the best performing \gls{senet}) model used throughout the main paper.
 Results for the other two, alongside \gls{senet} for comparison, are in the Supplemental Material.

### 2020-02-01 sections/introduction.tex
version complete

@@
     \include{tables/ethnic_splits}
-     \vspace{-5mm}
+    %  \vspace{-5mm}t
 
@@
 
-The contributions in this paper are as follows:\footnote{For data, code, more visit \scriptsize{\href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}}.}
+The contributions in this paper are as follows:\footnote{Data, code, more visit \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}.}
 \begin{enumerate}
@@
     \item We showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
-    \item We conducted human-evaluations to demonstrate bias in human perception. 
+    \item We conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
 \end{enumerate}

### 2020-02-01 sections/section3.tex
version complete

@@
 
-\subsection{Human Assessment\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}}\label{subsec:human-assessment}
+\subsection{Human Assessment}\label{subsec:human-assessment}
 We evaluated human on face pairs focusing on two racial groups: Chinese and Caucasians.

### 2020-02-01 supplemental.tex
version complete

@@
 
-\section{Other \gls{cnn} models.}
+\section{Other \gls{cnn} models.}~\label{app:sec:other:models}
 Variations in optimal threshold are shown shown to generalize to other models (Fig.~\ref{fig:sdm-appendix-a}).

### 2020-02-01 sections/analysis.tex
updated supplemental citations

@@
 The proposed shows improvement in all cases.}}\label{tab:ethnicy-far}
+ \vspace{-2mm}
 \include{tables/ethnicity_far}
@@
 \glsreset{far}
+ \vspace{-3mm}
+
 \end{table}
 
-The gender-based \gls{det} curve shows that there are different performances for \gls{m} and \gls{f} at a fixed threshold (dashed-line).
+The gender-based \gls{det} curve shows a difference in performances between \gls{m} and \gls{f} with a fixed threshold (dashed-line).
 Similar effects exist for the other curves as well (lines omitted to declutter).
-Many \gls{fr} applications, systems are operated at the highest \gls{fpr} allowed, so the line of constant threshold indicates that a single threshold produced at different operating points (\ie \gls{fpr}) \gls{m} and \gls{f}, which is undesirable.
-The difference in \gls{fpr} is approximately a factor of 2-- quite large indeed.
-If this kind of difference is common in industry, one would expect a factor of 3 more false positives to be reported based on one subgroup than another, which has a strong potential to be the cause of media articles reporting bias in
-\gls{fr}~\cite{england2019,snow2018}.
+For many \gls{fr} applications, systems are operated at the highest \gls{fpr} allowed, so the line of constant threshold indicates that a single threshold produces different operating points (\ie \gls{fpr}), which is undesirable.
+The difference in \gls{fpr} is approximately a factor of two-- quite large indeed.
+If this is the case in an industrial system, one would expect a difference in about double the false positives to be reported on based on subgroup alone.
+The potential ramifications of such a bias should not be overlooked, which it has not as of lately-- gaining interest of even main-stream media ~\cite{england2019,snow2018}.
 
@@
 	\centering    
-	\includegraphics[trim=0in 0.0in 0in 0in,clip, width=\linewidth] {images/human_eval.pdf}
+	\includegraphics[width=\linewidth] {images/human_eval.pdf}
 		\caption{\small{\textbf{Qualitative results of human assessment.} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}, which scores plotted next to each.

### 2020-02-01 supplemental.tex
updated supplemental citations

--- 2020-02-01 supplemental.tex
@@
 
-\section{Other \gls{cnn} models.}
-Variations in optimal threshold are shown shown to generalize to other models (Fig.~\ref{fig:sdm-appendix-a}).
-Like in Fig.~\ref{fig:detcurves}, the \gls{det} curves for three \gls{cnn}-based models, each trained on VGG2 with softmax but with different backbones.\footnote{Used pre-trained models and public Github, \href{https://github.com/rcmalli/keras-vggface}{https://github.com/rcmalli/keras-vggface}} Notice similar trends across subgroups and models, which is consistent with  ArcFace as well.
+\section{Other \gls{cnn} models.}~\label{app:sec:other:models}
+Variations in optimal threshold exists across different models (Fig.~\ref{fig:sdm-appendix-a}).
+Like in Fig.~\ref{fig:detcurves}, the \gls{det} curves for three \gls{cnn}-based models, each trained on VGG2 with softmax but with different backbones.\footnote{Used pre-trained models and public Github, \href{https://github.com/rcmalli/keras-vggface}{https://github.com/rcmalli/keras-vggface}} Notice similar trends across subgroups and models, which is consistent with  Sphereface as well (Fig.~\ref{fig:detcurves}).
 For example, the plots generated with ArcFace and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst).
-
 \begin{figure}[h!]
+\vspace{-2mm}
     \centering
-    \includegraphics[width=.9\linewidth, trim={0mm 0mm 0mm 0mm},clip]{images/SDM.pdf}\\
+    \includegraphics[width=.8\linewidth, trim={0mm 0mm 0mm 0mm},clip]{images/SDM.pdf}\\
     \caption{\textbf{\gls{det} curves for different CNN models}.
@@
 For each plot, \gls{wm} is the best performing curve, \gls{af} is the worst.
-The the ordering of the curves is roughly the same for each backbone.
-Very similar results are found for ArcFace in Fig.~\ref{fig:detcurves}.
-    }\label{fig:sdm-appendix-a}
+The the ordering of the curves is roughly the same for each backbone.}\label{fig:sdm-appendix-a}
+    \vspace{-3mm}
 \end{figure}
@@
 \begin{figure}[h!]
+\vspace{-2mm}
     \centering
@@
     \label{fig:montage:app}
+    \vspace{-3mm}
 \end{figure}
+
+\begin{thebibliography}{widest entry}
+ \bibitem[1]{Cao18} Cao, Qiong, Li Shen, Weidi Xie, Omkar M.
+Parkhi, and Andrew Zisserman.
+"Vggface2: A dataset for recognising faces across pose and age." In \textit{IEEE International Conference on Automatic Face & Gesture Recognition.} 2018.
+  \bibitem[2]{he2016deep} He, Kaiming, Xiangyu Zhang, Shaoqing Ren, and Jian Sun.
+"Deep residual learning for image recognition." In \textit{IEEE Conference on Computer Vision and Pattern Recognition.} 2016.
+ \bibitem[3]{hu2018squeeze} Hu, Jie, Li Shen, and Gang Sun. "Squeeze-and-excitation networks." In \textit{IEEE Conference on Computer Vision and Pattern Recognition.} 2018.
+\end{thebibliography}

### 2020-02-06 main.tex
version complete
updated supplemental citations

@@
 \usepackage{amsfonts}       % blackboard math symbols
-\usepackage{amssymb,amsthm}
+\usepackage{amssymb}
 \usepackage{amsmath}% assumes amsmath package installed
@@
 \usepackage{booktabs}       % professional-quality tables
-\usepackage{enumitem}
+%\usepackage{enumitem}
 \usepackage [english]{babel}
@@
 
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage[]{algorithm2e}
-\usepackage{amssymb}% http://ctan.org/pkg/amssymb
+%\usepackage{algorithm}
+%\usepackage{algorithmic}
+%\usepackage[]{algorithm2e}
+%\usepackage{amssymb}% http://ctan.org/pkg/amssymb
 \usepackage{pifont}% http://ctan.org/pkg/pifont
@@
 \newacronym{tp}{TP}{true-positive}
+\newacronym{fp}{FP}{false-positive}
+
 \newacronym{tpir}{TPIR}{true-positive identification rate}
@@
 \newacronym{eer}{EER}{Equal Error Rate}
-\newacronym{det}{DET}{Detection Error Trade-off}
 
@@
 \renewcommand{\arraystretch}{1.4}
-\newcommand{\xmark}{\ding{53}}%
+%\newcommand{\xmark}{\ding{53}}%
 
@@
 
-\def\FGPaperID{41} % *** Enter the FG 2019 Paper ID here
-
 \title{\LARGE \bf\vspace{-10mm}
@@
 }
-
-%use this in case of a single affiliation
-%\author{\parbox{16cm}{\centering
-%    {\large Huibert Kwakernaak}\\
-%    {\normalsize
-%    Faculty of Electrical Engineering, Mathematics and Computer Science, University of Twente, Enschede, The Netherlands\\}}
-%    \thanks{This work was not supported by any organization.}% <-this % stops a space
-%}
 
@@
     \hspace{-.4in}$^{1}$Northeastern University\hspace{.7in} $^{2}$ISM Connect}}
-    % $^1$ Northeastern University, Boston, MA, USA\\
-    % $^2$ ISM Connect, Cambridge, MA, USA}}
     \thanks{}% <-this % stops a space
@@
 
-% \author{Joseph P Robinson$^1$, Yuncheng Li$^{2}$, Ning Zhang$^{2}$, Yun Fu$^{1}$, and Sergey Tulyakov$^{2}$\\
-% \hspace{-.4in}$^{1}$Northeastern University\hspace{.7in} $^{2}$Snap Inc. \\\vspace{2mm}
-% } %\\
-
 \begin{document}
-\begin{bibunit}
+%\begin{bibunit}
 
@@
 Furthermore, we do a human evaluation to measure the bias in humans, which supports the hypothesis that such a bias exists in human perception.
+To download the \gls{bfw} database, source code, and more, visit \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}
 \end{abstract}
-
+\
 \input{sections/introduction.tex}
@@
 \begin{table}[t!]
+\begin{center}
     \caption{\small{\textbf{Quantitative of human assessment.} Different human subgroups listed per row.
@@
     \label{tab:humsn-eval-results} 
-\begin{center}
+     \vspace{-2mm}
+
 \footnotesize
@@
         &  CF  &      CM     &   WF     &   WM     &  Avg\\\midrule
-        CF &  \textbf{0.529}&  0.480&0.438&0.447 &0.474\pm0.041 \\
-        CM & 0.456 & \textbf{0.504}  & 0.444 &0.362 &0.441\pm0.059 \\
-        WF & 0.447 &0.438 &  \textbf{0.573}& 0.480 & 0.485\pm0.062 \\
-        WM & 0.301&0.474 &  0.453 & \textbf{0.561} & 0.447\pm0.108 \\\midrule
-        Avg &  0.433& 0.474&0.477 &0.463 &0.462\pm0.020\\
+        CF &  \textbf{0.529}&  0.480&0.438&0.447 &0.474$\pm$0.041 \\
+        CM & 0.456 & \textbf{0.504}  & 0.444 &0.362 &0.441$\pm$0.059 \\
+        WF & 0.447 &0.438 &  \textbf{0.573}& 0.480 & 0.485$\pm$0.062 \\
+        WM & 0.301&0.474 &  0.453 & \textbf{0.561} & 0.447$\pm$0.108 \\\midrule
+        Avg &  0.433& 0.474&0.477 &0.463 &0.462$\pm$0.020\\
  \end{tabular}}
  \end{center}
- \vspace{-5mm}
+ \vspace{-6mm}
 \end{table}
@@
 
-\putbib
-\end{bibunit}
+s\bibliographystyle{ieee}
+\bibliography{bias}
 
-% \newpage\newpage
-% \col
-\begin{bibunit}
-\input{supplemental.tex}
-\end{bibunit}
-% \end{comment}
+%\putbib
+%\end{bibunit}
+%\begin{bibunit}
+%\input{supplemental.tex}
+%\end{bibunit}
 

### 2020-02-06 sections/analysis.tex

@@
 An {\em off-the-shelf} \gls{cnn} was used throughout to control the model across all experiments.
-For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{https://github.com/clcarwin/sphereface_pytorch}{https://github.com/clcarwin/sphereface_pytorch}} Following~\cite{liu2017sphereface}, upon applying an affine transformation to align faces according to pre-defined eye locations, each face got fed through the network twice, the original and horizontally flipped.
+For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https:\/\/github.com\/clcarwin\/sphereface\_pytorch}} Following~\cite{liu2017sphereface}, upon applying an affine transformation to align faces according to pre-defined eye locations, each face got fed through the network twice, the original and horizontally flipped.
 The two features were fused by concatenation.
@@
  \vspace{-2mm}
-\include{tables/ethnicity_far}
+\begin{center}
+\begin{tabular}{l c c c c c}
+     \gls{far} & 0.3 & 0.1 & 0.01 & 0.001 & 0.0001\\\midrule
+    \multirow{2}{.1mm}{\textbf{\gls{af}}} &0.990 & 0.867 & 0.516 & 0.470 & 0.465\\[-4pt]
+        &1.000 & 0.882 & 0.524 & 0.478 & 0.474\\[-1pt]
+    \multirow{2}{3mm}{\textbf{\gls{am}}} &0.994 & 0.883 & 0.529 & 0.482 & 0.477\\[-4pt]
+        &1.000 & 0.890 & 0.533 & 0.486 & 0.482\\[-1pt]
+    \multirow{2}{3mm}{\textbf{\gls{bf}}} &0.991 & 0.870 & 0.524 & 0.479 & 0.473\\[-4pt]
+        &1.000 & 0.879 & 0.530 & 0.484 & 0.480\\[-1pt]
+    \multirow{2}{3mm}{\textbf{\gls{bm}}} &0.992 & 0.881 & 0.526 & 0.480 & 0.474\\[-4pt]
+        &1.000 & 0.891 & 0.532 & 0.485 & 0.480\\[-1pt]
+    \multirow{2}{3mm}{\textbf{\gls{if}}} &0.996 & 0.881 & 0.532 & 0.486 & 0.481\\[-4pt]
+        &1.000 & 0.884 & 0.534 & 0.488 & 0.484\\[-1pt]
+    \multirow{2}{3mm}{\textbf{\gls{im}}} &0.997 & 0.895 & 0.533 & 0.485 & 0.479\\[-4pt]
+        &1.000 & 0.898 & 0.535 & 0.486 & 0.481\\[-1pt]
+    \multirow{2}{3mm}{\textbf{\gls{wf}}} &0.988 & 0.878 & 0.517 & 0.469 & 0.464\\[-4pt]
+        &1.000 & 0.894 & 0.526 & 0.478 & 0.474\\[-1pt]
+    \multirow{2}{3mm}{\textbf{\gls{wm}}} &0.989 & 0.896 & 0.527 & 0.476 & 0.470\\[-4pt]
+        &1.000 & 0.910 & 0.535 & 0.483 & 0.478\\[-1pt]
+    \midrule
+    \multirow{2}{3mm}{\textbf{Avg.}} &0.992 & 0.881 & 0.526 & 0.478 & 0.474\\[-4pt]
+        &1.000 & 0.891 & 0.531 & 0.483 & 0.479\\[-10pt]
+\end{tabular}
+\end{center}
 \glsreset{tar}

### 2020-02-06 sections/introduction.tex
updated supplemental citations

@@
 For instance, \gls{fr} systems can be untrustworthy and racist in some cases~\cite{england2019,snow2018}.
-
-%The trust of people in machinery built to make decisions that range from casual-to-critical matters still takes time to build  .
+.
 
@@
 Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}
-% \emph{Bottom-left} average per gender.
-\emph{Top-left}: average face of \gls{bfw}.
-Notice the distinguishable variation in face features within and across subgroups.
-Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}
-% \caption{\small{\textbf{Data of BFW.} Subgroups depicted by average faces.
-\emph{Top-left}: average of BFW.
-\emph{Top-right}: averages for races,  Asian (A), Black (B), Indian (I), White (W).
-\emph{Bottom-left}: averages for genders, Female (F) Male (M).
-\emph{Bottom-Right}: combined, \gls{af}, \gls{am}, \gls{bf}, \gls{bm}, \gls{if}, \gls{im}, \gls{wf}, \gls{wm}.
-% \emph{Top-left}: average face of \gls{bfw}.
-% Notice the distinguishable variation in face features within and between subgroups.}}
 \label{fig:avg-faces}
@@
 \scriptsize
-    % \glsunset{bfw}
-    % \glsunset{rfw}
-    % \glsunset{dp}
     \centering
@@
     \gls{dp}~\cite{demogPairs}     & CASIA-W, VGG\&VGG2 & 10,800& 600 & 6 &\checkc& \checkc &\checkc \\
-    \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &\approx80,000&\approx12,000& 4 & \xmark & \checkc &\xmark \\
+    \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &$\approx$80,000&$\approx$12,000& 4 & \xmark & \checkc &\xmark \\
     \gls{bfw} (ours) & VGG2 & 20,000 & 800 &8 & \checkc & \checkc &\checkc \\\bottomrule
     \end{tabular}
-
     \label{tab:compared}
@@
 
-% Fig.~\ref{fig:avg-faces} depicts the average face of each subgroup.
-
 \begin{table*}[!t]
@@
 Furthermore, $F$ tend to perform inferior to that of $M$ (\ie up to 8\%).}}\label{tab:ethnic-splits}
-    \include{tables/ethnic_splits}
-    %  \vspace{-5mm}t
+    \scalebox{1}{
+     \resizebox{\textwidth}{!}{%
+    \begin{tabular}{r c c c c c c c c l}
+        \toprule
+        & \multicolumn{2}{c}{Asian (A)} & \multicolumn{2}{c}{Black (B)}  & \multicolumn{2}{c}{Indian (I)}& \multicolumn{2}{c}{White (W)}\\
+        \cmidrule(l){2-3} \cmidrule(l){4-5} \cmidrule(l){6-7}\cmidrule(l){8-9} % spanning less than the full width of the table - you can add (r) or (l) just before the opening curly bracket to shorten the rule on the left or right side
+         & Female (AF) & Male (AM) & BF & BM& IF & IM & WF & WM&Aggregated\\ % Column names row
+        \midrule
 
+       \# Faces  &  2,500&  2,500& 2,500 & 2,500& 2,500 & 2,500 & 2,500 & 2,500 &20,000 \\ % row 1
+        \# Subjects & 100& 100& 100  & 100  & 100  & 100& 100 &100&800  \\ % row 2
+        \# Faces / Subject  & 25 & 25    & 25 & 25 & 25  & 25  &  25 & 25 & 25\\ % row 3
+        % \midrule
+\specialrule{.01em}{.05em}{.05em}
+            \# Positive Pairs &  30,000&  30,000& 30,000 &30,000 & 30,000 &30,000&30,000 & 30,000 &240,000 \\ % row 1
+        \# Negative Pairs & 85,135&  85,232& 85,016  & 85,141  & 85,287  & 85,152& 85,223 &85,193&681,379  \\ % row 2
+
+        \# Pairs (Total) & 115,135 & 115,232    &115016 &115,141 & 115287  & 115,152  &  115,223& 115193 & 921,379\\ % row 3
+        % \midrule
+        \specialrule{.01em}{.05em}{.05em}
+        Acc$@t_g$ & 0.876 & 0.944  &0.934 &0.942 &0.922&0.949 &0.916  &0.918&0.925$\pm$0.022 \\
+        $t_o$ & 0.235 &  0.274 & 0.267&0.254  &0.299 & 0.295& 0.242 &0.222&0.261\\ % row 4
+        Acc$@t_o$ & 0.916 &0.964 &0.955 &0.971 &0.933 &0.958 & 0.969 &0.973 & 0.955 $\pm$ 0.018\\
+        \bottomrule
+    \end{tabular}}
+    \glsunset{af}
+    }
+    \glsunset{af}
+    \glsunset{am}
+    \glsunset{bf}
+    \glsunset{bm}
+    \glsunset{if}
+    \glsunset{im}
+    \glsunset{wf}
+    \glsunset{wm}
+    \vspace{-12pt}
 \end{table*}
 
-The contributions in this paper are as follows:\footnote{Data, code, more visit \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}.}
+The contributions in this paper are as follows:
 \begin{enumerate}

### 2020-02-06 sections/relatedworks.tex
updated supplemental citations

@@
 
-% \vspace{-5pt}
 \subsection{Bias in \gls{fr}}
-Problems of bias in \gls{fr} have been driven by different motivations and have thereforem been solved by different means.
+Problems of bias in \gls{fr} have been driven by different motivations and have therefore been solved by different means.
 To name a few: problems of data augmentation~\cite{yin2019feature}, one-shot learning~\cite{ding2018one}, demographic parity and fairness with priority on privacy~\cite{huang2018generative}, domain adaptation~\cite{wang2018racial}, differences in face-based attributes across demographics~\cite{wang2018they}, and even exploratory data analysis~\cite{muthukumar2019}.
@@
 
-% \vspace{-5pt}
 \subsection{Human bias in \gls{ml}}

### 2020-02-06 sections/analysis.tex
wip

@@
 
+% \vspace{-2mm}
 % \vspace{-2pt}
 \subsection{Score Analysis}
-Fig.~\ref{fig:detection-model} shows score distributions for faces of the same identity (\ie \emph{Genuine}) and different (\ie \emph{Imposter}), with \gls{sdm}s split by subgroup.
-Observe that scores of imposters  approximately distribute about a peak of zero independent of subgroup, with slightly different standard deviation.
-Observe that the broadness in score distribution of the \emph{genuine} pairs varies across subgroups.
+Fig.~\ref{fig:detection-model} shows score distributions for faces of the same (\ie \emph{Genuine}) and different (\ie \emph{Imposter}) identity, with a subgroup per \gls{sdm} plot.
+Notice that score distributions for imposters tends to peak about zero for all subgroups, and with minimal deviation comparing modes of the different plots.
+On the other hand, the score distribution of the \emph{genuine} pairs varies across subgroups in location (\ie score value) and spread (\ie overall shape).
 Asian Fem
@@
 \end{figure} 
-
+\vspace{-2mm}
 % \vspace{-5pt}

### 2020-02-06 .rebuttal/rebuttal.tex

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{fgrebuttal}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{subfig}
\usepackage{subcaption}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\usepackage{comment}
\usepackage{tabularx, booktabs, ragged2e}
\usepackage[acronym]{glossaries}
\usepackage{acronym}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

\def\cvprPaperID{41} 
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\newcommand{\xmark}{\ding{56}}%
\newcommand{\checkc}{\ding{51}}%
\newcommand{\NA}{---}
\newacronym{lfw}{LFW}{Labeled Faces in the Wild}
\newacronym{bfw}{BFW}{Balanced Faces In the Wild}
\newacronym{rfw}{RFW}{Racial Faces in-the-Wild:}
\newacronym{dp}{DemogPairs}{Demographic Pairs}

\newacronym{af}{AF}{\textit{Asian}-\textit{Female}}
\newacronym{am}{AM}{\textit{Asian}-\textit{Male}}
\newacronym{bf}{BF}{\textit{Black}-\textit{Female}}
\newacronym{bm}{BM}{\textit{Black}-\textit{Male}}
\newacronym{if}{IF}{\textit{Indian}-\textit{Female}}
\newacronym{im}{IM}{\textit{Indian}-\textit{Male}}
\newacronym{wf}{WF}{\textit{White}-\textit{Female}}
\newacronym{wm}{WM}{\textit{White}-\textit{Male}}
\newcommand{\com}[1]{\textcolor{red}{#1}}
\newcommand{\ie}{\textit{i}.\textit{e}., }
\newcommand{\eg}{\textit{e}.\textit{g}., }
\newcommand*{\etc}{etc.\@\xspace}

\providecommand{\cq}[1]{\textcolor{blue}{#1}}
\providecommand{\cqr}[1]{\sout{\textcolor{blue}{#1}}}

\begin{document}

\title{Face Recognition: Too Bias, or Not Too Bias?}

\maketitle

% \begin{figure*}[t!]
%     \centering
%         \includegraphics[width=.75\linewidth, trim={0mm 0mm 0mm 0mm},clip]{rebuttal/Figure1.pdf}
%     \caption{\textbf{DET curves for different CNN models}. FNR (vertical) vs FPR (horizontal) for different CNN-based models. The plots are displayed from lowest to best performance in terms of overall accuracy (\%) following the baseline (\ie global threshold). Notice that the better models have a larger variance in optimal thresholds from the global threshold.
%     }

%     \label{fig:fig1}
% \end{figure*}
\thispagestyle{empty}

\glsunset{bfw}
\glsunset{rfw}
\glsunset{dp}
% TODO DET curve: analyze threshold. Difference between models and global-- percent difference. Mention settings, simply how are they followed the same, but do state.

% \section{General Remarks}\label{sec:general}
% \noindent\textbf{Acknowledgement.} 
We thank reviewers for their time and comments. All reviewers agreed that bias in face recognition is important and well motivated in the paper. Two reviewers noted the value of the novel dataset to help find bias and improve performance. A reviewer noted the dataset will help motivate work in areas of bias, and another mentioned the urgency of finding a solution to bias and, thus, the importance of this work. Two reviewers noted that different optimal thresholds for subgroups is interesting, and the experiments showed it in results. A reviewer noted a thorough literature review. We begin by addressing four high-level issues noted by multiple reviewers, and then reply per reviewer:

\vspace{0.2mm}

\noindent\textbf{1. Additional CNN models} show bias results similar to ArcFace. As one example, (Fig.~\textcolor{red}{1}) shows DET curves (FNR vs FPR) for VggFace2 trained with 3 different backbones using (https://github.com/rcmalli/keras-vggface). Notice that the curves show similar trends across models, and consistent with  ArcFace (Fig. 2 in the paper). For example, the plots generated with ArcFace and VggFace2 all have the White Male curve at the bottom (best) and Asian Female on top (worst). In the paper, we initially chose to focus on ArcFace given the similarity of results to other networks. We will mention experiments with additional networks in paper, and refer readers to supplemental material for details.

\begin{table*}[t!]
\scriptsize
\glsreset{bfw}
\glsreset{rfw}
\glsreset{dp}
    \centering
    \begin{tabular}{rccccccc}%\toprule
    \multicolumn{2}{c}{Database} & \multicolumn{3}{c}{Number of}& \multicolumn{3}{c}{Balanced Labels}\\
    \cmidrule(lr){1-2}	\cmidrule(lr){3-5} \cmidrule(lr){6-8}
    Name & Source Data & Faces &  IDs & Subgroups & ID & Ethnicity & Gender\\\midrule
    \gls{dp}~\cite{demogPairs}     & CASIA-Web, VGG, VGG2 & 10,800& 600 & 6 &\checkc& \checkc &\checkc \\
    \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &\approx80,000&\approx12,000& 4 & \xmark & \checkc &\xmark \\
    \gls{bfw} (ours) & VGG2 & 20,000 & 800 &8 & \checkc & \checkc &\checkc \\\bottomrule
    \end{tabular}
    \caption{\textbf{\gls{bfw} compared to related datasets.} \gls{bfw} is exactly balanced across ID, gender, and ethnicity: 25 faces of 100 subjects of 8 ethnic groups generate a list of 240,000 positive pairs (\ie  30,000 per subgroup). It is an extension of \gls{dp}, providing more samples per subject and more subgroups, and without the issues of missing data that occur when building \gls{dp}, with a majority of black females from CASIA-Web. For the reduction of potential conflict between train and test sets, \gls{bfw} was built solely with VGG2. \gls{rfw}, on the other hand, supports a different task (\ie domain adaptation). Furthermore, \gls{rfw} is made up of pairs labeled by subgroup and pair (\ie \{{\em match}, {\em non-match}\}), while not considering distribution of identities. }
    \label{tab:table1}
    % \vspace{-5mm}
\end{table*}

% \begin{table*}[t!]
% \scriptsize
% \glsreset{bfw}
% \glsreset{rfw}
% \glsreset{dp}
%     \centering
%     \begin{tabular}{rcccccccl}%\toprule
%     \multicolumn{2}{c}{Database} & \multicolumn{3}{c}{Number of}& \multicolumn{3}{c}{Balanced Labels}&\hspace{8mm} Task\\
%     \cmidrule(lr){1-2}	\cmidrule(lr){3-5} \cmidrule(lr){6-8}
%     Name & Source Data & Faces &  IDs & Subgroups & ID & Ethnicity & Gender\\\midrule
%     \gls{dp}~\cite{demogPairs}     & CASIA-Web, VGG, VGG2 & 10,800& 600 & 6 &\checkc& \checkc &\checkc& ID Verification \\
%     \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &\approx80,000&\approx12,000& 4 & \xmark & \checkc &\xmark& Domain Adaptation \\
%     \gls{bfw} (ours) & VGG2 & 20,000 & 800 &8 & \checkc & \checkc &\checkc& ID Verification \\\bottomrule
%     \end{tabular}
%     \caption{\textbf{\gls{bfw} compared with related datasets.} \gls{bfw} is exactly balanced across ID, gender, and ethnicity. It is an extension of \gls{dp}, providing more samples per subject and the number of subgroups. Also, issues of missing data occur when building \gls{dp}, with a majority of black females from CASIA-Web. For this, the simplification, and the reduction of potential conflict between train and test sets, \gls{bfw} was built solely with VGG2. \gls{rfw}, on the other hand, supports a different task (\ie domain adaptation). Furthermore, \gls{rfw} is made up of pairs labeled by subgroup and pair (\ie \{{\em match}, {\em non-match}\}), while not including gender or considering the distribution of identities. In the end, \gls{bfw} is entirely balanced: 25 faces of 100 subjects of 8 ethnic groups are used to generate a list of 240,000 positive pairs (\ie  30,000 per subgroup).}
%     \label{tab:table1}
%     \vspace{-5mm}
% \end{table*}

\noindent\textbf{2. Data comparison} of the proposed \gls{bfw} with related datasets has been added to the paper (Table~\ref{tab:table1}). \gls{bfw} is designed such that each image includes identity, gender and ethnicity labels. Furthermore, there is an even number across all label-types. \gls{bfw}, unlike \gls{rfw}, splits ethnic-based subgroups by gender: subgroups based on ethnicity, on average, do vary with regards to optimal threshold. However, further splitting by gender yields gender-specific thresholds. Still, male faces are more separable than female counter-parts w.r.t subgroup, resulting in higher performances on average [line 620-622, col 1].

% (\ie \gls{dp}~\cite{demogPairs} and \gls{rfw}~\cite{wang2018racial}). 
\noindent\textbf{3. Details of human evaluation} regarding participants and survey were added to paper: To summarize, we recruited 120 subjects from multiple sources (\eg social media, email lists, and family/friends). Only white Americans ({\emph WA}) and Chinese from China ({\emph C}) were  targeted for this study. A total of 50 face pairs of non-famous subjects, with 20 ({\emph WA}) and 20 ({\emph C}) pairs (male and female split evenly). The other 10 pairs are of others (\eg Hispanic/ Latino, Japanese, African). Survey was created, distributed, and recorded via \href{https://paperform.co}{PaperForm}. The effort was supported by an IRB to ensure proper protocols were followed. A summary of results for every look-alike face pair is documented in the project repository, which will be made publicly available.

\noindent\textbf{4. Overarching Goal.} We created \gls{bfw} as a benchmark dataset for face verification of balanced data across subgroups. To aid this effort and ensure reproducibility, we created a Github repository, which includes easy-to-use notebook to regenerate all figures. We also have a project website with downloadable data, overview/ stats, and interactive gallery to browse \gls{bfw}. We are implementing \gls{bfw} as a Keras dataset for researchers. All will be released upon paper, publication: make entirely reproducible and maximize potential impact.

% The following sections are organized per reviewer. Please see the respective section for replies to individual remarks.

% \section{Review-Specific Rebuttal}
% \vspace{0.2mm}
\noindent\textbf{\textcolor{orange}{Reviewer1}}
\textbf{Q:} \textit{Analysis done with single CNN. The model bias can also be there to completely rely on the findings.Experiments using multiple CNNs would be great.} 
\textbf{A:} We have expanded the work with comparisons across multiple CNNs (Fig.~\textcolor{red}{1}). 
\textbf{Q:} \textit{How to improve performances of the FR system needs to be discussed.} 
\textbf{A:} We propose using variable thresholds based on the subgroup the individual belongs to. We compare the standard unique global threshold with the proposed variable threshold and demonstrate an improvement in performance [p6, Table II]. This shows significant improvement using our method (Table II). 
\textbf{Q:} \textit{We could do FR per subgroups as authors mention. Discussion or results from this for face verification (ROC) and face ID can be fruitful for research community.} 
\textbf{A:} We agree this is a promising research direction to build on work done here.

% \vspace{10mm}
\noindent\textbf{\textcolor{orange}{Reviewer2}}
\textbf{Q:} \textit{The presented threshold approach is evaluated in a very specific scenario. The extent of the evaluation is too narrow to derive general conclusions regarding bias in face verification neural networks. Only one network architecture is used to analyze the bias - ArcFace.} 
\textbf{A:} We have expanded the work with comparisons across multiple CNNs (Fig.~\textcolor{red}{1}). 
\textbf{Q:} \textit{ Only one similarity function is used to analyze the bias - cosine similarity.}  
\textbf{A:} We note that ArcFace produces L2-normalized embedding vectors for faces. As such, the cosine distance is proportional to the squared euclidean distance. Since distances are in the range [0, 1], the ranking of distances is preserved. As such, the metrics presented will not differ with euclidean distance, opposed to cosine distance. 
\textbf{Q:} \textit{Only one dataset is used to analyze the bias - BFW, introduced by the authors.} 
\textbf{A:} RFW provides balance in ethnicity labels and labels for pairs (\ie true or false). Thus, comparing results to another dataset is not possible, as we constructed this data by necessity (\ie RFW examines bias from the view of ethnicity classification, whereas we do identity verification).
Statistics and characteristics of the proposed \gls{bfw} to other face-based public datasets will be added to the paper to highlight the novelty in and need for data (Table~\ref{tab:table1}). 
\textbf{Q:} \textit{The different optimal thresholds for each subgroup is very interesting. However, the solution does not seem scalable to applications with growing, unbalanced datasets. Do we need to recompute when new identities are added to train set?} \textbf{A:} We note that if all subgroups are properly represented in the data when thresholds are computed, further data will not shift the overall distribution. We acknowledge that this method is difficult to employ when facing small datasets, as it may be difficult to compute the thresholds. However, this issue is encountered in many aspect of machine learning and isn't unique to our approach.
% \textbf{Q:} \textit{The human assessment experiment does not mention number of participants.} \textbf{A:} See (paragraph 3). 
\textbf{Q:} \textit{Face verification performance not compared to SOTA.} 
\textbf{A:} Our baseline is SOTA Arcface with a single global threshold, for which we show significant improvement over by using thresholds optimised per subgroup.
% facial verification with balance in ethnic backgrounds and gender-types. Thus, we propose this benchmark paper as a foundation for the research community.

\noindent\textbf{\textcolor{orange}{Reviewer3}}
\textbf{Q:} \textit{The paper is difficult to read and follow. There also exist several grammatical errors in the paper.} 
\textbf{A:} We have proofread and edited the paper to address this issue.
\textbf{Q:} \textit{they have proposed \gls{bfw} and that it is "complementary" to the existing \gls{rfw} dataset. It is not clear what additional contribution is being provided via the \gls{bfw} dataset, which extends the current literature. The authors must include a comparison with the \gls{rfw} dataset}
\textbf{A:} We have addressed this oversight by adding the requested comparison, see (Table~\ref{tab:table1}). We note that \gls{bfw} fundamentally differs from \gls{rfw} in its data and available types of label, although this was not made clear in our text. \gls{rfw} is intended as a benchmark in domain adaptation. Identity labels are not present, nor is a face verification benchmark. The task we address motivated us to build \gls{bfw}, a facial verification dataset with data balanced in terms of identity, gender, and ethnicity. \gls{bfw} supports the analysis and evaluation of 8 separate subgroups.
\textbf{Q:} \textit{There are no details with respect to training of Arcface when mentioned by authors.} 
\textbf{A:} We followed the same implementation as the authors of ArcFace, which achieves 99.83\% obtained on LFW. This will be clarified and detailed in paper.
% \textbf{Q:} \textit{In Section III.C, "Human performance was evaluated on face pairs focusing on two racial groups: Chinese and Caucasians." Where the data come from? Also, how many participants? How many pairs? How was study performed?}  \textbf{A:} See {\em General Remarks}, paragraph 3. 
\textbf{Q:} \textit{Fig. 4 presents the Rank-1 \% error. How is this calculated? Wasn't analysis drawn in terms of FAR and TAR?}  
\textbf{A:} FAR and TAR metrics are used to evaluate pairs of images. However, the Rank-1 analysis is not based on the pairs, but is the percent of faces for which the top-1 neighbor is of the same identity (effectively a top-1 accuracy). This analysis is performed within and across subgroups of gender and identity. We have explicitly stated and defined these metrics in the paper. 
\textbf{Q:} \textit{No clear description of how the category-based thresholds were obtained. A separate val set for obtaining thresholds?} 
\textbf{A:} We do a 5-fold cross validation protocol (Section IV.C, paragraph 2). Optimal thresholds were determined on train sets-- the threshold that resulted in lowest FPR w.r.t the desired FNR. All reported results are averaged across folds ($k=5$). Aforementioned will be updated and clearly stated.

{\footnotesize
\bibliographystyle{ieee}
\bibliography{rebuttal}
}
\end{document}

### 2020-02-06 supplemental.tex
wip

@@
     \centering
-    \includegraphics[width=.8\linewidth, trim={0mm 0mm 0mm 0mm},clip]{images/SDM.pdf}\\
+    \includegraphics[width=.8\linewidth, trim={0mm 0mm 0mm 0mm},clip]{figures/SDM.pdf}\\
     \caption{\textbf{\gls{det} curves for different CNN models}.
-\gls{fnr} (\%) (vertical) vs \gls{fnp} (log-scale)  (horizontal) for VGG2~\cite{Cao18} models with different backbones (vgg16, Resnet50~\cite{he2016deep}, SEnet50~\cite{hu2018squeeze}).
+\gls{fnr} (\%) (vertical) vs \gls{fpr} (log-scale)  (horizontal) for VGG2~\cite{Cao18} models with different backbones (vgg16, Resnet50~\cite{he2016deep}, SEnet50~\cite{hu2018squeeze}).
 Lower is better.
@@
     \centering
-    \includegraphics[width=.8\linewidth]{images/facemontage.pdf}
+    \includegraphics[width=.8\linewidth]{figures/facemontage.pdf}
     \caption{\textbf{Sample of \gls{bfw}}.
@@
 Parkhi, and Andrew Zisserman.
-"Vggface2: A dataset for recognising faces across pose and age." In \textit{IEEE International Conference on Automatic Face & Gesture Recognition.} 2018.
+"Vggface2: A dataset for recognising faces across pose and age." In \textit{IEEE International Conference on Automatic Face \& Gesture Recognition.} 2018.
   \bibitem[2]{he2016deep} He, Kaiming, Xiangyu Zhang, Shaoqing Ren, and Jian Sun.

### 2020-02-12 sections/analysis.tex
final changes for arxiv

@@
 	\centering    
-	\includegraphics[width=\linewidth]{images/confusion.pdf}
+	\includegraphics[width=\linewidth]{figures/confusion.pdf}
 		\caption{\small{\textbf{Confusion matrix.} Percent error (Rank 1) for all faces of \gls{bfw} queried against all others.
@@
 	\centering    
-	\includegraphics[width=\linewidth] {images/human_eval.pdf}
-		\caption{\small{\textbf{Qualitative results of human assessment.} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}, which scores plotted next to each.
-Humans tend to be more successful at recognizing their own subgroup, with few exceptions (bottom).
-}}
+	\includegraphics[width=\linewidth] {figures/human_eval.pdf}
+		\caption{\small{\textbf{Qualitative results of human assessment.} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}.
+Accuracy scores shown as bar plots.
+Humans are most successful at recognizing their own subgroup, with a few exceptions (\eg bottom).}}
 		% (\eg bottom row). } Result summary in Table~\ref{tab:human-eval}.

### 2020-02-12 sections/introduction.tex
final changes for arxiv

@@
 \centering
-         \includegraphics[trim=0in 0.2in 0in 0in,clip,width=.9\linewidth]{images/montage.pdf}  
+         \includegraphics[trim=0in 0.2in 0in 0in,clip,width=.9\linewidth]{figures/montage.pdf}
 \caption{\textbf{\gls{bfw}.} The average face of its different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender.
@@
 \textit{Top-row:} Statistics of \gls{bfw}.
-\textit{Middle-row:} Number of pairs for each partition.
-\textit{Bottom-row:} Accuracy when applying a global threshold $t_g$, the optimal threshold $t_o$, and accuracy with $t_o$ per subgroup.
+\textit{Middle:} Number of pairs for each partition.
+\textit{Bottom:} Accuracy with a global threshold $t_g$, the optimal threshold $t_o$, and accuracy with $t_o$ per subgroup.
+Specifically, $t_g$=0.259$\pm$0.002 when averaged across the five folds.
 Columns grouped by race and then further split by gender.
 Out of millions of pairs, accuracy is inconsistent across subgroups.
-Furthermore, $F$ tend to perform inferior to that of $M$ (\ie up to 8\%).}}\label{tab:ethnic-splits}
+Furthermore, $F$ tend to perform inferior to that of $M$ for all but B.}}\label{tab:ethnic-splits}
     \scalebox{1}{
@@
         \specialrule{.01em}{.05em}{.05em}
-        Acc$@t_g$ & 0.876 & 0.944  &0.934 &0.942 &0.922&0.949 &0.916  &0.918&0.925$\pm$0.022 \\
-        $t_o$ & 0.235 &  0.274 & 0.267&0.254  &0.299 & 0.295& 0.242 &0.222&0.261\\ % row 4
-        Acc$@t_o$ & 0.916 &0.964 &0.955 &0.971 &0.933 &0.958 & 0.969 &0.973 & 0.955 $\pm$ 0.018\\
+        Acc$@t_g$ & 0.941 & 0.952  & 0.967 & 0.960 & 0.958 &0.962  & 0.973 & 0.981 &0.962 $\pm$0.012 \\
+        $t_o$ & 0.255 &  0.246 & 0.277 &0.242  &0.297 & 0.264& 0.233 &0.239 &0.257$\pm$0.006\\ % row 4
+
+        Acc$@t_o$ & 0.942 &0.952 &0.968 &0.961 &0.960 &0.962 & 0.974 &0.982 & 0.963 $\pm$ 0.012\\
         \bottomrule

### 2020-02-15 manuscript.tex
name changes to more generic filenames
wip
Updates from Overleaf

--- 2020-02-06 main.tex
@@
 \usepackage{makecell}
-
+\usepackage{balance}
 % \usepackage[small]{caption}
@@
 
+% \usepackage{minipage}
 \usepackage{array}
@@
 \newacronym{nmse}{NMSE}{Normalized Mean Square Error}
-\newacronym{det}{DET}{Detection Error Trade-off}
+\newacronym{det}{DET}{Detection error trade-off}
 \newacronym{tp}{TP}{true-positive}
@@
 \newacronym{fpr}{FPR}{false-positive rate}
+
 \newacronym{tpr}{TPR}{true-positive rate}
@@
 
-\title{\LARGE \bf\vspace{-10mm}
+\title{\LARGE \bf
 Face Recognition: Too Bias, or Not Too Bias?
 }
+
+% \title{\LARGE \bf\vspace{-10mm}
+% Face Recognition: Too Bias, or Not Too Bias?
+% }
 
@@
 We reveal critical insights into problems of bias in state-of-the-art \gls{fr} systems using a novel \gls{bfw} dataset: data balanced for gender and ethnic groups.
-Classic signal detection theory revealed trends in the underlying score distribution across subgroups.
-Specifically, we show variations in the optimal scoring threshold varies for face-pairs across different subgroups.
+We show variations in the optimal scoring threshold for face-pairs across different subgroups.
 Thus, the conventional approach of learning a  global threshold for all pairs resulting in performance gaps among subgroups.
@@
 Furthermore, we do a human evaluation to measure the bias in humans, which supports the hypothesis that such a bias exists in human perception.
-To download the \gls{bfw} database, source code, and more, visit \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}
+For the \gls{bfw} database, source code, and more, visit \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}
 \end{abstract}
@@
     \label{tab:humsn-eval-results} 
-     \vspace{-2mm}
-
+    %  \vspace{-2mm}
 \footnotesize
@@
  \end{center}
- \vspace{-6mm}
+%  \vspace{-7mm}
 \end{table}
@@
 \glsresetall
-\section{CONCLUSION}
+\section{Conclusion}
 We introduce a new data set \gls{bfw} with eight subgroups balanced across gender and ethnicity.
-With this, and upon highlighting the challenges and shortcomings of grouping subjects as a single subset, we provide evidence that forming subgroups is meaningful, as the \gls{fr} algorithm rarely makes mistakes across subgroups.
-We trained Arc-Face net on MSCeleb, expecting that the results would suffer from bias because of the imbalanced train-set.
-Once established that the results do suffer from problems of bias, we observed that the same threshold across ethnic and gender subgroups leads to differences in the \gls{fpr} up to a factor of two, which is seemingly the cause of the frenzy about bias in \gls{fr} in main-stream media.
-Furthermore, we ameliorate these differences with a per-subgroup threshold, leveling out \gls{fpr}, and achieving a higher \gls{tpr}.
+With this, and upon highlighting the challenges and shortcomings of grouping subjects as a single subset, we provide evidence that forming subgroups is meaningful, as the FR algorithm rarely makes mistakes across subgroups.
+We used an \textit{off-the-shelf} Sphereface, hypothesizing this SOTA CNN suffers from bias because of the imbalanced train-set.
+Once established that the results do suffer from problems of bias, we observed that the same threshold across ethnic and gender subgroups leads to differences in the FPR up to a factor of two, which is seemingly the cause of the frenzy about bias in FR in main-stream media.
+Furthermore, we ameliorate these differences with a per-subgroup threshold, leveling out FPR, and achieving a higher TPR.
 We hypothesized that most humans grown amongst more than their own demographic and, therefore, effectively learn from imbalanced datasets-- a human evaluation validated that humans are biased, as most recognized their personal demographic best.
@@
 
-s\bibliographystyle{ieee}
-\bibliography{bias}
+% \vspace{-3mm}
+\bibliographystyle{ieee}
+\balance
+\bibliography{references}
 
-%\putbib
-%\end{bibunit}
-%\begin{bibunit}
-%\input{supplemental.tex}
-%\end{bibunit}
+% \putbib
+% \end{bibunit}
+\begin{bibunit}
+\input{supplemental.tex}
+\end{bibunit}
 

### 2020-02-15 sections/analysis.tex
Updates from Overleaf

@@
 		\label{fig:confusion} 
-		\vspace{-5mm}
 \end{figure}
@@
 An {\em off-the-shelf} \gls{cnn} was used throughout to control the model across all experiments.
-For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https:\/\/github.com\/clcarwin\/sphereface\_pytorch}} Following~\cite{liu2017sphereface}, upon applying an affine transformation to align faces according to pre-defined eye locations, each face got fed through the network twice, the original and horizontally flipped.
-The two features were fused by concatenation.
+For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https://github.com/clcarwin/sphereface\_pytorch}} Following~\cite{liu2017sphereface}, upon applying an affine transformation to align faces according to pre-defined eye locations, each face got fed through the network twice, the original and horizontally flipped.
+The two features were fused by average pooling (\ie 512 D).
 
@@
 
-% \vspace{-2mm}
-% \vspace{-2pt}
-\subsection{Score Analysis}
+\subsection{Score analysis}
 Fig.~\ref{fig:detection-model} shows score distributions for faces of the same (\ie \emph{Genuine}) and different (\ie \emph{Imposter}) identity, with a subgroup per \gls{sdm} plot.
@@
 The proposed shows improvement in all cases.}}\label{tab:ethnicy-far}
- \vspace{-2mm}
+%  \vspace{-2mm}
 \begin{center}
@@
 \glsreset{far}
- \vspace{-3mm}
+%  \vspace{-3mm}
 
@@
 More importantly, each subgroup has the desired \gls{fpr}, so that substantial differences in \gls{fpr} will remain unfounded.
-We further experimented using ethnicity estimators and using the ethnicity of both query and test image, which yielded similar results to those reported here (results not shown).
+We further experimented using ethnicity estimators and using the ethnicity of both query and test image, which yielded similar results to those reported here.
 
@@
 		\label{fig:human-eval} 
-		\vspace{-5mm}
+% 		\vspace{-1mm}
 \end{figure} 
-\vspace{-2mm}
+% \vspace{-4mm}
 % \vspace{-5pt}
@@
  Quantitative and qualitative results are in Table~\ref{tab:humsn-eval-results} and Fig.~\ref{fig:human-eval}, respectfully.
-One might expect that the most exposure to others would be within the same subgroup, and, therefore, would be best at labeling their own.
-Secondarily, they would be best at labeling images of the same ethnicity, but opposite gender.
+Subjects of a subgroup likely have mostly been exposured to others of the same.
+Therefore, it is expected they would be best at labeling their own; similar for the same ethnicity, but other gender.
 Our findings concur.

### 2020-02-15 sections/introduction.tex
Updates from Overleaf

@@
 \centering
-         \includegraphics[trim=0in 0.2in 0in 0in,clip,width=.9\linewidth]{figures/montage.pdf}
+         \includegraphics[trim=0in 0.2in 0in 0in,clip,width=\linewidth]{figures/montage.pdf}
 \caption{\textbf{\gls{bfw}.} The average face of its different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender.
@@
 \label{fig:avg-faces}
-% \vspace{-3pt}
 \end{figure}
@@
     \label{tab:compared}
-    \vspace{-5mm}
+    % \vspace{-4mm}
 \end{table*}
@@
     \glsunset{wm}
-    \vspace{-12pt}
+    % \vspace{-12pt}
 \end{table*}
@@
 \begin{enumerate}
-    \item We built a balanced dataset as a proxy to measure verification performance per subgroup.
-    \item We analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup significantly boost and balances performances.
-    \item We showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
-    \item We conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
+    \item Built a balanced dataset as a proxy to measure verification performance per subgroup.
+    \item Analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup significantly boost and balances performances.
+    \item Showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
+    \item Conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
 \end{enumerate}

### 2020-02-15 sections/relatedworks.tex
Updates from Overleaf

@@
 A common factor of these efforts are claims of insufficient data supply to support studies on bias in \gls{fr}, and the introduction of new metadata on demographics for image sets parsed out of existing collections.
-Specifically, \cite{demogPairs} curated a set of faces based on racial demographics (\ie Asian, Black, and White) called \gls{dp}, while~\cite{srinivas2019face} honed in on adults versus children called gls{itwcc}.
+Specifically, \cite{demogPairs} curated a set of faces based on racial demographics (\ie Asian, Black, and White) called \gls{dp}, while~\cite{srinivas2019face} honed in on adults versus children called \gls{itwcc}.
 Like the proposed \gls{bfw}, both were built by sampling existing databases, but with the addition of tags for the respective subgroups of interest.

### 2020-02-15 sections/section3.tex
updated supplemental citations
cropped margins from PDFs
Updates from Overleaf

@@
 Inspired by \gls{dp}~\cite{demogPairs}, the specification of \gls{bfw} follows in suit.
-\gls{bfw} includes additional subgroups (\ie \gls{if} and \gls{im}), an increased in the number subjects per subgroup, with many more pairs (Table~\ref{tab:ethnic-splits}).
+\gls{bfw} includes additional subgroups (\ie \gls{if} and \gls{im}), an increase in the number subjects per subgroup, and with many more pairs (Table~\ref{tab:ethnic-splits}).
 
@@
 Faces within detected bounding box (BB) regions extended out 130\% in each direction, with zero-padding as the boundary condition made-up one set.
-The second set were faces aligned and cropped for Arcface~\cite{deng2019arcface} (see the next step).
-Also, coordinates of the BB and the five landmarks per MTCNN were stored as part of the static, raw data.
-Samples with multiple face detections had the BB area times the confidence score of the MTCNN to determine the instance most likely to be the face, with others set aside and labeled \textit{miss-detection}.
+The second set were faces aligned and cropped for Sphereface~\cite{liu2017sphereface} (see the next step).
+Also, coordinates of the BB and the five landmarks from \gls{mtcnn} were stored as part of the static, raw data.
+For samples with multiple face detections, we used the BB area times the confidence score of the \gls{mtcnn} to determine the face most likely to be the subject of interest, with the others set aside and labeled \textit{miss-detection}.
 
@@
   \glsunset{fnr}
-	\includegraphics[trim=7mm 0.1in 5mm 0in,clip,width=\linewidth]{images/detcurve-improved.pdf}
+	\includegraphics[width=\linewidth]{figures/detcurve-improved.pdf}
 		\caption{\textbf{\gls{det} curves.} \emph{Top-left}: per gender.
@@
 \label{fig:detcurves} 
-\vspace{-5mm}
 \end{figure}
@@
 \noindent\textbf{Validating labels.} 
-Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Arcface~\cite{deng2019arcface}.
+Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Sphereface~\cite{liu2017sphereface}.
 A matrix of cosine similarity scores was then generated for each subject, and removed samples (\ie rows) with median scores below threshold $\theta=0.2$.
-Mathematically, the $j^{th}$ subject with $N_j$ faces is removed if the ordinal rank of its score $n = \frac{P\times N}{100}\geq\theta$, with the percentile $P=50$, $\theta=0.2$ set manually, and an ordered list of scores.
+Mathematically, the $n^{th}$ sample for the $j^{th}$ subject with $N_j$ faces was removed if the ordinal rank of its score $n = \frac{P\times N}{100}\geq\theta$, with the percentile $P=50$, fell below a threshold of $\theta=0.2$ (set manually).
 This allowed us to quickly prune \gls{fp} face detections.
 Following~\cite{robinson2016families, robinson2018visual}, we built a JAVA tool to visually validate the remaining faces.
-The faces were first ordered from most-to-least confidence, with confidence set as the average score, and then displayed as image icons on top toggling buttons arranged as a grid in a sliding pane window.
+For this, the faces were ordered by decreasing confidence, with confidence set as the average score, and then displayed as image icons on top toggling buttons arranged as a grid in a sliding pane window.
 Labeling then consisted of going subject-by-subject and flagging faces of \emph{imposters}.
-% Each subject underwent two passes by different annotators.
-
-%  by flagging \emph{imposters}
 
@@
 	\centering
-	\includegraphics[trim=0.3in 0.1in 0.2in 0.2in,clip,angle=-0,origin=c,width=1\linewidth]{images/violinplots.pdf}
+	\includegraphics[width=1\linewidth]{figures/violinplots.pdf}
 		\caption{\textbf{\Gls{sdm} across subgroups.} The scores for \emph{imposters} have medians about 0.3 but with variation in upper percentiles, while \emph{genuine} pairs vary in both (\eg \gls{af} has more area of overlap).
 A varying threshold varying across different subgroups yields a constant \gls{fpr}.} \label{fig:detection-model}
+
 \end{figure}
 
-\vspace{-5pt}
+% \vspace{-5pt}
 \subsection{Problem formulation}\label{subsec:pf} 
@@
 
-\subsection{Human Assessment}\label{subsec:human-assessment}
+\subsection{Human assessment}\label{subsec:human-assessment}
 We evaluated human on face pairs focusing on two racial groups: Chinese and Caucasians.

### 2020-02-15 supplemental.tex
Updates from Overleaf

@@
 %
-% \newpage
+\newpage
 \onecolumn
+% \begin{minipage*}[\linewidth]{width of the minipage}
 \renewcommand{\thesection}{\alph{section}}
@@
 \section{Other \gls{cnn} models.}~\label{app:sec:other:models}
-Variations in optimal threshold exists across different models (Fig.~\ref{fig:sdm-appendix-a}).
+Variations in optimal threshold exist across models (Fig.~\ref{fig:sdm-appendix-a}).
 Like in Fig.~\ref{fig:detcurves}, the \gls{det} curves for three \gls{cnn}-based models, each trained on VGG2 with softmax but with different backbones.\footnote{Used pre-trained models and public Github, \href{https://github.com/rcmalli/keras-vggface}{https://github.com/rcmalli/keras-vggface}} Notice similar trends across subgroups and models, which is consistent with  Sphereface as well (Fig.~\ref{fig:detcurves}).
-For example, the plots generated with ArcFace and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst).
+For example, the plots generated with Spherface and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst).
+% ArcFace (not shown) shown similar results. We would have to cite + will incorporate later .... a blank statement and sphereface is comparable ... up to you i do not care
+% \begin{figure}[h!]
+% % \vspace{-2mm}
+%     \centering
+%     \includegraphics[width=.8\linewidth, trim={0mm 0mm 0mm 0mm},clip]{figures/SDM.pdf}\\
+%     \caption{\textbf{\gls{det} curves for different CNN models}.
+\gls{fnr} (\%) (vertical) vs \gls{fpr} (log-scale)  (horizontal) for VGG2~\cite{Cao18} models with different backbones (vgg16, Resnet50~\cite{he2016deep}, SEnet50~\cite{hu2018squeeze}).
+Lower is better.
+For each plot, \gls{wm} is the best performing curve, \gls{af} is the worst.
+The the ordering of the curves is roughly the same for each backbone.}\label{fig:sdm-appendix-a}
+%     \vspace{-3mm}
+% \end{figure}
+
 \begin{figure}[h!]
-\vspace{-2mm}
+\vspace{-1mm}
     \centering
-    \includegraphics[width=.8\linewidth, trim={0mm 0mm 0mm 0mm},clip]{figures/SDM.pdf}\\
+    \begin{subfig}
+    \includegraphics[width=.3\linewidth]{figures/curve_vgg16_subgroups-crop.pdf}
+    \end{subfig}
+    \begin{subfig}
+    \includegraphics[width=.28\linewidth]{figures/curve_resnet50_subgroups-crop.pdf}
+    \end{subfig}
+        \begin{subfig}
+    \includegraphics[width=.28\linewidth]{figures/curve_senet50_subgroups-crop.pdf}
+    \end{subfig}
     \caption{\textbf{\gls{det} curves for different CNN models}.
@@
 The the ordering of the curves is roughly the same for each backbone.}\label{fig:sdm-appendix-a}
-    \vspace{-3mm}
+    \vspace{-1mm}
 \end{figure}
@@
     \centering
-    \includegraphics[width=.8\linewidth]{figures/facemontage.pdf}
+    \includegraphics[width=.75\linewidth]{figures/facemontage.pdf}
     \caption{\textbf{Sample of \gls{bfw}}.
@@
     \label{fig:montage:app}
-    \vspace{-3mm}
+    \vspace{-1mm}
 \end{figure}
-
+{
+\scriptsize
 \begin{thebibliography}{widest entry}
@@
 \end{thebibliography}
+}
+
+% \end{minipage*}

### 2020-02-16 sections/analysis.tex
Version for arXiv

@@
 \section{Results and Analysis}
-An {\em off-the-shelf} \gls{cnn} was used throughout to control the model across all experiments.
-For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https://github.com/clcarwin/sphereface\_pytorch}} Following~\cite{liu2017sphereface}, upon applying an affine transformation to align faces according to pre-defined eye locations, each face got fed through the network twice, the original and horizontally flipped.
-The two features were fused by average pooling (\ie 512 D).
+A single \gls{cnn} was used as a means to control the experiments.
+For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https://github.com/clcarwin/sphereface\_pytorch}} As reported in~\cite{wang2019racial}, \gls{lfw} has about a 13\%, 14\%, 3\%, and 70\% ratio in Asians, Blacks, Indians, and Whites, respectively; furthermore, CASIA-Web is even more unbalanced (again, as reported in~\cite{wang2019racial}), with about  3\%, 11\%, 2\%, and 85\% for the same subgroups.
 
@@
  Quantitative and qualitative results are in Table~\ref{tab:humsn-eval-results} and Fig.~\ref{fig:human-eval}, respectfully.
-Subjects of a subgroup likely have mostly been exposured to others of the same.
+Subjects of a subgroup likely have mostly been exposed to others of the same.
 Therefore, it is expected they would be best at labeling their own; similar for the same ethnicity, but other gender.
@@
 On average, subgroups are comparable at labeling images.
+Analogous to the \gls{fr} system, performance ratings differ when analyzing within and across subgroups.
+In other words, performance on \gls{bfw} improved with subgroup-specific thresholds.
+Similarly, humans tend to better recognize individuals by facial cues of the same or similar demographics.
+Put differently, as the recognition performances drops with a global threshold optimized for one subgroup and deployed on another, human performance tends to drop when across subgroups (\ie most familiar with the same subgroup; performance seems to drop when faced with another, less familiar subgroup).
+
+%  for it better adapts wh specific to subgroup) shift in sensitivity across subgroups when deep encodings are compared, the underlying mechanism

### 2020-02-16 sections/section3.tex
Version for arXiv

@@
 \noindent\textbf{Validating labels.} 
-Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Sphereface~\cite{liu2017sphereface}.
-A matrix of cosine similarity scores was then generated for each subject, and removed samples (\ie rows) with median scores below threshold $\theta=0.2$.
-Mathematically, the $n^{th}$ sample for the $j^{th}$ subject with $N_j$ faces was removed if the ordinal rank of its score $n = \frac{P\times N}{100}\geq\theta$, with the percentile $P=50$, fell below a threshold of $\theta=0.2$ (set manually).
+Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Sphereface~\cite{liu2017sphereface}: applying an affine transformation to align faces according to pre-defined eye locations, each face fed through twice (\ie the original and horizontally flipped), with the final encoding obtained by fusing the two features by average pooling (\ie 512 D).
+A matrix of cosine similarity scores was then generated for each subject, and removed samples (\ie rows) with median scores below threshold $\theta=0.2$ (set manually).
+Mathematically, the $n^{th}$ sample for the $j^{th}$ subject with $N_j$ faces was removed if the ordinal rank of its score $n = \frac{P\times N}{100}\geq\theta$, where $P=50$.
+In other words, the median (\ie 50 percentile) of all scores for a faces with respect to all of faces for the respective subject must pass a threshold of $\theta=0.2$; else, the face is dropped.
 This allowed us to quickly prune \gls{fp} face detections.

### 2020-02-16 supplemental.tex
arXiv

@@
     \centering
-    \begin{subfig}
-    \includegraphics[width=.3\linewidth]{figures/curve_vgg16_subgroups-crop.pdf}
-    \end{subfig}
-    \begin{subfig}
-    \includegraphics[width=.28\linewidth]{figures/curve_resnet50_subgroups-crop.pdf}
-    \end{subfig}
-        \begin{subfig}
-    \includegraphics[width=.28\linewidth]{figures/curve_senet50_subgroups-crop.pdf}
-    \end{subfig}
+    \begin{subfigure}[t]{.3\linewidth}
+    \includegraphics[width=.75\linewidth]{figures/curve_vgg16_subgroups-crop.pdf}
+    \caption{VGG16~\cite{simonyan2014very}}
+ \end{subfigure}
+    \begin{subfigure}[t]{.27\linewidth}
+    \includegraphics[width=.8\linewidth]{figures/curve_resnet50_subgroups-crop.pdf}
+    \caption{ResNet50~\cite{he2016deep}}
+   \end{subfigure}
+    \begin{subfigure}[t]{.27\linewidth}
+    \includegraphics[width=.8\linewidth]{figures/curve_senet50_subgroups-crop.pdf}
+    \caption{SENet~\cite{hu2018squeeze}}
+    \end{subfigure}
     \caption{\textbf{\gls{det} curves for different CNN models}.
-\gls{fnr} (\%) (vertical) vs \gls{fpr} (log-scale)  (horizontal) for VGG2~\cite{Cao18} models with different backbones (vgg16, Resnet50~\cite{he2016deep}, SEnet50~\cite{hu2018squeeze}).
+\gls{fnr} (\%) (vertical) vs \gls{fpr} (log-scale)  (horizontal) for VGG2~\cite{Cao18} models with different backbones (VGG16, Resnet50, SEnet50).
 Lower is better.
@@
     \centering
-    \includegraphics[width=.75\linewidth]{figures/facemontage.pdf}
+    \includegraphics[width=.7\linewidth]{figures/facemontage.pdf}
     \caption{\textbf{Sample of \gls{bfw}}.
@@
 \begin{thebibliography}{widest entry}
- \bibitem[1]{Cao18} Cao, Qiong, Li Shen, Weidi Xie, Omkar M.
+ \bibitem[1]{Cao18} Qiong Cao, Li Shen, Weidi Xie, Omkar M.
 Parkhi, and Andrew Zisserman.
-"Vggface2: A dataset for recognising faces across pose and age." In \textit{IEEE International Conference on Automatic Face \& Gesture Recognition.} 2018.
-  \bibitem[2]{he2016deep} He, Kaiming, Xiangyu Zhang, Shaoqing Ren, and Jian Sun.
-"Deep residual learning for image recognition." In \textit{IEEE Conference on Computer Vision and Pattern Recognition.} 2016.
- \bibitem[3]{hu2018squeeze} Hu, Jie, Li Shen, and Gang Sun. "Squeeze-and-excitation networks." In \textit{IEEE Conference on Computer Vision and Pattern Recognition.} 2018.
+``Vggface2: A dataset for recognising faces across pose and age.'' In \textit{IEEE International Conference on Automatic Face \& Gesture Recognition.} 2018.
+ \bibitem[2]{simonyan2014very} Karen Simonyan and Andrew Zisserman. ``Very deep convolutional networks for large-scale image recognition'' \textit{arXiv preprint arXiv:1409.1556.} 2014.
+  \bibitem[3]{he2016deep} He, Kaiming, Xiangyu Zhang, Shaoqing Ren, and Jian Sun.
+``Deep residual learning for image recognition.'' In \textit{IEEE Conference on Computer Vision and Pattern Recognition.} 2016.
+ \bibitem[4]{hu2018squeeze} Hu, Jie, Li Shen, and Gang Sun. ``Squeeze-and-excitation networks.'' In \textit{IEEE Conference on Computer Vision and Pattern Recognition.} 2018.
 \end{thebibliography}

### 2020-03-11 cvprworkshop.tex
added cvpr template and started filling in

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
% \usepackage[breaklinks=true,bookmarks=false]{hyperref}
\usepackage{sidecap}
\usepackage{wrapfig}
\usepackage[subrefformat=parens,labelformat=parens]{subcaption} %% provides subfigure
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{float}% If comment this, figure moves
\usepackage{geometry}

% \usepackage{FG2019}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage[acronym]{glossaries}
\usepackage{acronym}

\newcommand{\minitab}[2][c]{\begin{tabular}{#1}#2\end{tabular}}

% symbols for 1/2, etc.
% \usepackage{microtype}      % microtypography
% \usepackage{amsfonts}       % blackboard math symbols
% \usepackage[utf8]{inputenc} % allow utf-8 input
% \usepackage[T1]{fontenc}    % use 8-bit T1 fonts
% % \usepackage{hyperref}       % hyperlinks
% \usepackage[colorlinks = true,
%             linkcolor = blue,
%             urlcolor  = blue,
%             citecolor = blue,
%             anchorcolor = blue]{hyperref}
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \usepackage{mathptmx} % assumes new font selection scheme installed
% \usepackage{lipsum}
% \usepackage{slashbox}
\usepackage{booktabs}       % professional-quality tables
%\usepackage{enumitem}
\usepackage [english]{babel}
\usepackage [autostyle, english = american]{csquotes}
\MakeOuterQuote{"}
\usepackage[font=itshape]{quoting}
\usepackage{lipsum}
% \usepackage{subfig}
%% add
\usepackage{graphics}
\usepackage{xcolor}
\usepackage{color, colortbl}
\usepackage{comment}
\usepackage{tabularx, booktabs, ragged2e}
% \usepackage{arydshln}
% \usepackage{todonotes}
% \usepackage{ctable}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{balance}

\usepackage{pifont}% http://ctan.org/pkg/pifont

\graphicspath{{figures/}} %Setting the graphicspath
\usepackage{bibunits}

\defaultbibliography{bias}
\defaultbibliographystyle{ieee_fullname}

\usepackage[acronym]{glossaries}
\usepackage{acronym}

\usepackage{array}
\newcolumntype{x}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
\newcommand{\ie}{\textit{i}.\textit{e}., }
\newcommand{\eg}{\textit{e}.\textit{g}., }
\newcommand*{\etc}{etc.\@\xspace}

\newcommand{\xmark}{\ding{56}}%
\newcommand{\checkc}{\ding{51}}%
\newcommand{\NA}{---}

\newacronym{ml}{ML}{machine learning}
\newacronym{fr}{FR}{facial recognition}
\newacronym{fv}{FV}{facial verification}

\newacronym{cnn}{CNN}{convolutional neural network}
\newacronym{nn}{NN}{neural network}
\newacronym{mtcnn}{MTCNN}{\emph{multi-task \gls{cnn}}}

\newacronym{gan}{GAN}{generative adversarial network}
\newacronym{se}{SE}{\emph{Squeeze-and-Excitation}}
\newacronym{d}{$D$}{discriminator}
\newacronym{g}{$G$}{generator}
\newacronym{dbvae}{DB-VAE}{Debiasing Variational Autoencoder}

\newacronym{lut}{LUT}{Look-Up-Table}
\newacronym{soa}{SOTA}{state-of-the-art}

\newacronym{fiw}{FIW}{Families In the Wild}
\newacronym{lfw}{LFW}{Labeled Faces in the Wild}
\newacronym{bfw}{BFW}{Balanced Faces In the Wild}
\newacronym{rfw}{RFW}{Racial Faces in-the-Wild:}
\newacronym{dp}{DemogPairs}{Demographic Pairs}
\newacronym{itwcc}{ITWCC}{Wild Child Celebrity}

\newacronym{m}{M}{\textit{Male}}
\newacronym{f}{F}{\textit{Female}}
\newacronym{a}{A}{\textit{Asian}}
\newacronym{b}{B}{\textit{Black}}
\newacronym{i}{I}{\textit{Indian}}
\newacronym{w}{W}{\textit{White}}
\newacronym{af}{AF}{\textit{Asian}-\textit{Female}}
\newacronym{am}{AM}{\textit{Asian}-\textit{Male}}
\newacronym{bf}{BF}{\textit{Black}-\textit{Female}}
\newacronym{bm}{BM}{\textit{Black}-\textit{Male}}
\newacronym{if}{IF}{\textit{Indian}-\textit{Female}}
\newacronym{im}{IM}{\textit{Indian}-\textit{Male}}
\newacronym{wf}{WF}{\textit{White}-\textit{Female}}
\newacronym{wm}{WM}{\textit{White}-\textit{Male}}

\newacronym{fd}{FD}{Face Discrimination}
\newacronym{bb}{BB}{bounding box}

\newacronym{sdm}{SDM}{signal detection model}
\newacronym{roc}{ROC}{receiver operating characteristic}
\newacronym{nmse}{NMSE}{Normalized Mean Square Error}
\newacronym{det}{DET}{Detection error trade-off}
\newacronym{tp}{TP}{true-positive}
\newacronym{fp}{FP}{false-positive}

\newacronym{tpir}{TPIR}{true-positive identification rate}
\newacronym{frir}{FRIR}{false-reject identification rate}
\newacronym{fpir}{FRIR}{false-positive identification rate}

\newacronym{fn}{FN}{false-negative}
\newacronym{frr}{FRR}{false-reject rate}
\newacronym{fnr}{FNR}{false-negative rate}

\newacronym{fpr}{FPR}{false-positive rate}

\newacronym{tpr}{TPR}{true-positive rate}

\newacronym{tar}{TAR}{True Acceptance Rate}
\newacronym{far}{FAR}{False Acceptance Rate}
\newacronym{eer}{EER}{Equal Error Rate}

\newacronym{cs}{CS}{Cosine Similiarity}

\newacronym{lime}{LIME}{Local Interpretable Model-Agnostic Explanations}
\newacronym{nas}{NAS}{Neural Architecture Search}
\newacronym{gapf}{GAPF}{Generative Adversarial Privacy and Fairness}

                                                         % paper
%\setlist[itemize]{leftmargin=*}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\newcommand{\mc}[2]{\multicolumn{#1}{c}{#2}}
\definecolor{Gray}{gray}{0.85}
\definecolor{LightCyan}{rgb}{0.88,1,1}

\newcolumntype{a}{>{\columncolor{Gray}}c}
\newcolumntype{b}{>{\columncolor{white}}c}
\renewcommand{\arraystretch}{1.4}
%\newcommand{\xmark}{\ding{53}}%

\newcommand{\highlightb}[1]{%
\colorbox{blue!30}{$\displaystyle#1$}}
\newcommand{\vo}{\vec{o}\@ifnextchar{^}{\,}{}}

\newcommand{\vx}{\vec{x}\@ifnextchar{^}{\,}{}}

% \let\vec\mathbf

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{0002} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
%\setcounter{page}{4321}
\begin{document}

%%%%%%%%% TITLE
\title{Face Recognition: Too Bias, or Not Too Bias?}

% \author{First Author\\
% Institution1\\
% Institution1 address\\
% {\tt\small firstauthor@i1.org}
% % For a paper whose authors are all at the same institution,
% % omit the following lines up until the closing ``}''.
% % Additional authors and addresses can be added with ``\and'',
% % just like the second author.
% % To save space, use either the email address or home page, not both
% \and
% Second Author\\
% Institution2\\
% First line of institution2 address\\
% {\tt\small secondauthor@i2.org}
% }

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
We reveal critical insights into problems of bias in state-of-the-art \gls{fr} systems using a novel \gls{bfw} dataset: data balanced for gender and ethnic groups. We show variations in the optimal scoring threshold for face-pairs across different subgroups. Thus, the conventional approach of learning a  global threshold for all pairs resulting in performance gaps among subgroups. By learning subgroup-specific thresholds, we not only mitigate problems in performance gaps but also show a notable boost in the overall performance. Furthermore, we do a human evaluation to measure the bias in humans, which supports the hypothesis that such a bias exists in human perception. For the \gls{bfw} database, source code, and more, visit \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{bibunit}

\glsresetall
\section{Introduction}
    As more of society becomes integrated with \gls{ml}, bias, fairness, and the formalization of \gls{ml} standards attract more attention~\cite{10.1007/978-3-030-13469-3_68, anne2018women, wang2018racial}. Thus, an effect of the growing dependency on \gls{ml} are growing concerns for biased and unfair algorithms, \eg untrustworthy and racist \gls{fr} systems~\cite{england2019,snow2018}.

    Typically, \glspl{cnn} are trained on faces identified by a detection system. Specifically, for \gls{fr}, the goal is to encode faces in an N-dimensional space such that samples of the same identity are neighbors and those of different identity of further apart. Thus, the goal is to identify subjects with face encodings. We can deploy a \gls{cnn} to encode faces, and then compare via a similarity score: assume pairs with a high enough score are \emph{genuine}; else, an \emph{imposter}.

% \begin{figure}
%     \centering
%     \includegraphics[width=\linewidth]{figures/fig1-crop.pdf}
%     \caption{Caption}
%     \label{fig:my_label}
% \end{figure}
    \begin{figure}[t!]
        \centering
        \includegraphics[width=.9\linewidth]{figures/montage.pdf}
        \caption{\textbf{\gls{bfw}.} The average face of its different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender. The others represent the ethnicity and gender of the race and gender, respectively. Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}
        \label{fig:avg-faces}
    \end{figure}

    When comparing scores, typically, a fixed threshold sets the decision boundary. Thus, features of the same identity must satisfy a criterion via a single value~\cite{deng2019arcface, liu2017sphereface, wang2018additive, wang2018cosface}. However, we found that a single (\ie global) threshold is a crude measure that leads to skewed errors - the held-out set used to determine the threshold tends to share the same distribution with the test data, which favors specific demographics that are a majority. That skew, the difference in the performance of an algorithm of certain demographics, is our definition of bias. A key question is: \emph{is \gls{fr} too biased, or not?}

\begin{table*}[!t]
    \centering
    \caption{\small{\textbf{Database stats and nomenclature, optimal thresholds ($t_o$), and accuracy scores.} \textit{Header:} Subgroup definitions. \textit{Top-row:} Statistics of \gls{bfw}. \textit{Middle:} Number of pairs for each partition. \textit{Bottom:} Accuracy with a global threshold $t_g$, the optimal threshold $t_o$, and accuracy with $t_o$ per subgroup. Specifically, $t_g$=0.259$\pm$0.002 when averaged across the five folds. Columns grouped by race and then further split by gender. Out of millions of pairs, accuracy is inconsistent across subgroups. Furthermore, $F$ tends to perform inferior to that of $M$ for all but B.}}\label{tab:ethnic-splits}
    \scalebox{.75}{
     \resizebox{\textwidth}{!}{%
    \begin{tabular}{r c c c c c c c c l}
        \toprule
        & \multicolumn{2}{c}{Asian (A)} & \multicolumn{2}{c}{Black (B)}  & \multicolumn{2}{c}{Indian (I)}& \multicolumn{2}{c}{White (W)}\\
        \cmidrule(l){2-3} \cmidrule(l){4-5} \cmidrule(l){6-7}\cmidrule(l){8-9} % spanning less than the full width of the table - you can add (r) or (l) just before the opening curly bracket to shorten the rule on the left or right side
         & Female (AF) & Male (AM) & BF & BM& IF & IM & WF & WM&Aggregated\\ % Column names row
        \midrule

       \# Faces  &  2,500&  2,500& 2,500 & 2,500& 2,500 & 2,500 & 2,500 & 2,500 &20,000 \\ % row 1
        \# Subjects & 100& 100& 100  & 100  & 100  & 100& 100 &100&800  \\ % row 2
        \# Faces / Subject  & 25 & 25    & 25 & 25 & 25  & 25  &  25 & 25 & 25\\ % row 3
        % \midrule
\specialrule{.01em}{.05em}{.05em}
            \# Positive Pairs &  30,000&  30,000& 30,000 &30,000 & 30,000 &30,000&30,000 & 30,000 &240,000 \\ % row 1
        \# Negative Pairs & 85,135&  85,232& 85,016  & 85,141  & 85,287  & 85,152& 85,223 &85,193&681,379  \\ % row 2

        \# Pairs (Total) & 115,135 & 115,232    &115016 &115,141 & 115287  & 115,152  &  115,223& 115193 & 921,379\\ % row 3
        % \midrule
        % \specialrule{.01em}{.05em}{.05em}
        % Acc$@t_g$ & 0.941 & 0.952  & 0.967 & 0.960 & 0.958 &0.962  & 0.973 & 0.981 &0.962 $\pm$0.012 \\
        % $t_o$ & 0.255 &  0.246 & 0.277 &0.242  &0.297 & 0.264& 0.233 &0.239 &0.257$\pm$0.006\\ % row 4

        % Acc$@t_o$ & 0.942 &0.952 &0.968 &0.961 &0.960 &0.962 & 0.974 &0.982 & 0.963 $\pm$ 0.012\\
        \bottomrule
    \end{tabular}}
    % \glsunset{af}
    }
    % % \glsunset{af}
    % \glsunset{am}
    % \glsunset{bf}
    % \glsunset{bm}
    % \glsunset{if}
    % \glsunset{im}
    % \glsunset{wf}
    % \glsunset{wm}
    % \vspace{-12pt}
\end{table*}

    Making matters more challenging is that race and ethnicity are loosely defined. For example, the US Census Bureau allows an individual to self-identify race.\footnote{\scriptsize\href{https://www.census.gov/mso/www/training/pdf/race-ethnicity-onepager.pdf}{www.census.gov/mso/www/training/pdf/race-ethnicity-onepager}} For this work, we define subgroups as specific sub-populations with face characteristics similar to those found in a region.

    The adverse effects of a global threshold are two-fold: \textbf{(1)} the mapping produced by a \gls{cnn} is nonuniform. Therefore, distances between pairs of faces in different demographics vary in distribution of similarity scores (Fig~\ref{fig:detection-model}); \textbf{(2)} the evaluation set is imbalanced as well. Particular demographics making up a majority of the population will carry most weight on the reported performance ratings. Reported results favor the common traits over the underrepresented. Alas, demographics like gender, ethnicity, race, and age are underrepresented in most public datasets~\cite{merler2019diversity, wang2018racial}.

    % The result is various types of biases in FR systems in favor of or against particular demographics remain a question.

    For \textbf{(1)} we propose subgroup-specific (\ie optimal) thresholds; to address \textbf{(2)}, we introduce a new benchmark for measuring bias in \gls{fr}, \gls{bfw} (Table~\ref{tab:compared} \&~\ref{tab:ethnic-splits}). \gls{bfw} allows for a fair evaluation of \gls{fr} systems, while enabling demographic-specific ratings to be reported. We use \gls{bfw} to gain a deeper understanding of the extent of bias present in \gls{soa} \gls{cnn} used to encode faces. Then, we suggest a mechanism to mitigate problems of bias with more balanced performance ratings for different demographics. Specifically, we propose using an adaptive threshold that varies depending on the characteristics of detected facial attributes (\ie gender and ethnicity, Fig~\ref{fig:avg-faces}). We show an increase in accuracy with a balanced performance for different subgroups. Similarly, we show a positive effect of adjusting the similarity threshold based on the facial features of matched faces. Thus, selective use of similarity thresholds in current \gls{soa} \gls{fr} systems provides more intuition in \gls{fr} research with a method easy to adopt in practice.

    The contributions in this paper are 3-fold: (1) Built a balanced dataset as a proxy to measure verification performance per subgroup; (2) Analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup balances performance; (3) Showed bias causes skewness in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds; (4) Conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
    % \begin{enumerate}
    %     \item Built a balanced dataset as a proxy to measure verification performance per subgroup.
    %     \item Analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup significantly boost and balance performances.
    %     \item Showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
    %     \item Conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
    % \end{enumerate}

    \begin{table*}[t!]

        \centering
        \caption{\textbf{\gls{bfw} compared to related datasets.} Our \gls{bfw} is balanced across ID, gender, and ethnicity (Table~\ref{tab:ethnic-splits}). Compared with \gls{dp}, \gls{bfw} provides more samples per subject and subgroups per set, while only using a single resource, VGG2. \gls{rfw}, on the other hand, supports a different task (\ie domain adaptation), and focuses on race-distribution, but not the distribution of identities.}
        \scriptsize
        \begin{tabular}{rccccccc}%x{10mm}x{3mm}x{10mm}x{8mm}}%\toprule

            \multicolumn{2}{c}{Database} & \multicolumn{3}{c}{Number of}& \multicolumn{3}{c}{Balanced Labels}\\
            \cmidrule(lr){1-2}	\cmidrule(lr){3-5} \cmidrule(lr){6-8}
            Name & Source Data & Faces &  IDs & Subgroups & ID & Ethnicity & Gender\\\midrule
            \gls{dp}~\cite{demogPairs}     & CASIA-W~\cite{yi2014learning}, VGG~\cite{schroff2015facenet} \&VGG2~\cite{Cao18} & 10,800& 600 & 6 &\checkc& \checkc &\checkc \\
            \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &$\approx$80,000&$\approx$12,000& 4 & \xmark & \checkc &\xmark \\
            \gls{bfw} (ours) & VGG2 & 20,000 & 800 &8 & \checkc & \checkc &\checkc \\\bottomrule
        \end{tabular}
        \label{tab:compared}
    \end{table*}

\section{Background Information}
    \subsection{Bias in \gls{ml}}
        The progress and the commercial value of \gls{ml} is exciting; however, the trust of society in \gls{ml} will take time to build due to inherent biases. The exact definitions and implications of bias vary between sources, as do its sources and types. A common theme is that bias hinders performance ratings in ways that skew to a particular sub-population-- the source varies, whether from humans~\cite{windmann1998subconscious}, data and label types~\cite{tommasi2017deeper}, \gls{ml} models~\cite{amini2019uncovering, kim2019learning}, or evaluation protocols~\cite{stock2018convnets}. For instance, a vehicle-detection model might miss cars if training data was mostly trucks. In practice, many \gls{ml} systems learn on biased data, which could be detrimental for society.

    \subsection{Bias in \gls{fr}}
        Problems of bias in \gls{fr} have been driven by different motivations; therefore, solutions have been posed for different matters. To name a few: data augmentation~\cite{yin2019feature}, one-shot learning~\cite{ding2018one}, demographic parity and fairness with priority on privacy~\cite{huang2018generative}, domain adaptation~\cite{wang2018racial}, differences in face-based attributes across demographics~\cite{wang2018they}, and even data exploration~\cite{muthukumar2019}. Yin~\etal proposed to augment the feature space of underrepresented classes using other classes with a diverse collection of samples to encourage distributions of underrepresented classes to more closely resemble that of the others~\cite{yin2019feature}. Similarly, others formulated the imbalanced class problem as one-shot learning, where a \gls{gan} was trained to generate face features to augment classes with fewer samples~\cite{ding2018one}. \gls{gapf} was proposed to create fair representations of the data in a quantifiable way, allowing for the finding of a de-correlation scheme from the data without access to its statistics~\cite{huang2018generative}. Wang~\etal defined subgroups at a finer-level (\ie Chinese, Japanese, Korean), and determined the familiarity of faces inter-subgroup~\cite{wang2018they}. Genders have also been used to make subgroups (\eg for analysis of gender-based face encodings~\cite{muthukumar2019}). Most recently,~\cite{wang2018racial} proposed adapting domains to bridge the bias gap by knowledge transfer, which was supported by a novel data collection, \gls{rfw}. The release of \gls{rfw} occurred after \gls{bfw} was built - although similar in terms of demographics, \gls{rfw} uses faces from MSCeleb~\cite{guo2016ms} for testing, and assumes CASIA-Face~\cite{yi2014learning} and VGG2~\cite{Cao18} were used to train. In contrast, our \gls{bfw} assumes VGG2 as the test set. 
        % As shown in \cite{wang2018racial}, MSCeleb is highly imbalanced, primarily consisting of images of white individuals. 
        % For this, we expect bias from data, for this is the training set. 
        Furthermore, \gls{bfw} balances gender and race: \gls{rfw} splits subgroups by gender and race, while \gls{bfw} splits by gender, race, or gender and race). 
        % Thus, \gls{rfw} and \gls{bfw} complement one another. % - both add demographic for subjects with faces in renowned, large-scale \gls{fr} datasets.

Most similar to us is~\cite{das2018, demogPairs, lopez2019dataset, srinivas2019face} - each were motivated by insufficient paired data for studying bias in \gls{fr}; then, problems were addressed using labeled data from existing image collections. Specifically, Hupont~\etal curated a set of faces based on racial demographics (\ie \gls{a}, \gls{b}, and \gls{w}) called \gls{dp}~\cite{demogPairs}, while~\cite{srinivas2019face} honed in on adults versus children called \gls{itwcc}. Like the proposed \gls{bfw}, both were built by sampling existing databases, but with the addition of tags for the respective subgroups of interest. Besides, the additional data of \gls{bfw} (\ie added an additional subgroup \gls{i}, along with additional subjects with more faces for all subgroups), we also further split subgroups by gender. Furthermore, we focus on the problem of facial verification and the different levels of sensitivity in cosine similarity scores per subgroup.

\subsection{Human bias in \gls{ml}}
Bias is not unique to \gls{ml} - humans are also susceptible to a perceived bias. In fact, biases exist across race, gender, and even age~\cite{10.1007/978-3-030-13469-3_68, bar2006, meissner2001, nicholls2018}. Wang~\etal showed machines surpass human performance in discriminating between Japanese, Chinese, or Korean faces by nearly 150\%~\cite{wang2018they}, as humans just pass random (\ie 38.89\% accuracy).

We expect human bias to skew to their own genders and races. For this, we measure the human perception with face pairs of different subgroups (Section~\ref{subsec:human-assessment}). The results concur with~\cite{wang2018they}, as we also recorded overall averages below random ($<$50\%). %The details on settings and number of submissions per demographics are discussed in Section~\ref{}.

\glsunset{if}\glsunset{im}\glsunset{af}\glsunset{am}\glsunset{bf}\glsunset{bm}\glsunset{wf}\glsunset{wm}
\section{The BFW Benchmark and Dataset}

We now discuss the \gls{bfw} dataset (\ie collection and preparation), along with the evaluation protocols. We conclude this section by reviewing the settings followed to survey bias in humans.

\subsection{The data}
Problems of bias in \gls{fr} motivated us to build \gls{bfw}. The data evenly represent various subgroups partitioned by demographic. Inspired by \gls{dp}~\cite{demogPairs}, the specification of \gls{bfw} follows in suit, but with additional subgroups (\ie \gls{if} and \gls{im}), an increase in the number of subjects per subgroup, and many more pairs (Table~\ref{tab:ethnic-splits}).

\vspace{1mm}
\noindent\textbf{Compiling subject list.} 
Subjects were sampled from VGG2~\cite{Cao18} - unlike others that depend on multiple sources, \gls{bfw} has fewer potential conflicts in train and test overlap with existing models. We used pre-trained ethnicity~\cite{ambekar2009name} and gender~\cite{levi2015age} classifiers to find candidates for the different subgroups.

\vspace{1mm}
\noindent\textbf{Detecting faces.} Faces were detected using MTCNN~\cite{zhang2016joint}.\footnote{\href{https://github.com/polarisZhao/mtcnn-pytorch}{https://github.com/polarisZhao/mtcnn-pytorch}} Then, assigned into one of two sets. Faces within detected bounding box (BB) regions extended out 130\% in each direction, with zero-padding as the boundary condition made-up one set. The second set were faces aligned and cropped for Sphereface~\cite{liu2017sphereface} (see the next step). Also, coordinates of the BB and the five landmarks from \gls{mtcnn} were stored as part of the static, raw data. For samples with multiple face detections, we used the BB area times the confidence score of the \gls{mtcnn} to determine the face most likely to be the subject of interest, with the others set aside and labeled \textit{miss-detection}.

\begin{figure}[!t] 
	\centering    
 \glsunset{fpr}
  \glsunset{fnr}
	\includegraphics[width=\linewidth]{figures/detcurve-improved.pdf}
		\caption{\textbf{\gls{det} curves.} \emph{Top-left}: per gender. \emph{Top-right}: per ethnicity. \emph{Bottom}: per subgroup (\ie combined). A dashed line shows a difference by about 2$\times$ in \gls{fpr} for the same threshold $\theta_{const}$. \gls{fnr} is the match error count (closer to the bottom is better).}
		\glsreset{det}\glsreset{fpr}
\label{fig:detcurves} 
\end{figure}

\vspace{1mm}
\noindent\textbf{Validating labels.} 
Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Sphereface~\cite{liu2017sphereface}: applying an affine transformation to align faces according to predefined eye locations, each face fed through twice (\ie the original and horizontally flipped), with the final encoding obtained by fusing the two features by average pooling (\ie 512 D). A matrix of cosine similarity scores was then generated for each subject and removed samples (\ie rows) with median scores below threshold $\theta=0.2$ (set manually). Mathematically, the $n^{th}$ sample for the $j^{th}$ subject with $N_j$ faces was removed if the ordinal rank of its score $n = \frac{P\times N}{100}\geq\theta$, where $P=50$. In other words, the median (\ie 50 percentile) of all scores for a faces with respect to all of faces for the respective subject must pass a threshold of $\theta=0.2$; else, the face is dropped. This allowed us to quickly prune \gls{fp} face detections. Following~\cite{robinson2016families, robinson2018visual}, we built a JAVA tool to visually validate the remaining faces. For this, the faces were ordered by decreasing confidence, with confidence set as the average score, and then displayed as image icons on top toggling buttons arranged as a grid in a sliding pane window. Labeling then consisted of going subject-by-subject and flagging faces of \emph{imposters}.

\vspace{1mm}
\noindent\textbf{Sampling faces and creating folds.} We created lists of pairs in five-folds with subjects split evenly per subject and without overlap across folds. Furthermore, a balance in the number of faces per was obtained by sampling twenty-five faces at random from each. Next, we generated a list of all the face pairs per subject, resulting in $\sum_{l=1}^{L}\sum_{k=1}^{K_d} {N_k \choose 2}$ positive pairs, where the number of faces of all $K_l$ subjects $N_k=25$  for each of the $L$ subgroups (Table~\ref{tab:ethnic-splits}). Next, we assigned subjects to a fold. To preserve balance across folds, we sorted subjects by the number of pairs and then started assigning to alternating folds from the one with the most samples. Note, this left no overlap in identity between folds. Later, a negative set from samples within the same subgroup randomly matched until the count met that of the positive. Finally, we doubled the total count with negative pairs from across subgroups but in the same fold.
\begin{figure}[t!] 
	\glsunset{fpr}
	\glsunset{sdm}
	\centering
	\centering
	\includegraphics[width=1\linewidth]{figures/violinplots.pdf}
		\caption{\textbf{\Gls{sdm} across subgroups.} Scores of \emph{imposters} have medians around 0.3 but with variations in upper percentiles; \emph{genuine} pairs vary in mean and spread (\eg \gls{af} has more area of overlap). A threshold varying across different subgroups yields a constant \gls{fpr}.} \label{fig:detection-model} 
		\glsreset{sdm}
\end{figure}

% \vspace{-5pt}
\subsection{Problem formulation}\label{subsec:pf} 
% \gls{lfw}~\cite{LFWTech}, a commonly used benchmark for ~\gls{fr}, reserves specific train-test face-pair lists. 
\Gls{fv} is the special case of the two-class (\ie boolean) classification. Hence, pairs are labeled as the ``same'' or ``different'' \textit{genuine} pairs (\ie \textit{match}) or \textit{imposter} (\ie \textit{mismatch}), respectively. This formulation (\ie \gls{fv}) is highly practical for applications like access control, re-identification, and surveillance. Typically, training a separate model for each unique subject is infeasible. Firstly, the computational costs compound as the number of subjects increase.  Secondly, such a scheme would require model retraining each time a new person is added. Instead, we train models to encode facial images in a feature space that captures the uniqueness of a face, to then determine the outcome based on the output of a scoring (or distance) function. Formally put:
\begin{equation}\label{eg:matcher}
    f_{boolean}(\vec{x}_i, \vec{x}_j) = d(\vec{x}_i, \vec{x}_j) \leq \theta
\end{equation}

where $f_{boolean}$ is the \textit{matcher} of the feature vector $\vec{x}$ for the $i^{th}$ and $j^{th}$ sample~\cite{LFWTech}.

Cosine similarity is used as the \emph{matcher} in Eq~\ref{eg:matcher} the closeness of $i^{th}$ and $j^{th}$ features, \ie
$
s_l= \frac{f_i\cdot f_j}{||f_i||_2||f_j||_2}
$ is the closeness of the $l^{th}$ pair.

\subsection{Human assessment}\label{subsec:human-assessment}
We evaluated human on face pairs focusing on two racial groups: Chinese and Caucasians. To focus on the experiment, we honed-in on two groups, white Americans (W) and Chinese from China (C). The purpose was to minimize variability by only analyzing the subsets of the broader groups of whites and Asians.

Samples were collected by recruiting subjects from multiple sources (\eg social media, email lists, and family/friends)-- a total of 120 participants were sampled at random from all the submissions that were (1) complete and (2) from a W or C participant. Specifically, there were 60 W and 60 C, both with \gls{m} and \gls{f} split evenly. A total of 50 face pairs of non-famous ``look-alikes'' were collected from the internet, with 20 ({\emph WA}) and 20 ({\emph C}) pairs (male and female split evenly). The other 10 pairs are of others (\eg Hispanic/ Latino, Japanese, African). The survey was created, distributed, and recorded via \href{https://paperform.co}{PaperForm}.

\begin{figure}[t!]
	\centering    
	\includegraphics[width=.85\linewidth]{figures/confusion.pdf}
		\caption{\small{\textbf{Confusion matrix.} Percent error (Rank 1) for all faces of \gls{bfw} queried against all others. Notice errors concentrate within a subgroup, consistent with the \gls{sdm} in Fig.~\ref{fig:detection-model} (\ie \gls{af} show worst performance, and is mostly confused with faces of the same demographic). This plot is evidence that while race/ethnicity may be challenging to define, the subgroups are meaningful.}}
		\label{fig:confusion} 
\end{figure}

\section{Results and Analysis}
A single \gls{cnn} was used as a means to control the experiments. For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https://github.com/clcarwin/sphereface\_pytorch}} As reported in~\cite{wang2019racial}, \gls{lfw} has about a 13\%, 14\%, 3\%, and 70\% ratio in Asians, Blacks, Indians, and Whites, respectively; furthermore, CASIA-Web is even more unbalanced (again, as reported in~\cite{wang2019racial}), with about  3\%, 11\%, 2\%, and 85\% for the same subgroups.

% Specifically, we analyze three \glspl{cnn} variants: we encode faces using VGG-16~\cite{simonyan2014very},  50-layer residual network (\ie ResNet-50)~\cite{he2016deep}, and \gls{senet} (\ie SENet-50)~\cite{hu2018squeeze}, with results of only the latter (\ie the best performing \gls{senet}) model used throughout the main paper. Results for the other two, alongside \gls{senet} for comparison, are in the Supplemental Material.

\glsunset{am}\glsunset{af}\glsunset{bm}\glsunset{bf}\glsunset{im}\glsunset{if}\glsunset{wm}\glsunset{wf}

\subsection{Score analysis}
Fig.~\ref{fig:detection-model} shows score distributions for faces of the same (\ie \emph{Genuine}) and different (\ie \emph{Imposter}) identity, with a subgroup per \gls{sdm} plot. Notice that score distributions for imposters tend to peak about zero for all subgroups, and with minimal deviation comparing modes of the different plots. On the other hand, the score distribution of the \emph{genuine} pairs varies across subgroups in location (\ie score value) and spread (\ie overall shape). Asian Fem
Fig.~\ref{fig:confusion} shows the confusion matrix of the subgroups. A vast majority of errors occur in the intra-subgroup. It is interesting to note that while the definition of  each group  based on ethnicity and race may not be crisply defined, the confusion matrix indicates that in practice the \gls{cnn} finds that the groups are effectively separate. The categories are, therefore, meaningful in the context of \gls{fr}.

\subsection{\gls{det} analysis}
\glsunset{m}
\glsunset{f}

\begin{figure*}[h!]
    \centering
    \begin{subfigure}[t]{.3\linewidth}
    \includegraphics[width=.75\linewidth]{figures/curve_vgg16_subgroups-crop.pdf}
    \caption{VGG16~\cite{simonyan2014very}}
 \end{subfigure}
    \begin{subfigure}[t]{.27\linewidth}
    \includegraphics[width=.8\linewidth]{figures/curve_resnet50_subgroups-crop.pdf}
    \caption{ResNet50~\cite{he2016deep}}
   \end{subfigure}
    \begin{subfigure}[t]{.27\linewidth}
    \includegraphics[width=.8\linewidth]{figures/curve_senet50_subgroups-crop.pdf}
    \caption{SENet~\cite{hu2018squeeze}}
    \end{subfigure}
    \caption{\textbf{\gls{det} curves for different CNNs}. \gls{fnr} (\%) (vertical) vs \gls{fpr}  (horizontal, log-scale) for VGG2~\cite{Cao18} models with different backbones (VGG16, Resnet50, SEnet50). Lower is better. For each plot, \gls{wm} is the top-performer, while \gls{af} is the worst. The the ordering of the curves is roughly the same for each backbone.}\label{fig:sdm-appendix-a}
\end{figure*}

\gls{det} curves averaged across 5-folds show per-subgroup trade-offs (Fig.~\ref{fig:detcurves}). Note that \gls{m} performs better than \gls{f}, precisely as one would expect from the tails of score-distributions for \emph{genuine} pairs (Fig.~\ref{fig:detection-model}). \Gls{af} and \gls{if} perform the worst.

\begin{table}[b!]
\glsunset{tar}
\glsunset{far}
\caption{\small{\textbf{\gls{tar} at intervals of \gls{far}}. For each subgroup, and overall average \gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom). Higher is better. The proposed shows improvement in all cases.}}\label{tab:ethnicy-far} 
%  \vspace{-2mm}
\begin{center}
\scriptsize
\begin{tabular}{l c c c c c}
     \gls{far} & 0.3 & 0.1 & 0.01 & 0.001 & 0.0001\\\midrule
    \multirow{2}{.1mm}{\textbf{\gls{af}}} &0.990 & 0.867 & 0.516 & 0.470 & 0.465\\[-4pt]
        &1.000 & 0.882 & 0.524 & 0.478 & 0.474\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{am}}} &0.994 & 0.883 & 0.529 & 0.482 & 0.477\\[-4pt]
        &1.000 & 0.890 & 0.533 & 0.486 & 0.482\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{bf}}} &0.991 & 0.870 & 0.524 & 0.479 & 0.473\\[-4pt]
        &1.000 & 0.879 & 0.530 & 0.484 & 0.480\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{bm}}} &0.992 & 0.881 & 0.526 & 0.480 & 0.474\\[-4pt]
        &1.000 & 0.891 & 0.532 & 0.485 & 0.480\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{if}}} &0.996 & 0.881 & 0.532 & 0.486 & 0.481\\[-4pt]
        &1.000 & 0.884 & 0.534 & 0.488 & 0.484\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{im}}} &0.997 & 0.895 & 0.533 & 0.485 & 0.479\\[-4pt]
        &1.000 & 0.898 & 0.535 & 0.486 & 0.481\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{wf}}} &0.988 & 0.878 & 0.517 & 0.469 & 0.464\\[-4pt]
        &1.000 & 0.894 & 0.526 & 0.478 & 0.474\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{wm}}} &0.989 & 0.896 & 0.527 & 0.476 & 0.470\\[-4pt]
        &1.000 & 0.910 & 0.535 & 0.483 & 0.478\\[-1pt]
    \midrule
    \multirow{2}{3mm}{\textbf{Avg.}} &0.992 & 0.881 & 0.526 & 0.478 & 0.474\\[-4pt]
        &1.000 & 0.891 & 0.531 & 0.483 & 0.479\\[-10pt]
\end{tabular}
\end{center}
\glsreset{tar}
\glsreset{far}
%  \vspace{-3mm}

\end{table}

The gender-based \gls{det} curve shows a difference in performances between \gls{m} and \gls{f} with a fixed threshold (dashed line). Similar effects exist for the other curves as well (lines omitted to declutter). For many \gls{fr} applications, systems are operated at the highest \gls{fpr} allowed, so the line of constant threshold indicates that a single threshold produces different operating points (\ie \gls{fpr}), which is undesirable.  The difference in \gls{fpr} is approximately a factor of two-- quite large indeed. If this is the case in an industrial system, one would expect a difference in about double the false positives to be reported based on subgroup alone. The potential ramifications of such a bias should not be overlooked, which it has not as of lately-- gaining the interest in even main-stream media ~\cite{england2019,snow2018}.

% \subsubsection{Other \gls{cnn} models.}~\label{app:sec:other:models}
Variations in optimal threshold exist across models (Fig.~\ref{fig:sdm-appendix-a}). Like in Fig.~\ref{fig:detcurves}, the \gls{det} curves for three \gls{cnn}-based models, each trained on VGG2 with softmax but with different backbones.\footnote{\href{https://github.com/rcmalli/keras-vggface}{https://github.com/rcmalli/keras-vggface}} Notice similar trends across subgroups and models, which is consistent with  Sphereface as well (Fig.~\ref{fig:detcurves}). For example, the plots generated with Spherface and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst). Thus, the dditional \gls{cnn}-based models demonstrate the same phenomena: proportional to the overall model performance, exact in which the ordering subgroups insensitivity in scores space.

\subsection{Verification threshold} \label{subsec:analysis:verification}
We seek to reduce the bias between subgroups. Such that an operating point (\ie \gls{fpr}) is constant across subgroups. To accomplish that, we used a per subgroup threshold. In \gls{fv}, we consider one image as the query and all others as the test. For this, the ethnicity of the query image is assumed. We can then examine the \gls{det} curves and pick the best threshold per subgroup for a certain \gls{fpr}.

We evaluated \gls{tar} for specific \gls{far} values. As described in Section~\ref{subsec:pf}, the verification experiments were 5-fold, with no overlap in subject ID between folds. Results reported are averaged across folds in all cases and are shown in Table~\ref{tab:ethnicy-far}. For each subgroup, the \gls{tar} of using a global threshold is reported (upper row), as well as using the optimal per subgroup threshold (lower row).

Even for lower \gls{far}, there are notable improvements, often of the order of 1\%, which can be challenging to achieve when \gls{far} is near $\geq$90\%. More importantly, each subgroup has the desired \gls{fpr}, so that substantial differences in \gls{fpr} will remain unfounded. We experimented with ethnicity estimators on both the query and test image, which yielded similar results to those reported here.

\begin{figure}[t!] 
	\centering    
	\includegraphics[width=.85\linewidth] {figures/human_eval.pdf}
		\caption{\small{\textbf{Qualitative results of human assessment.} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}. Accuracy scores shown as bar plots. Humans are most successful at recognizing their own subgroup, with a few exceptions (\eg bottom).}}
		% (\eg bottom row). } Result summary in Table~\ref{tab:human-eval}.
		\label{fig:human-eval} 
% 		\vspace{-1mm}
\end{figure} 
% \vspace{-4mm}
% \vspace{-5pt}

\subsection{Human evaluation}
 Quantitative and qualitative results are in Table~\ref{tab:humsn-eval-results} and Fig.~\ref{fig:human-eval}, respectively. Subjects of a subgroup likely have mostly been exposed to others of the same. Therefore, it is expected they would be best at labeling their own; similar for the same ethnicity, but another gender. Our findings concur. Each subgroup is best at labeling their type, and then second best at labeling the same ethnicity but opposite sex. Interestingly, each group of images is best tagged by the corresponding subgroup, with the second-best having the same ethnicity and opposite gender. On average, subgroups are comparable at labeling images. Analogous to the \gls{fr} system, performance ratings differ when analyzing within and across subgroups. In other words, performance on \gls{bfw} improved with subgroup-specific thresholds. Similarly, humans tend to better recognize individuals by facial cues of the same or similar demographics. Put differently, as the recognition performances drop with a global threshold optimized for one subgroup and deployed on another, human performance tends to drop when across subgroups (\ie most familiar with the same subgroup; performance seems to drop when faced with another, less familiar subgroup).

%  for it better adapts wh specific to subgroup) shift in sensitivity across subgroups when deep encodings are compared, the underlying mechanism

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t!]
\begin{center}
    \caption{\small{\textbf{Quantitative of human assessment.} Different subgroups listed per row and column. Note that each subject is best within their subgroup, and second-best within the same subgroup but different gender. CF shows the least variation, but with the lowest accuracy. CM shows the best accuracy, but second to \gls{wm} in deviation from the mean. Thus, scores for males vary more than females.}}
    \label{tab:humsn-eval-results} 
    %  \vspace{-2mm}
\footnotesize
\scalebox{0.8}{
\begin{tabular}{l c c c c c}
        &  CF  &      CM     &   WF     &   WM     &  Avg\\\midrule
        CF &  \textbf{\textcolor{blue}{52.9}}&   \textbf{\textcolor{red}{48.0}}&43.8&44.7 &47.4$\pm$4.1 \\
        CM &  \textbf{\textcolor{red}{45.6}} & \textbf{\textcolor{blue}{0.504}}  & 44.4 &36.2 &44.1$\pm$5.9 \\
        WF & 44.7 &43.8 & \textbf{\textcolor{blue}{0.573}}& \textbf{\textcolor{red}{48.0}} & 48.5$\pm$6.2 \\
        WM & 30.1&\textbf{\textcolor{red}{47.4}} &  45.3 & \textbf{\textcolor{blue}{56.1}} & 44.7$\pm$10.8 \\\midrule
        Avg &  43.3& 47.4&47.7 &46.3 &46.2$\pm$2.0\\
 \end{tabular}}
 \end{center}
%  \vspace{-7mm}
\end{table}

\glsresetall
\section{Conclusion}
We introduce a new data set \gls{bfw} with eight subgroups balanced across gender and ethnicity. With this, and upon highlighting the challenges and shortcomings of grouping subjects as a single subset, we provide evidence that forming subgroups is meaningful, as the FR algorithm rarely makes mistakes across subgroups. We used an \textit{off-the-shelf} Sphereface, hypothesizing this SOTA CNN suffers from bias because of the imbalanced train-set. Once established that the results do suffer from problems of bias, we observed that the same threshold across ethnic and gender subgroups leads to differences in the FPR up to a factor of two, which is seemingly the cause of the frenzy about bias in FR in main-stream media. Furthermore, we ameliorate these differences with a per-subgroup threshold, leveling out FPR, and achieving a higher TPR. We hypothesized that most humans grew amongst more than their own demographic and, therefore, effectively learn from imbalanced datasets-- a human evaluation validated that humans are biased, as most recognized their personal demographic best. The focused research findings presented here, along with the public database included, are extendable in vast ways. Thus, we see this as the slither to a much larger problem of bias in ML.

{\small
\bibliographystyle{ieee_fullname}
\bibliography{references}
}

% \putbib
% \end{bibunit}
% \begin{bibunit}
\input{supplemental.tex}
% \end{bibunit}

\end{document}

### 2020-03-11 manuscript.tex
arXiv
added cvpr template and started filling in

@@
- \documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
+%  \documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4
+\documentclass[10pt,twocolumn,letterpaper]{article}
+
+\usepackage{cvpr}
 
@@
 
-\usepackage{FG2019}
+% \usepackage{FG2019}
 \usepackage{multirow}
+\usepackage{graphicx}
+
 \newcommand{\minitab}[2][c]{\begin{tabular}{#1}#2\end{tabular}}
@@
 \usepackage{bigstrut}
+\usepackage{cvpr}
+\usepackage{times}
+\usepackage{epsfig}
+\usepackage{graphicx}
+\usepackage{amsmath}
+\usepackage{amssymb}
 % symbols for 1/2, etc.
@@
 \usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-% \usepackage{hyperref}       % hyperlinks
-\usepackage[colorlinks = true,
-            linkcolor = blue,
-            urlcolor  = blue,
-            citecolor = blue,
-            anchorcolor = blue]{hyperref}
+% % \usepackage{hyperref}       % hyperlinks
+% \usepackage[colorlinks = true,
+%             linkcolor = blue,
+%             urlcolor  = blue,
+%             citecolor = blue,
+%             anchorcolor = blue]{hyperref}
+\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
 
@@
 
-\graphicspath{{images/}} %Setting the graphicspath
+% \graphicspath{{figures/}} %Setting the graphicspath
 \usepackage{bibunits}
@@
 \newcommand{\com}[1]{\textcolor{red}{#1}}
-\newcommand{\ie}{\textit{i}.\textit{e}., }
-\newcommand{\eg}{\textit{e}.\textit{g}., }
-\newcommand*{\etc}{etc.\@\xspace}
+% \newcommand{\ie}{\textit{i}.\textit{e}., }
+% \newcommand{\eg}{\textit{e}.\textit{g}., }
+% \newcommand*{\etc}{etc.\@\xspace}
 
@@
 
-\FGfinalcopy % *** Uncomment this line for the final submission
+% \FGfinalcopy % *** Uncomment this line for the final submission
 \let\vec\mathbf
@@
 
-\title{\LARGE \bf
-Face Recognition: Too Bias, or Not Too Bias?
-}
-
-% \title{\LARGE \bf\vspace{-10mm}
+% \title{\LARGE \bf
 % Face Recognition: Too Bias, or Not Too Bias?
@@
 
-%use this in case of several affiliations
-\author{\parbox{16cm}{\centering
-    {\large Joseph P Robinson$^1$, Gennady Livitz$^2$, Yann Henon$^2$, Can Qin$^1$,\\ Yun Fu$^1$, and Samson Timoner$^2$}\\
-    {\normalsize
-    \hspace{-.4in}$^{1}$Northeastern University\hspace{.7in} $^{2}$ISM Connect}}
-    \thanks{}% <-this % stops a space
+% % \title{\LARGE \bf\vspace{-10mm}
+% % Face Recognition: Too Bias, or Not Too Bias?
+% % }
+
+% %use this in case of several affiliations
+% \author{\parbox{16cm}{\centering
+%     {\large Joseph P Robinson$^1$, Gennady Livitz$^2$, Yann Henon$^2$, Can Qin$^1$,\\ Yun Fu$^1$, and Samson Timoner$^2$}\\
+%     {\normalsize
+%     \hspace{-.4in}$^{1}$Northeastern University\hspace{.7in} $^{2}$ISM Connect}}
+%     \thanks{}% <-this % stops a space
+% }
+
+% \cvprfinalcopy % *** Uncomment this line for the final submission
+
+\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
+\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
+
+% Pages are numbered in submission mode, and unnumbered in camera-ready
+\ifcvprfinal\pagestyle{empty}\fi
+\begin{document}
+
+%%%%%%%%% TITLE
+\title{Face Recognition: Too Bias, or Not Too Bias?}
+
+\author{First Author\\
+Institution1\\
+Institution1 address\\
+{\tt\small firstauthor@i1.org}
+% For a paper whose authors are all at the same institution,
+% omit the following lines up until the closing ``}''.
+% Additional authors and addresses can be added with ``\and'',
+% just like the second author.
+% To save space, use either the email address or home page, not both
+\and
+Second Author\\
+Institution2\\
+First line of institution2 address\\
+{\tt\small secondauthor@i2.org}
 }
 
-\begin{document}
+\maketitle
+
+% \begin{document}
 %\begin{bibunit}
 
-\def\angle{0}
-\def\radius{3}
-\def\cyclelist{{"orange","blue","red","green","magenta","cyan"}}
-\newcount\cyclecount \cyclecount=-1
-\newcount\ind \ind=-1
+% \def\angle{0}
+% \def\radius{3}
+% \def\cyclelist{{"orange","blue","red","green","magenta","cyan"}}
+% \newcount\cyclecount \cyclecount=-1
+% \newcount\ind \ind=-1
 
-\ifFGfinal
-\thispagestyle{empty}
-\pagestyle{empty}
-\else
-\author{Anonymous FG 2019 submission\\ Paper ID \FGPaperID \\}
-\pagestyle{plain}
-\fi
-\maketitle
+% \ifFGfinal
+% \thispagestyle{empty}
+% \pagestyle{empty}
+% \else
+% \author{Anonymous FG 2019 submission\\ Paper ID \FGPaperID \\}
+% \pagestyle{plain}
+% \fi
+% \maketitle
 
@@
 \end{abstract}
-\
+
 \input{sections/introduction.tex}
@@
 
-% \vspace{-3mm}
-\bibliographystyle{ieee}
-\balance
+% % \vspace{-3mm}
+% \bibliographystyle{ieee}
+% \balance
+% \bibliography{references}
+
+{\small
+\bibliographystyle{ieee_fullname}
 \bibliography{references}
+}
 
@@
 % \end{bibunit}
-\begin{bibunit}
-\input{supplemental.tex}
-\end{bibunit}
+% \begin{bibunit}
+% \input{supplemental.tex}
+% \end{bibunit}
 

### 2020-03-11 supplemental.tex
added cvpr template and started filling in

@@
 \onecolumn
+
 % \begin{minipage*}[\linewidth]{width of the minipage}
-\renewcommand{\thesection}{\alph{section}}
+% \renewcommand{\thesection}{\alph{section}}
 \glsresetall
-\setcounter{section}{0}
-
+% \setcounter{section}{0}
+\begin{comment}
 \section{Other \gls{cnn} models.}~\label{app:sec:other:models}
@@
 \begin{figure}[h!]
-\vspace{-1mm}
     \centering
@@
 The the ordering of the curves is roughly the same for each backbone.}\label{fig:sdm-appendix-a}
-    \vspace{-1mm}
+\end{figure}
+\end{comment}
+
+\begin{table}
+\begin{center}
+\glsunsetall
+\caption{Intended \gls{fpr} per subgroup.
+Each tuple represents percentage the respective rate was achieved (\textcolor{green}{+\%} or \textcolor{red}{-\%}) paired with FNR.
+Each subgroup shows the use of $t_g$ (top) and individual $t_o$ (bottom).}
+\begin{tabular}{l c c c c c}
+     \gls{fpr} & 0.3 & 0.1 & 0.01 & 0.001 & 0.0001\\\midrule
+    \multirow{2}{.1mm}{\textbf{\gls{af}}} & (\textcolor{red}{-4.5}, )&(\textcolor{red}{-13.4}, )  &(\textcolor{red}{-25.8}, ) &(\textcolor{red}{-25.8}, )  &(\textcolor{red}{-25.5}, )  \\[-4pt]
+        & (0.0, ) &(0.0, )  & (0.0, ) &  (\textcolor{green}{0.1}, )&(\textcolor{red}{-5.6}, )  \\[-1pt]
+
+    \multirow{2}{3mm}{\textbf{\gls{am}}}&(\textcolor{green}{2.3}, )&(\textcolor{green}{4.1}, )&(\textcolor{green}{11.2}, )&(\textcolor{green}{14.5}, )&(\textcolor{green}{52.1}, ) \\[-4pt]
+        & (0.0, ) &(0.0, )  & (\textcolor{green}{0.2}, ) & (\textcolor{green}{0.2}, ) & (\textcolor{red}{5.4}, )\\[-1pt]
+
+    \multirow{2}{3mm}{\textbf{\gls{bf}}}& (\textcolor{red}{-1.6}, ) &(\textcolor{red}{-8.2}, ) &(\textcolor{red}{-23.4}, ) &(\textcolor{red}{-27.7}, ) &(\textcolor{red}{-41.8}, ) \\[-4pt]
+        & (0.0, ) &(\textcolor{red}{-0.1}, )  & (0.0, ) & (0.0, ) & (\textcolor{red}{5.7}, )\\[-1pt]
+
+    \multirow{2}{3mm}{\textbf{\gls{bm}}}&(\textcolor{green}{2.3}, )&(\textcolor{green}{6.5}, )&(\textcolor{green}{17.6}, )&(\textcolor{green}{29.8}, )&(\textcolor{green}{95.8}, ) \\[-4pt]
+        & (0.0, ) &(\textcolor{red}{-0.1}, )  & (0.0, ) & (\textcolor{green}{0.1}, ) & (\textcolor{red}{5.5}, )\\[-1pt]
+
+    \multirow{2}{3mm}{\textbf{\gls{if}}} &(\textcolor{red}{-9.0}, ) &(\textcolor{red}{-26.1}, ) &(\textcolor{red}{-58.6}, ) &(\textcolor{red}{-46.1}, ) &(\textcolor{red}{-60.9}, )\\[-4pt]
+        & (0.0, ) &(0.0, )  & (0.0, ) & (\textcolor{green}{0.2}, ) & (\textcolor{red}{5.4}, )\\[-1pt]
+
+    \multirow{2}{3mm}{\textbf{\gls{im}}} &(\textcolor{green}{2.4}, )&(\textcolor{green}{4.3}, )&(\textcolor{red}{-3.4}, )&(\textcolor{red}{-16.4}, )&(\textcolor{red}{-76.2}, )\\[-4pt]
+        & (0.0, ) &(0.0, )  & (\textcolor{green}{0.1}, ) & (\textcolor{green}{0.1}, )  &(\textcolor{red}{5.5}, ) \\[-1pt]
+
+    \multirow{2}{3mm}{\textbf{\gls{wf}}} &(\textcolor{red}{-1.1},) &(\textcolor{green}{5.8}, ) &(\textcolor{green}{50.6}, ) &(\textcolor{green}{52.6}, ) &(\textcolor{green}{158.0}, )\\[-4pt]
+       & (0.0, ) &(\textcolor{red}{-0.1}, )  & (\textcolor{green}{0.3}, ) & (\textcolor{green}{0.2}, )  &(\textcolor{red}{5.5}, )  \\[-1pt]
+
+    \multirow{2}{3mm}{\textbf{\gls{wm}}} & (\textcolor{green}{10.5}, ) & (\textcolor{green}{38.6}, )& (\textcolor{green}{120.0}, )& (\textcolor{green}{83.4}, )& (\textcolor{green}{200.0}, ) \\[-4pt]
+
+        & (0.0, ) &(\textcolor{green}{0.2}, ) &(\textcolor{green}{0.2}, ) & (\textcolor{green}{0.2}, )&(\textcolor{red}{-5.5}, )\\[-1pt]
+    % \midrule
+    % \multirow{2}{3mm}{\textbf{Avg.}} \\[-4pt]
+    %   & (0.0, ) &(0.0, )  & (\textcolor{green}{-}, ) &  \\[-10pt]
+\end{tabular}
+\end{center}
+\end{table}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/SDM-crop.pdf}
+    \caption{Caption}
+    \label{fig:my_label}
 \end{figure}
@@
 \begin{figure}[h!]
-\vspace{-2mm}
     \centering
@@
     \label{fig:montage:app}
-    \vspace{-1mm}
 \end{figure}
-{
+\begin{comment}
 \scriptsize
@@
 \end{thebibliography}
-}
+
+\end{comment}
 

### 2020-04-13 cvpr_workshop_backup.tex
v2

--- 2020-03-11 cvprworkshop.tex
@@
 \usepackage{comment}
-\usepackage{tabularx, booktabs, ragged2e}
+\usepackage{booktabs, ragged2e}
 % \usepackage{arydshln}
@@
 
-\newacronym{cs}{CS}{Cosine Similiarity}
+\newacronym{cs}{CS}{Cosine Similarity}
 
@@
 % \let\vec\mathbf
+% \usepackage[T1]{fontenc}
+% \usepackage[latin9]{inputenc}
+
+\def\colorModel{hsb} %You can use rgb or hsb
+\usepackage{babel}
+\usepackage[table]{xcolor}
+\usepackage{collcell}
+\def\colorModel{hsb} %You can use rgb or hsb
+
+\usepackage{hhline}
+\newcommand\ColCell[1]{
+  \pgfmathparse{#1<50?1:0}  %Threshold for changing the font color into the cells
+    \ifnum\pgfmathresult=0\relax\color{white}\fi
+  \pgfmathsetmacro\compA{0}      %Component R or H
+  \pgfmathsetmacro\compB{#1/100} %Component G or S
+  \pgfmathsetmacro\compC{1}      %Component B or B
+  \edef\x{\noexpand\centering\noexpand\cellcolor[\colorModel]{\compA,\compB,\compC}}\x #1
+  } 
+\newcolumntype{E}{>{\collectcell\ColCell}m{0.4cm}<{\endcollectcell}}  %Cell width
+\newcommand*\rot{\rotatebox{90}}
 
@@
 
-\def\cvprPaperID{0002} % *** Enter the CVPR Paper ID here
+\def\cvprPaperID{0001} % *** Enter the CVPR Paper ID here
 \def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
@@
 Furthermore, we do a human evaluation to measure the bias in humans, which supports the hypothesis that such a bias exists in human perception.
-For the \gls{bfw} database, source code, and more, visit \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}
+For the \gls{bfw} database, source code, and more, visit \href{http://cvpr2020.thecvf.com/submission/main-conference/author-guidelines#policies}{github}.
+% \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}
 \end{abstract}
@@
         \centering
-        \includegraphics[width=.9\linewidth]{figures/montage.pdf}
-        \caption{\textbf{\gls{bfw}.} The average face of its different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender.
+        \includegraphics[width=.8\linewidth]{figures/montage.pdf}
+        \caption{\small{\textbf{\gls{bfw}.} Average face of the different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender.
 The others represent the ethnicity and gender of the race and gender, respectively.
-Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}
+Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}}
         \label{fig:avg-faces}
@@
 
-    The contributions in this paper are 3-fold: (1) Built a balanced dataset as a proxy to measure verification performance per subgroup; (2) Analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup balances performance; (3) Showed bias causes skewness in ratings across demographics-- the typical use of a global threshold unfavorable.
-We mitigate the problem via adaptive thresholds; (4) Conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
+    The contributions of this work are 3-fold.
+(1) We built a balanced dataset as a proxy to measure verification performance per subgroup for studies of bias in \gls{fr}.
+(2) We revealed an unwanted bias in scores of face pairs - a bias that causes ratings to skew across demographics.
+For this, we showed that an adaptive threshold per subgroup balances performance (\ie the typical use of a global threshold unfavorable, which we address via optimal thresholds).
+(3) We surveyed humans to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}.}%19-09-08}.}
     % \begin{enumerate}
@@
     % \end{enumerate}
-
-    \begin{table*}[t!]
-
-        \centering
-        \caption{\textbf{\gls{bfw} compared to related datasets.} Our \gls{bfw} is balanced across ID, gender, and ethnicity (Table~\ref{tab:ethnic-splits}).
-Compared with \gls{dp}, \gls{bfw} provides more samples per subject and subgroups per set, while only using a single resource, VGG2.
-\gls{rfw}, on the other hand, supports a different task (\ie domain adaptation), and focuses on race-distribution, but not the distribution of identities.}
-        \scriptsize
-        \begin{tabular}{rccccccc}%x{10mm}x{3mm}x{10mm}x{8mm}}%\toprule
-
-            \multicolumn{2}{c}{Database} & \multicolumn{3}{c}{Number of}& \multicolumn{3}{c}{Balanced Labels}\\
-            \cmidrule(lr){1-2}	\cmidrule(lr){3-5} \cmidrule(lr){6-8}
-            Name & Source Data & Faces &  IDs & Subgroups & ID & Ethnicity & Gender\\\midrule
-            \gls{dp}~\cite{demogPairs}     & CASIA-W~\cite{yi2014learning}, VGG~\cite{schroff2015facenet} \&VGG2~\cite{Cao18} & 10,800& 600 & 6 &\checkc& \checkc &\checkc \\
-            \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &$\approx$80,000&$\approx$12,000& 4 & \xmark & \checkc &\xmark \\
-            \gls{bfw} (ours) & VGG2 & 20,000 & 800 &8 & \checkc & \checkc &\checkc \\\bottomrule
-        \end{tabular}
-        \label{tab:compared}
-    \end{table*}
 
@@
         % For this, we expect bias from data, for this is the training set. 
-        Furthermore, \gls{bfw} balances gender and race: \gls{rfw} splits subgroups by gender and race, while \gls{bfw} splits by gender, race, or gender and race). 
+        Furthermore, \gls{bfw} balances gender and race: \gls{rfw} splits subgroups by gender and race, while \gls{bfw} has gender, race, or both). 
         % Thus, \gls{rfw} and \gls{bfw} complement one another. % - both add demographic for subjects with faces in renowned, large-scale \gls{fr} datasets.
@@
 Furthermore, we focus on the problem of facial verification and the different levels of sensitivity in cosine similarity scores per subgroup.
+
+    \begin{table*}[t!]
+
+        \centering
+        \caption{\small{\textbf{\gls{bfw} compared to related datasets.} Our \gls{bfw} is balanced across ID, gender, and ethnicity (Table~\ref{tab:ethnic-splits}).
+Compared with \gls{dp}, \gls{bfw} provides more samples per subject and subgroups per set, while only using a single resource, VGG2.
+\gls{rfw}, on the other hand, supports a different task (\ie domain adaptation), and focuses on race-distribution, but not the distribution of identities.}}
+        \scriptsize
+        \begin{tabular}{rccccccc}%x{10mm}x{3mm}x{10mm}x{8mm}}%\toprule
+
+            \multicolumn{2}{c}{Database} & \multicolumn{3}{c}{Number of}& \multicolumn{3}{c}{Balanced Labels}\\
+            \cmidrule(lr){1-2}	\cmidrule(lr){3-5} \cmidrule(lr){6-8}
+            Name & Source Data & Faces &  IDs & Subgroups & ID & Ethnicity & Gender\\\midrule
+            \gls{dp}~\cite{demogPairs}     & CASIA-W~\cite{yi2014learning}, VGG~\cite{schroff2015facenet} \&VGG2~\cite{Cao18} & 10,800& 600 & 6 &\checkc& \checkc &\checkc \\
+            \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &$\approx$80,000&$\approx$12,000& 4 & \xmark & \checkc &\xmark \\
+            \gls{bfw} (ours) & VGG2 & 20,000 & 800 &8 & \checkc & \checkc &\checkc \\\bottomrule
+        \end{tabular}
+        \label{tab:compared}
+    \end{table*}
 
@@
 \noindent\textbf{Compiling subject list.} 
-Subjects were sampled from VGG2~\cite{Cao18} - unlike others that depend on multiple sources, \gls{bfw} has fewer potential conflicts in train and test overlap with existing models.
+Subjects were sampled from VGG2~\cite{Cao18} - unlike others built from multiple sources, \gls{bfw} has fewer potential conflicts in train and test overlap with existing models.
 We used pre-trained ethnicity~\cite{ambekar2009name} and gender~\cite{levi2015age} classifiers to find candidates for the different subgroups.
@@
 For samples with multiple face detections, we used the BB area times the confidence score of the \gls{mtcnn} to determine the face most likely to be the subject of interest, with the others set aside and labeled \textit{miss-detection}.
-
-\begin{figure}[!t] 
-	\centering    
- \glsunset{fpr}
-  \glsunset{fnr}
-	\includegraphics[width=\linewidth]{figures/detcurve-improved.pdf}
-		\caption{\textbf{\gls{det} curves.} \emph{Top-left}: per gender.
-\emph{Top-right}: per ethnicity.
-\emph{Bottom}: per subgroup (\ie combined).
-A dashed line shows a difference by about 2$\times$ in \gls{fpr} for the same threshold $\theta_{const}$.
-\gls{fnr} is the match error count (closer to the bottom is better).}
-		\glsreset{det}\glsreset{fpr}
-\label{fig:detcurves} 
-\end{figure}
 
@@
 
+\begin{figure}[!t] 
+	\centering    
+ \glsunset{fpr}
+  \glsunset{fnr}
+	\includegraphics[width=.9\linewidth]{figures/detcurve-improved.pdf}
+		\caption{\small{\textbf{\gls{det} curves.} \emph{Top-left}: per gender.
+\emph{Top-right}: per ethnicity.
+\emph{Bottom}: per subgroup (\ie combined).
+Dashed line shows about 2$\times$ difference in \gls{fpr} for the same threshold $\theta_{const}$.
+\gls{fnr} is the match error count (closer to the bottom is better).}}
+		\glsreset{det}\glsreset{fpr}
+\label{fig:detcurves} 
+\end{figure} 
 \vspace{1mm}
@@
 Later, a negative set from samples within the same subgroup randomly matched until the count met that of the positive.
-Finally, we doubled the total count with negative pairs from across subgroups but in the same fold.
-\begin{figure}[t!] 
-	\glsunset{fpr}
-	\glsunset{sdm}
-	\centering
-	\centering
-	\includegraphics[width=1\linewidth]{figures/violinplots.pdf}
-		\caption{\textbf{\Gls{sdm} across subgroups.} Scores of \emph{imposters} have medians around 0.3 but with variations in upper percentiles; \emph{genuine} pairs vary in mean and spread (\eg \gls{af} has more area of overlap).
-A threshold varying across different subgroups yields a constant \gls{fpr}.} \label{fig:detection-model}
-		\glsreset{sdm}
-\end{figure}
+Finally, we doubled the number with negative pairs from across subgroups but in the same fold.
 
@@
 \begin{equation}\label{eg:matcher}
-    f_{boolean}(\vec{x}_i, \vec{x}_j) = d(\vec{x}_i, \vec{x}_j) \leq \theta
+    f_{boolean}(\vec{x}_i, \vec{x}_j) = d(\vec{x}_i, \vec{x}_j) \leq \theta,
 \end{equation}
@@
 $ is the closeness of the $l^{th}$ pair.
+
+\begin{figure}[t!] 
+	\glsunset{fpr}
+	\glsunset{sdm}
+	\centering
+	\centering
+	\includegraphics[width=1\linewidth]{figures/violinplots.pdf}
+		\caption{\small{\textbf{\Gls{sdm} across subgroups.} Scores of \emph{imposters} have medians around 0.3 but with variations in upper percentiles; \emph{genuine} pairs vary in mean and spread (\eg \gls{af} has more area of overlap).
+A threshold varying across different subgroups yields a constant \gls{fpr}.}} \label{fig:detection-model}
+		\glsreset{sdm}
+\end{figure}
 
@@
 	\centering    
-	\includegraphics[width=.85\linewidth]{figures/confusion.pdf}
-		\caption{\small{\textbf{Confusion matrix.} Percent error (Rank 1) for all faces of \gls{bfw} queried against all others.
-Notice errors concentrate within a subgroup, consistent with the \gls{sdm} in Fig.~\ref{fig:detection-model} (\ie \gls{af} show worst performance, and is mostly confused with faces of the same demographic).
+	\includegraphics[width=.75\linewidth]{figures/confusion.pdf}
+		\caption{\small{\textbf{Confusion matrix.} Percent error (Rank 1, \%) for all faces of \gls{bfw} versus all others.
+Error concentrates within a subgroup - consistent with the \gls{sdm} (Fig.~\ref{fig:detection-model}), as \gls{af} performs the worst, confusing mostly within subgroup.
 This plot is evidence that while race/ethnicity may be challenging to define, the subgroups are meaningful.}}
@@
 
+\begin{figure}[t!]
+
+    % \begin{subfigure}[t]{\linewidth}
+
+       \centering
+%     \includegraphics[width=.85\linewidth]{figures/global_threshold_fpr_percent_diff.pdf}
+%     \caption{Global threshold ($t_g$)}
+%  \end{subfigure}
+%     \begin{subfigure}[t]{\linewidth}
+%       \centering
+    \includegraphics[width=.95\linewidth]{figures/adaptive_fpr_percent_diff_complete.pdf}
+    % \caption{Optimal threshold ($t_o$)}
+%   \end{subfigure}
+    \caption{\small{\textbf{Percent difference from intended \gls{fpr}.} \emph{Top:} $t_g$ yields \gls{fpr} the span as large as 2$\times$ (\ie 200\%) that intended (\ie \gls{wm} for 1e-4).
+Furthermore, \gls{f} subgroups tend to perform worse than intended for all cases (while \gls{m} tend to overshoot intended performance, with exception of \gls{im} in for \gls{fpr}=1e-4).
+\emph{Bottom:} Subgroup-specific thresholds reduces this difference to near zero, where there are small differences, the percent difference across different subgroups is fair (\ie FPR=1e-4).}}\label{fig:percent:difference}
+\end{figure}
 \subsection{\gls{det} analysis}
@@
     \begin{subfigure}[t]{.3\linewidth}
-    \includegraphics[width=.75\linewidth]{figures/curve_vgg16_subgroups-crop.pdf}
+    \includegraphics[width=.71\linewidth]{figures/curve_vgg16_subgroups-crop.pdf}
     \caption{VGG16~\cite{simonyan2014very}}
@@
     \begin{subfigure}[t]{.27\linewidth}
-    \includegraphics[width=.8\linewidth]{figures/curve_resnet50_subgroups-crop.pdf}
+    \includegraphics[width=.75\linewidth]{figures/curve_resnet50_subgroups-crop.pdf}
     \caption{ResNet50~\cite{he2016deep}}
@@
     \begin{subfigure}[t]{.27\linewidth}
-    \includegraphics[width=.8\linewidth]{figures/curve_senet50_subgroups-crop.pdf}
+    \includegraphics[width=.75\linewidth]{figures/curve_senet50_subgroups-crop.pdf}
     \caption{SENet~\cite{hu2018squeeze}}
     \end{subfigure}
-    \caption{\textbf{\gls{det} curves for different CNNs}.
+    \caption{\small{\textbf{\gls{det} curves for different CNNs}.
 \gls{fnr} (\%) (vertical) vs \gls{fpr}  (horizontal, log-scale) for VGG2~\cite{Cao18} models with different backbones (VGG16, Resnet50, SEnet50).
@@
 For each plot, \gls{wm} is the top-performer, while \gls{af} is the worst.
-The the ordering of the curves is roughly the same for each backbone.}\label{fig:sdm-appendix-a}
+The the ordering of the curves is roughly the same for each backbone.}}\label{fig:sdm-appendix-a}
 \end{figure*}
@@
 \caption{\small{\textbf{\gls{tar} at intervals of \gls{far}}.
-For each subgroup, and overall average \gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom).
-Higher is better.
-The proposed shows improvement in all cases.}}\label{tab:ethnicy-far}
+gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom).
+Higher is better.}}\label{tab:ethnicy-far}
 %  \vspace{-2mm}
@@
 For example, the plots generated with Spherface and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst).
-Thus, the dditional \gls{cnn}-based models demonstrate the same phenomena: proportional to the overall model performance, exact in which the ordering subgroups insensitivity in scores space.
+Thus, the additional \gls{cnn}-based models demonstrate the same phenomena: proportional to the overall model performance, exact in which the ordering subgroups insensitivity in scores space.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[t!]
+\begin{center}
+    \caption{\small{\textbf{Human assessment (quantitative).} Subgroups listed per row (\ie human) and column (\ie image).
+Note, most do the best intra-subgroup (\textcolor{blue}{blue}), and second-best (\textcolor{red}{red}) intra-subgroup but inter-gender.
+WF performs the best; WF pairs are most correctly matched.}}
+    % CF shows the least variation, but with the lowest accuracy. WF shows the best accuracy, in human performance and on face pairs of that subgroup.}}
+    \label{tab:humsn-eval-results} 
+    % \vspace{-5mm}
+    %  \vspace{-2mm}
+\footnotesize
+\scalebox{0.94}{
+\begin{tabular}{c}
+\begin{tabular}{c l c c c cc}
+&&\multicolumn{4}{c}{\textbf{Image}}\\
+   \multicolumn{2}{c}{(Acc, \%)}& CF  & CM & WF &WM& Avg\\
+\end{tabular}\\
+\begin{tabular}{c l| r r r r| r}
+\cline{3-7}
+       &CF &\  \textbf{\textcolor{blue}{52.9}}&   \textbf{\textcolor{red}{48.0}}&43.8&44.7 &47.4 \\
+         \multirow{\items}{*}{\rotatebox{90}{\textbf{Human}}} \hspace{-5mm}
+        &CM &  \textbf{\textcolor{red}{45.6}} & \textbf{\textcolor{blue}{50.4}}  & 44.4 &36.2 &44.1 \\
+
+        &WF & 44.7 &43.8 & \textbf{\textcolor{blue}{57.3}}& \textbf{\textcolor{red}{48.0}} & \textbf{48.5} \\
+        &WM & 30.1&\textbf{\textcolor{red}{47.4}} &  45.3 & \textbf{\textcolor{blue}{56.1}} & 44.7\\\cline{3-7}
+        &Avg &  43.3& 47.4&\textbf{47.7} &46.3 &46.2\\
+ \end{tabular}
+ \end{tabular}
+ }
+
+ \end{center}
+%  \vspace{-7mm}
+\end{table}
+
+\subsection{Human evaluation}
+Subjects of a subgroup likely have mostly been exposed to others of the same (Table~\ref{tab:humsn-eval-results} and Fig.~\ref{fig:human-eval}).
+Therefore, it is expected they would be best at labeling their own; similar for the same ethnicity, but another gender.
+Our findings concur.
+Each subgroup is best at labeling their type, and then second best at labeling the same ethnicity but opposite sex.
+Interestingly, each group of images is best tagged by the corresponding subgroup, with the second-best having the same ethnicity and opposite gender.
+On average, subgroups are comparable at labeling images.
+Analogous to the \gls{fr} system, performance ratings differ when analyzing within and across subgroups.
+In other words, performance on \gls{bfw} improved with subgroup-specific thresholds.
+Similarly, humans tend to better recognize individuals by facial cues of the same or similar demographics.
+Put differently, as the recognition performances drop with a global threshold optimized for one subgroup and deployed on another, human performance tends to drop when across subgroups (\ie most familiar intra-subgroup; a performance drop for other, less familiar subgroups).
+
+%  for it better adapts wh specific to subgroup) shift in sensitivity across subgroups when deep encodings are compared, the underlying mechanism
+
 \begin{figure}[t!] 
 	\centering    
-	\includegraphics[width=.85\linewidth] {figures/human_eval.pdf}
-		\caption{\small{\textbf{Qualitative results of human assessment.} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}.
+	\includegraphics[width=.69\linewidth] {figures/human_eval.pdf}
+		\caption{\small{\textbf{Human assessment (qualitative).} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}.
 Accuracy scores shown as bar plots.
@@
 
-\subsection{Human evaluation}
- Quantitative and qualitative results are in Table~\ref{tab:humsn-eval-results} and Fig.~\ref{fig:human-eval}, respectively.
-Subjects of a subgroup likely have mostly been exposed to others of the same.
-Therefore, it is expected they would be best at labeling their own; similar for the same ethnicity, but another gender.
-Our findings concur.
-Each subgroup is best at labeling their type, and then second best at labeling the same ethnicity but opposite sex.
-Interestingly, each group of images is best tagged by the corresponding subgroup, with the second-best having the same ethnicity and opposite gender.
-On average, subgroups are comparable at labeling images.
-Analogous to the \gls{fr} system, performance ratings differ when analyzing within and across subgroups.
-In other words, performance on \gls{bfw} improved with subgroup-specific thresholds.
-Similarly, humans tend to better recognize individuals by facial cues of the same or similar demographics.
-Put differently, as the recognition performances drop with a global threshold optimized for one subgroup and deployed on another, human performance tends to drop when across subgroups (\ie most familiar with the same subgroup; performance seems to drop when faced with another, less familiar subgroup).
-
-%  for it better adapts wh specific to subgroup) shift in sensitivity across subgroups when deep encodings are compared, the underlying mechanism
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}[t!]
-\begin{center}
-    \caption{\small{\textbf{Quantitative of human assessment.} Different subgroups listed per row and column.
-Note that each subject is best within their subgroup, and second-best within the same subgroup but different gender.
-CF shows the least variation, but with the lowest accuracy.
-CM shows the best accuracy, but second to \gls{wm} in deviation from the mean.
-Thus, scores for males vary more than females.}}
-    \label{tab:humsn-eval-results} 
-    %  \vspace{-2mm}
-\footnotesize
-\scalebox{0.8}{
-\begin{tabular}{l c c c c c}
-        &  CF  &      CM     &   WF     &   WM     &  Avg\\\midrule
-        CF &  \textbf{\textcolor{blue}{52.9}}&   \textbf{\textcolor{red}{48.0}}&43.8&44.7 &47.4$\pm$4.1 \\
-        CM &  \textbf{\textcolor{red}{45.6}} & \textbf{\textcolor{blue}{0.504}}  & 44.4 &36.2 &44.1$\pm$5.9 \\
-        WF & 44.7 &43.8 & \textbf{\textcolor{blue}{0.573}}& \textbf{\textcolor{red}{48.0}} & 48.5$\pm$6.2 \\
-        WM & 30.1&\textbf{\textcolor{red}{47.4}} &  45.3 & \textbf{\textcolor{blue}{56.1}} & 44.7$\pm$10.8 \\\midrule
-        Avg &  43.3& 47.4&47.7 &46.3 &46.2$\pm$2.0\\
- \end{tabular}}
- \end{center}
-%  \vspace{-7mm}
-\end{table}
+% \scalebox{0.8}{
+% \begin{tabular}{l c c c c c}
+% \multicolumn{6}{c}{\textbf{Image}}\\
+%         &  CF  &      CM     &   WF     &   WM     &  Avg\\\midrule
+%         CF &  \textbf{\textcolor{blue}{52.9}}&   \textbf{\textcolor{red}{48.0}}&43.8&44.7 &47.4$\pm$4.1 \\
+%         CM &  \textbf{\textcolor{red}{45.6}} & \textbf{\textcolor{blue}{0.504}}  & 44.4 &36.2 &44.1$\pm$5.9 \\
+%         WF & 44.7 &43.8 & \textbf{\textcolor{blue}{0.573}}& \textbf{\textcolor{red}{48.0}} & 48.5$\pm$6.2 \\
+%         WM & 30.1&\textbf{\textcolor{red}{47.4}} &  45.3 & \textbf{\textcolor{blue}{56.1}} & 44.7$\pm$10.8 \\\midrule
+%         Avg &  43.3& 47.4&47.7 &46.3 &46.2$\pm$2.0\\
+%  \end{tabular}}
+%  \end{center}
+% %  \vspace{-7mm}
+% \end{table} 
+% \newcommand\items{4}   %Number of classes
+% \arrayrulecolor{white} %Table line colors
+% \noindent\begin{tabular}{cc*{\items}{|E}|}
+% \multicolumn{1}{c}{} &\multicolumn{1}{c}{} &\multicolumn{\items}{c}{Image} \\ \hhline{~*\items{|-}|}
+% \multicolumn{1}{c}{} & 
+% \multicolumn{1}{c}{} & 
+% \multicolumn{1}{c}{CF} & 
+% \multicolumn{1}{c}{{CM}} & 
+% \multicolumn{1}{c}{{WF}}&
+% \multicolumn{1}{c}{{WM}}\\ \hhline{~*\items{|-}|}
+% \multirow{\items}{*}{\rotatebox{90}{Labeler}} 
+% &Class A  & 100   & 0  & 10   \\ \hhline{~*\items{|-}|}
+% &Class B  & 10   & 80  & 10   \\ \hhline{~*\items{|-}|}
+% &Class C  & 30   & 0   & 70   \\ \hhline{~*\items{|-}|}
+% &Class C  & 30   & 0   & 70   \\ \hhline{~*\items{|-}|}
+% \end{tabular}
 
@@
 We used an \textit{off-the-shelf} Sphereface, hypothesizing this SOTA CNN suffers from bias because of the imbalanced train-set.
-Once established that the results do suffer from problems of bias, we observed that the same threshold across ethnic and gender subgroups leads to differences in the FPR up to a factor of two, which is seemingly the cause of the frenzy about bias in FR in main-stream media.
+Once established that the results do suffer from problems of bias, we observed that the same threshold across ethnic and gender subgroups leads to differences in the \gls{fpr} up to a factor of two.
+Furthermore, the percent difference from in which is seemingly the cause of the frenzy about bias in FR in main-stream media.
 Furthermore, we ameliorate these differences with a per-subgroup threshold, leveling out FPR, and achieving a higher TPR.

### 2020-04-13 cvprworkshop.tex
moved back into section (ie organized by file)

@@
 \usepackage{acronym}
-
+\usepackage{collcell}
 \newcommand{\minitab}[2][c]{\begin{tabular}{#1}#2\end{tabular}}
@@
 % \let\vec\mathbf
+% \usepackage[T1]{fontenc}
+% \usepackage[latin9]{inputenc}
 
-% \cvprfinalcopy % *** Uncomment this line for the final submission
+\def\colorModel{hsb} %You can use rgb or hsb
+\usepackage{babel}
+% \usepackage[table]{xcolor}
+% 
+% \def\colorModel{hsb} %You can use rgb or hsb
 
-\def\cvprPaperID{0002} % *** Enter the CVPR Paper ID here
+\usepackage{hhline}
+\newcommand\ColCell[1]{
+  \pgfmathparse{#1<50?1:0}  %Threshold for changing the font color into the cells
+    \ifnum\pgfmathresult=0\relax\color{white}\fi
+  \pgfmathsetmacro\compA{0}      %Component R or H
+  \pgfmathsetmacro\compB{#1/100} %Component G or S
+  \pgfmathsetmacro\compC{1}      %Component B or B
+  \edef\x{\noexpand\centering\noexpand\cellcolor[\colorModel]{\compA,\compB,\compC}}\x #1
+  } 
+\newcolumntype{E}{>{\collectcell\ColCell}m{0.4cm}<{\endcollectcell}}  %Cell width
+\newcommand*\rot{\rotatebox{90}}
+
+\cvprfinalcopy % *** Uncomment this line for the final submission
+
+% \def\cvprPaperID{0001} % *** Enter the CVPR Paper ID here
 \def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
@@
 \title{Face Recognition: Too Bias, or Not Too Bias?}
+
+%use this in case of several affiliations
+\author{\parbox{16cm}{\centering
+    {\large Joseph P Robinson$^1$, Gennady Livitz$^2$, Yann Henon$^2$, Can Qin$^1$,\\ Yun Fu$^1$, and Samson Timoner$^2$}\\
+    {\normalsize
+    \hspace{-.4in}$^{1}$Northeastern University\hspace{.7in} $^{2}$ISM Connect}}
+    % \thanks{}% <-this % stops a space
+}
 
@@
 Furthermore, we do a human evaluation to measure the bias in humans, which supports the hypothesis that such a bias exists in human perception.
-For the \gls{bfw} database, source code, and more, visit \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}
+For the \gls{bfw} database, source code, and more, visit \href{https://github.com/visionjo/facerec-bias-bfw}{https://github.com/visionjo/facerec-bias-bfw}.
+% \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}
 \end{abstract}
@@
 % \begin{bibunit}
+\input{sections/introduction}
+\input{sections/relatedworks}
 
-\glsresetall
-\section{Introduction}
-    As more of society becomes integrated with \gls{ml}, bias, fairness, and the formalization of \gls{ml} standards attract more attention~\cite{10.1007/978-3-030-13469-3_68, anne2018women, wang2018racial}.
-Thus, an effect of the growing dependency on \gls{ml} are growing concerns for biased and unfair algorithms, \eg untrustworthy and racist \gls{fr} systems~\cite{england2019,snow2018}.
-
-    Typically, \glspl{cnn} are trained on faces identified by a detection system.
-Specifically, for \gls{fr}, the goal is to encode faces in an N-dimensional space such that samples of the same identity are neighbors and those of different identity of further apart.
-Thus, the goal is to identify subjects with face encodings.
-We can deploy a \gls{cnn} to encode faces, and then compare via a similarity score: assume pairs with a high enough score are \emph{genuine}; else, an \emph{imposter}.
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[width=\linewidth]{figures/fig1-crop.pdf}
-%     \caption{Caption}
-%     \label{fig:my_label}
-% \end{figure}
 \begin{figure}[t!]
     \centering
-        \includegraphics[width=.9\linewidth]{figures/montage.pdf}
-        \caption{\textbf{\gls{bfw}.} The average face of its different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender.
+    \includegraphics[width=.8\linewidth]{figures/montage.pdf}
+    \caption{\small{\textbf{\gls{bfw}.} Average face of the different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender.
 The others represent the ethnicity and gender of the race and gender, respectively.
-Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}
+Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}}
     \label{fig:avg-faces}
@@
 
-    When comparing scores, typically, a fixed threshold sets the decision boundary.
-Thus, features of the same identity must satisfy a criterion via a single value~\cite{deng2019arcface, liu2017sphereface, wang2018additive, wang2018cosface}.
-However, we found that a single (\ie global) threshold is a crude measure that leads to skewed errors - the held-out set used to determine the threshold tends to share the same distribution with the test data, which favors specific demographics that are a majority.
-That skew, the difference in the performance of an algorithm of certain demographics, is our definition of bias.
-A key question is: \emph{is \gls{fr} too biased, or not?}
-
-\begin{table*}[!t]
-    \centering
-    \caption{\small{\textbf{Database stats and nomenclature, optimal thresholds ($t_o$), and accuracy scores.} \textit{Header:} Subgroup definitions.
-\textit{Top-row:} Statistics of \gls{bfw}.
-\textit{Middle:} Number of pairs for each partition.
-\textit{Bottom:} Accuracy with a global threshold $t_g$, the optimal threshold $t_o$, and accuracy with $t_o$ per subgroup.
-Specifically, $t_g$=0.259$\pm$0.002 when averaged across the five folds.
-Columns grouped by race and then further split by gender.
-Out of millions of pairs, accuracy is inconsistent across subgroups.
-Furthermore, $F$ tends to perform inferior to that of $M$ for all but B.}}\label{tab:ethnic-splits}
-    \scalebox{.75}{
-     \resizebox{\textwidth}{!}{%
-    \begin{tabular}{r c c c c c c c c l}
-        \toprule
-        & \multicolumn{2}{c}{Asian (A)} & \multicolumn{2}{c}{Black (B)}  & \multicolumn{2}{c}{Indian (I)}& \multicolumn{2}{c}{White (W)}\\
-        \cmidrule(l){2-3} \cmidrule(l){4-5} \cmidrule(l){6-7}\cmidrule(l){8-9} % spanning less than the full width of the table - you can add (r) or (l) just before the opening curly bracket to shorten the rule on the left or right side
-         & Female (AF) & Male (AM) & BF & BM& IF & IM & WF & WM&Aggregated\\ % Column names row
-        \midrule
-
-       \# Faces  &  2,500&  2,500& 2,500 & 2,500& 2,500 & 2,500 & 2,500 & 2,500 &20,000 \\ % row 1
-        \# Subjects & 100& 100& 100  & 100  & 100  & 100& 100 &100&800  \\ % row 2
-        \# Faces / Subject  & 25 & 25    & 25 & 25 & 25  & 25  &  25 & 25 & 25\\ % row 3
-        % \midrule
-\specialrule{.01em}{.05em}{.05em}
-            \# Positive Pairs &  30,000&  30,000& 30,000 &30,000 & 30,000 &30,000&30,000 & 30,000 &240,000 \\ % row 1
-        \# Negative Pairs & 85,135&  85,232& 85,016  & 85,141  & 85,287  & 85,152& 85,223 &85,193&681,379  \\ % row 2
-
-        \# Pairs (Total) & 115,135 & 115,232    &115016 &115,141 & 115287  & 115,152  &  115,223& 115193 & 921,379\\ % row 3
-        % \midrule
-        % \specialrule{.01em}{.05em}{.05em}
-        % Acc$@t_g$ & 0.941 & 0.952  & 0.967 & 0.960 & 0.958 &0.962  & 0.973 & 0.981 &0.962 $\pm$0.012 \\
-        % $t_o$ & 0.255 &  0.246 & 0.277 &0.242  &0.297 & 0.264& 0.233 &0.239 &0.257$\pm$0.006\\ % row 4
-
-        % Acc$@t_o$ & 0.942 &0.952 &0.968 &0.961 &0.960 &0.962 & 0.974 &0.982 & 0.963 $\pm$ 0.012\\
-        \bottomrule
-    \end{tabular}}
-    % \glsunset{af}
-    }
-    % % \glsunset{af}
-    % \glsunset{am}
-    % \glsunset{bf}
-    % \glsunset{bm}
-    % \glsunset{if}
-    % \glsunset{im}
-    % \glsunset{wf}
-    % \glsunset{wm}
-    % \vspace{-12pt}
-\end{table*}
-
-    Making matters more challenging is that race and ethnicity are loosely defined.
-For example, the US Census Bureau allows an individual to self-identify race.\footnote{\scriptsize\href{https://www.census.gov/mso/www/training/pdf/race-ethnicity-onepager.pdf}{www.census.gov/mso/www/training/pdf/race-ethnicity-onepager}} For this work, we define subgroups as specific sub-populations with face characteristics similar to those found in a region.
-
-    The adverse effects of a global threshold are two-fold: \textbf{(1)} the mapping produced by a \gls{cnn} is nonuniform.
-Therefore, distances between pairs of faces in different demographics vary in distribution of similarity scores (Fig~\ref{fig:detection-model}); \textbf{(2)} the evaluation set is imbalanced as well.
-Particular demographics making up a majority of the population will carry most weight on the reported performance ratings.
-Reported results favor the common traits over the underrepresented.
-Alas, demographics like gender, ethnicity, race, and age are underrepresented in most public datasets~\cite{merler2019diversity, wang2018racial}.
-
-    % The result is various types of biases in FR systems in favor of or against particular demographics remain a question.
-
-    For \textbf{(1)} we propose subgroup-specific (\ie optimal) thresholds; to address \textbf{(2)}, we introduce a new benchmark for measuring bias in \gls{fr}, \gls{bfw} (Table~\ref{tab:compared} \&~\ref{tab:ethnic-splits}).
-\gls{bfw} allows for a fair evaluation of \gls{fr} systems, while enabling demographic-specific ratings to be reported.
-We use \gls{bfw} to gain a deeper understanding of the extent of bias present in \gls{soa} \gls{cnn} used to encode faces.
-Then, we suggest a mechanism to mitigate problems of bias with more balanced performance ratings for different demographics.
-Specifically, we propose using an adaptive threshold that varies depending on the characteristics of detected facial attributes (\ie gender and ethnicity, Fig~\ref{fig:avg-faces}).
-We show an increase in accuracy with a balanced performance for different subgroups.
-Similarly, we show a positive effect of adjusting the similarity threshold based on the facial features of matched faces.
-Thus, selective use of similarity thresholds in current \gls{soa} \gls{fr} systems provides more intuition in \gls{fr} research with a method easy to adopt in practice.
-
-    The contributions in this paper are 3-fold: (1) Built a balanced dataset as a proxy to measure verification performance per subgroup; (2) Analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup balances performance; (3) Showed bias causes skewness in ratings across demographics-- the typical use of a global threshold unfavorable.
-We mitigate the problem via adaptive thresholds; (4) Conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
-    % \begin{enumerate}
-    %     \item Built a balanced dataset as a proxy to measure verification performance per subgroup.
-    %     \item Analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup significantly boost and balance performances.
-    %     \item Showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
-    %     \item Conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
-    % \end{enumerate}
-
-    \begin{table*}[t!]
-
-        \centering
-        \caption{\textbf{\gls{bfw} compared to related datasets.} Our \gls{bfw} is balanced across ID, gender, and ethnicity (Table~\ref{tab:ethnic-splits}).
-Compared with \gls{dp}, \gls{bfw} provides more samples per subject and subgroups per set, while only using a single resource, VGG2.
-\gls{rfw}, on the other hand, supports a different task (\ie domain adaptation), and focuses on race-distribution, but not the distribution of identities.}
-        \scriptsize
-        \begin{tabular}{rccccccc}%x{10mm}x{3mm}x{10mm}x{8mm}}%\toprule
-
-            \multicolumn{2}{c}{Database} & \multicolumn{3}{c}{Number of}& \multicolumn{3}{c}{Balanced Labels}\\
-            \cmidrule(lr){1-2}	\cmidrule(lr){3-5} \cmidrule(lr){6-8}
-            Name & Source Data & Faces &  IDs & Subgroups & ID & Ethnicity & Gender\\\midrule
-            \gls{dp}~\cite{demogPairs}     & CASIA-W~\cite{yi2014learning}, VGG~\cite{schroff2015facenet} \&VGG2~\cite{Cao18} & 10,800& 600 & 6 &\checkc& \checkc &\checkc \\
-            \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &$\approx$80,000&$\approx$12,000& 4 & \xmark & \checkc &\xmark \\
-            \gls{bfw} (ours) & VGG2 & 20,000 & 800 &8 & \checkc & \checkc &\checkc \\\bottomrule
-        \end{tabular}
-        \label{tab:compared}
-    \end{table*}
-
-\section{Background Information}
-    \subsection{Bias in \gls{ml}}
-        The progress and the commercial value of \gls{ml} is exciting; however, the trust of society in \gls{ml} will take time to build due to inherent biases.
-The exact definitions and implications of bias vary between sources, as do its sources and types.
-A common theme is that bias hinders performance ratings in ways that skew to a particular sub-population-- the source varies, whether from humans~\cite{windmann1998subconscious}, data and label types~\cite{tommasi2017deeper}, \gls{ml} models~\cite{amini2019uncovering, kim2019learning}, or evaluation protocols~\cite{stock2018convnets}.
-For instance, a vehicle-detection model might miss cars if training data was mostly trucks.
-In practice, many \gls{ml} systems learn on biased data, which could be detrimental for society.
-
-    \subsection{Bias in \gls{fr}}
-        Problems of bias in \gls{fr} have been driven by different motivations; therefore, solutions have been posed for different matters.
-To name a few: data augmentation~\cite{yin2019feature}, one-shot learning~\cite{ding2018one}, demographic parity and fairness with priority on privacy~\cite{huang2018generative}, domain adaptation~\cite{wang2018racial}, differences in face-based attributes across demographics~\cite{wang2018they}, and even data exploration~\cite{muthukumar2019}.
-Yin~\etal proposed to augment the feature space of underrepresented classes using other classes with a diverse collection of samples to encourage distributions of underrepresented classes to more closely resemble that of the others~\cite{yin2019feature}.
-Similarly, others formulated the imbalanced class problem as one-shot learning, where a \gls{gan} was trained to generate face features to augment classes with fewer samples~\cite{ding2018one}.
-\gls{gapf} was proposed to create fair representations of the data in a quantifiable way, allowing for the finding of a de-correlation scheme from the data without access to its statistics~\cite{huang2018generative}.
-Wang~\etal defined subgroups at a finer-level (\ie Chinese, Japanese, Korean), and determined the familiarity of faces inter-subgroup~\cite{wang2018they}.
-Genders have also been used to make subgroups (\eg for analysis of gender-based face encodings~\cite{muthukumar2019}).
-Most recently,~\cite{wang2018racial} proposed adapting domains to bridge the bias gap by knowledge transfer, which was supported by a novel data collection, \gls{rfw}.
-The release of \gls{rfw} occurred after \gls{bfw} was built - although similar in terms of demographics, \gls{rfw} uses faces from MSCeleb~\cite{guo2016ms} for testing, and assumes CASIA-Face~\cite{yi2014learning} and VGG2~\cite{Cao18} were used to train.
-In contrast, our \gls{bfw} assumes VGG2 as the test set.
-        % As shown in \cite{wang2018racial}, MSCeleb is highly imbalanced, primarily consisting of images of white individuals. 
-        % For this, we expect bias from data, for this is the training set. 
-        Furthermore, \gls{bfw} balances gender and race: \gls{rfw} splits subgroups by gender and race, while \gls{bfw} splits by gender, race, or gender and race). 
-        % Thus, \gls{rfw} and \gls{bfw} complement one another. % - both add demographic for subjects with faces in renowned, large-scale \gls{fr} datasets.
-
-Most similar to us is~\cite{das2018, demogPairs, lopez2019dataset, srinivas2019face} - each were motivated by insufficient paired data for studying bias in \gls{fr}; then, problems were addressed using labeled data from existing image collections.
-Specifically, Hupont~\etal curated a set of faces based on racial demographics (\ie \gls{a}, \gls{b}, and \gls{w}) called \gls{dp}~\cite{demogPairs}, while~\cite{srinivas2019face} honed in on adults versus children called \gls{itwcc}.
-Like the proposed \gls{bfw}, both were built by sampling existing databases, but with the addition of tags for the respective subgroups of interest.
-Besides, the additional data of \gls{bfw} (\ie added an additional subgroup \gls{i}, along with additional subjects with more faces for all subgroups), we also further split subgroups by gender.
-Furthermore, we focus on the problem of facial verification and the different levels of sensitivity in cosine similarity scores per subgroup.
-
-\subsection{Human bias in \gls{ml}}
-Bias is not unique to \gls{ml} - humans are also susceptible to a perceived bias.
-In fact, biases exist across race, gender, and even age~\cite{10.1007/978-3-030-13469-3_68, bar2006, meissner2001, nicholls2018}.
-Wang~\etal showed machines surpass human performance in discriminating between Japanese, Chinese, or Korean faces by nearly 150\%~\cite{wang2018they}, as humans just pass random (\ie 38.89\% accuracy).
-
-We expect human bias to skew to their own genders and races.
-For this, we measure the human perception with face pairs of different subgroups (Section~\ref{subsec:human-assessment}).
-The results concur with~\cite{wang2018they}, as we also recorded overall averages below random ($<$50\%).
-%The details on settings and number of submissions per demographics are discussed in Section~\ref{}.
-
 \glsunset{if}\glsunset{im}\glsunset{af}\glsunset{am}\glsunset{bf}\glsunset{bm}\glsunset{wf}\glsunset{wm}
-\section{The BFW Benchmark and Dataset}
-
-We now discuss the \gls{bfw} dataset (\ie collection and preparation), along with the evaluation protocols. We conclude this section by reviewing the settings followed to survey bias in humans.
-
-\subsection{The data}
-Problems of bias in \gls{fr} motivated us to build \gls{bfw}.
-The data evenly represent various subgroups partitioned by demographic.
-Inspired by \gls{dp}~\cite{demogPairs}, the specification of \gls{bfw} follows in suit, but with additional subgroups (\ie \gls{if} and \gls{im}), an increase in the number of subjects per subgroup, and many more pairs (Table~\ref{tab:ethnic-splits}).
-
-\vspace{1mm}
-\noindent\textbf{Compiling subject list.} 
-Subjects were sampled from VGG2~\cite{Cao18} - unlike others that depend on multiple sources, \gls{bfw} has fewer potential conflicts in train and test overlap with existing models.
-We used pre-trained ethnicity~\cite{ambekar2009name} and gender~\cite{levi2015age} classifiers to find candidates for the different subgroups.
-
-\vspace{1mm}
-\noindent\textbf{Detecting faces.} Faces were detected using MTCNN~\cite{zhang2016joint}.\footnote{\href{https://github.com/polarisZhao/mtcnn-pytorch}{https://github.com/polarisZhao/mtcnn-pytorch}} Then, assigned into one of two sets.
-Faces within detected bounding box (BB) regions extended out 130\% in each direction, with zero-padding as the boundary condition made-up one set.
-The second set were faces aligned and cropped for Sphereface~\cite{liu2017sphereface} (see the next step).
-Also, coordinates of the BB and the five landmarks from \gls{mtcnn} were stored as part of the static, raw data.
-For samples with multiple face detections, we used the BB area times the confidence score of the \gls{mtcnn} to determine the face most likely to be the subject of interest, with the others set aside and labeled \textit{miss-detection}.
-
-\begin{figure}[!t] 
-	\centering    
- \glsunset{fpr}
-  \glsunset{fnr}
-	\includegraphics[width=\linewidth]{figures/detcurve-improved.pdf}
-		\caption{\textbf{\gls{det} curves.} \emph{Top-left}: per gender.
-\emph{Top-right}: per ethnicity.
-\emph{Bottom}: per subgroup (\ie combined).
-A dashed line shows a difference by about 2$\times$ in \gls{fpr} for the same threshold $\theta_{const}$.
-\gls{fnr} is the match error count (closer to the bottom is better).}
-		\glsreset{det}\glsreset{fpr}
-\label{fig:detcurves} 
-\end{figure}
-
-\vspace{1mm}
-\noindent\textbf{Validating labels.} 
-Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Sphereface~\cite{liu2017sphereface}: applying an affine transformation to align faces according to predefined eye locations, each face fed through twice (\ie the original and horizontally flipped), with the final encoding obtained by fusing the two features by average pooling (\ie 512 D).
-A matrix of cosine similarity scores was then generated for each subject and removed samples (\ie rows) with median scores below threshold $\theta=0.2$ (set manually).
-Mathematically, the $n^{th}$ sample for the $j^{th}$ subject with $N_j$ faces was removed if the ordinal rank of its score $n = \frac{P\times N}{100}\geq\theta$, where $P=50$.
-In other words, the median (\ie 50 percentile) of all scores for a faces with respect to all of faces for the respective subject must pass a threshold of $\theta=0.2$; else, the face is dropped.
-This allowed us to quickly prune \gls{fp} face detections.
-Following~\cite{robinson2016families, robinson2018visual}, we built a JAVA tool to visually validate the remaining faces.
-For this, the faces were ordered by decreasing confidence, with confidence set as the average score, and then displayed as image icons on top toggling buttons arranged as a grid in a sliding pane window.
-Labeling then consisted of going subject-by-subject and flagging faces of \emph{imposters}.
-
-\vspace{1mm}
-\noindent\textbf{Sampling faces and creating folds.} We created lists of pairs in five-folds with subjects split evenly per subject and without overlap across folds.
-Furthermore, a balance in the number of faces per was obtained by sampling twenty-five faces at random from each.
-Next, we generated a list of all the face pairs per subject, resulting in $\sum_{l=1}^{L}\sum_{k=1}^{K_d} {N_k \choose 2}$ positive pairs, where the number of faces of all $K_l$ subjects $N_k=25$  for each of the $L$ subgroups (Table~\ref{tab:ethnic-splits}).
-Next, we assigned subjects to a fold.
-To preserve balance across folds, we sorted subjects by the number of pairs and then started assigning to alternating folds from the one with the most samples.
-Note, this left no overlap in identity between folds.
-Later, a negative set from samples within the same subgroup randomly matched until the count met that of the positive.
-Finally, we doubled the total count with negative pairs from across subgroups but in the same fold.
-\begin{figure}[t!] 
-	\glsunset{fpr}
-	\glsunset{sdm}
-	\centering
-	\centering
-	\includegraphics[width=1\linewidth]{figures/violinplots.pdf}
-		\caption{\textbf{\Gls{sdm} across subgroups.} Scores of \emph{imposters} have medians around 0.3 but with variations in upper percentiles; \emph{genuine} pairs vary in mean and spread (\eg \gls{af} has more area of overlap).
-A threshold varying across different subgroups yields a constant \gls{fpr}.} \label{fig:detection-model}
-		\glsreset{sdm}
-\end{figure}
-
-% \vspace{-5pt}
-\subsection{Problem formulation}\label{subsec:pf} 
-% \gls{lfw}~\cite{LFWTech}, a commonly used benchmark for ~\gls{fr}, reserves specific train-test face-pair lists. 
-\Gls{fv} is the special case of the two-class (\ie boolean) classification.
-Hence, pairs are labeled as the ``same'' or ``different'' \textit{genuine} pairs (\ie \textit{match}) or \textit{imposter} (\ie \textit{mismatch}), respectively.
-This formulation (\ie \gls{fv}) is highly practical for applications like access control, re-identification, and surveillance.
-Typically, training a separate model for each unique subject is infeasible.
-Firstly, the computational costs compound as the number of subjects increase.
-Secondly, such a scheme would require model retraining each time a new person is added.
-Instead, we train models to encode facial images in a feature space that captures the uniqueness of a face, to then determine the outcome based on the output of a scoring (or distance) function.
-Formally put:
-\begin{equation}\label{eg:matcher}
-    f_{boolean}(\vec{x}_i, \vec{x}_j) = d(\vec{x}_i, \vec{x}_j) \leq \theta
-\end{equation}
-
-where $f_{boolean}$ is the \textit{matcher} of the feature vector $\vec{x}$ for the $i^{th}$ and $j^{th}$ sample~\cite{LFWTech}.
-
-Cosine similarity is used as the \emph{matcher} in Eq~\ref{eg:matcher} the closeness of $i^{th}$ and $j^{th}$ features, \ie
-$
-s_l= \frac{f_i\cdot f_j}{||f_i||_2||f_j||_2}
-$ is the closeness of the $l^{th}$ pair.
-
-\subsection{Human assessment}\label{subsec:human-assessment}
-We evaluated human on face pairs focusing on two racial groups: Chinese and Caucasians.
-To focus on the experiment, we honed-in on two groups, white Americans (W) and Chinese from China (C).
-The purpose was to minimize variability by only analyzing the subsets of the broader groups of whites and Asians.
-
-Samples were collected by recruiting subjects from multiple sources (\eg social media, email lists, and family/friends)-- a total of 120 participants were sampled at random from all the submissions that were (1) complete and (2) from a W or C participant.
-Specifically, there were 60 W and 60 C, both with \gls{m} and \gls{f} split evenly.
-A total of 50 face pairs of non-famous ``look-alikes'' were collected from the internet, with 20 ({\emph WA}) and 20 ({\emph C}) pairs (male and female split evenly).
-The other 10 pairs are of others (\eg Hispanic/ Latino, Japanese, African).
-The survey was created, distributed, and recorded via \href{https://paperform.co}{PaperForm}.
+\input{sections/section3}
 
@@
        \centering
-	\includegraphics[width=.85\linewidth]{figures/confusion.pdf}
-		\caption{\small{\textbf{Confusion matrix.} Percent error (Rank 1) for all faces of \gls{bfw} queried against all others.
-Notice errors concentrate within a subgroup, consistent with the \gls{sdm} in Fig.~\ref{fig:detection-model} (\ie \gls{af} show worst performance, and is mostly confused with faces of the same demographic).
-This plot is evidence that while race/ethnicity may be challenging to define, the subgroups are meaningful.}}
-		\label{fig:confusion} 
+    \includegraphics[width=.95\linewidth]{figures/adaptive_fpr_percent_diff_complete.pdf}
+    \caption{\small{\textbf{Percent difference from intended \gls{fpr}.} \emph{Top:} $t_g$ yields \gls{fpr} the span as large as 2$\times$ (\ie 200\%) that intended (\ie \gls{wm} for 1e-4).
+Furthermore, \gls{f} subgroups tend to perform worse than intended for all cases (while \gls{m} tend to overshoot intended performance, with exception of \gls{im} in for \gls{fpr}=1e-4).
+\emph{Bottom:} Subgroup-specific thresholds reduces this difference to near zero, where there are small differences, the percent difference across different subgroups is fair (\ie FPR=1e-4).}}\label{fig:percent:difference}
 \end{figure}
 
-\section{Results and Analysis}
-A single \gls{cnn} was used as a means to control the experiments.
-For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https://github.com/clcarwin/sphereface\_pytorch}} As reported in~\cite{wang2019racial}, \gls{lfw} has about a 13\%, 14\%, 3\%, and 70\% ratio in Asians, Blacks, Indians, and Whites, respectively; furthermore, CASIA-Web is even more unbalanced (again, as reported in~\cite{wang2019racial}), with about  3\%, 11\%, 2\%, and 85\% for the same subgroups.
+\input{sections/analysis}
 
-% Specifically, we analyze three \glspl{cnn} variants: we encode faces using VGG-16~\cite{simonyan2014very},  50-layer residual network (\ie ResNet-50)~\cite{he2016deep}, and \gls{senet} (\ie SENet-50)~\cite{hu2018squeeze}, with results of only the latter (\ie the best performing \gls{senet}) model used throughout the main paper.
-Results for the other two, alongside \gls{senet} for comparison, are in the Supplemental Material.
-
-\glsunset{am}\glsunset{af}\glsunset{bm}\glsunset{bf}\glsunset{im}\glsunset{if}\glsunset{wm}\glsunset{wf}
-
-\subsection{Score analysis}
-Fig.~\ref{fig:detection-model} shows score distributions for faces of the same (\ie \emph{Genuine}) and different (\ie \emph{Imposter}) identity, with a subgroup per \gls{sdm} plot.
-Notice that score distributions for imposters tend to peak about zero for all subgroups, and with minimal deviation comparing modes of the different plots.
-On the other hand, the score distribution of the \emph{genuine} pairs varies across subgroups in location (\ie score value) and spread (\ie overall shape).
-Asian Fem
-Fig.~\ref{fig:confusion} shows the confusion matrix of the subgroups.
-A vast majority of errors occur in the intra-subgroup.
-It is interesting to note that while the definition of  each group  based on ethnicity and race may not be crisply defined, the confusion matrix indicates that in practice the \gls{cnn} finds that the groups are effectively separate.
-The categories are, therefore, meaningful in the context of \gls{fr}.
-
-\subsection{\gls{det} analysis}
-\glsunset{m}
-\glsunset{f}
-
-\begin{figure*}[h!]
-    \centering
-    \begin{subfigure}[t]{.3\linewidth}
-    \includegraphics[width=.75\linewidth]{figures/curve_vgg16_subgroups-crop.pdf}
-    \caption{VGG16~\cite{simonyan2014very}}
- \end{subfigure}
-    \begin{subfigure}[t]{.27\linewidth}
-    \includegraphics[width=.8\linewidth]{figures/curve_resnet50_subgroups-crop.pdf}
-    \caption{ResNet50~\cite{he2016deep}}
-   \end{subfigure}
-    \begin{subfigure}[t]{.27\linewidth}
-    \includegraphics[width=.8\linewidth]{figures/curve_senet50_subgroups-crop.pdf}
-    \caption{SENet~\cite{hu2018squeeze}}
-    \end{subfigure}
-    \caption{\textbf{\gls{det} curves for different CNNs}.
-\gls{fnr} (\%) (vertical) vs \gls{fpr}  (horizontal, log-scale) for VGG2~\cite{Cao18} models with different backbones (VGG16, Resnet50, SEnet50).
-Lower is better.
-For each plot, \gls{wm} is the top-performer, while \gls{af} is the worst.
-The the ordering of the curves is roughly the same for each backbone.}\label{fig:sdm-appendix-a}
-\end{figure*}
-
-\gls{det} curves averaged across 5-folds show per-subgroup trade-offs (Fig.~\ref{fig:detcurves}).
-Note that \gls{m} performs better than \gls{f}, precisely as one would expect from the tails of score-distributions for \emph{genuine} pairs (Fig.~\ref{fig:detection-model}).
-\Gls{af} and \gls{if} perform the worst.
-
-\begin{table}[b!]
-\glsunset{tar}
-\glsunset{far}
-\caption{\small{\textbf{\gls{tar} at intervals of \gls{far}}.
-For each subgroup, and overall average \gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom).
-Higher is better.
-The proposed shows improvement in all cases.}}\label{tab:ethnicy-far}
-%  \vspace{-2mm}
-\begin{center}
-\scriptsize
-\begin{tabular}{l c c c c c}
-     \gls{far} & 0.3 & 0.1 & 0.01 & 0.001 & 0.0001\\\midrule
-    \multirow{2}{.1mm}{\textbf{\gls{af}}} &0.990 & 0.867 & 0.516 & 0.470 & 0.465\\[-4pt]
-        &1.000 & 0.882 & 0.524 & 0.478 & 0.474\\[-1pt]
-    \multirow{2}{3mm}{\textbf{\gls{am}}} &0.994 & 0.883 & 0.529 & 0.482 & 0.477\\[-4pt]
-        &1.000 & 0.890 & 0.533 & 0.486 & 0.482\\[-1pt]
-    \multirow{2}{3mm}{\textbf{\gls{bf}}} &0.991 & 0.870 & 0.524 & 0.479 & 0.473\\[-4pt]
-        &1.000 & 0.879 & 0.530 & 0.484 & 0.480\\[-1pt]
-    \multirow{2}{3mm}{\textbf{\gls{bm}}} &0.992 & 0.881 & 0.526 & 0.480 & 0.474\\[-4pt]
-        &1.000 & 0.891 & 0.532 & 0.485 & 0.480\\[-1pt]
-    \multirow{2}{3mm}{\textbf{\gls{if}}} &0.996 & 0.881 & 0.532 & 0.486 & 0.481\\[-4pt]
-        &1.000 & 0.884 & 0.534 & 0.488 & 0.484\\[-1pt]
-    \multirow{2}{3mm}{\textbf{\gls{im}}} &0.997 & 0.895 & 0.533 & 0.485 & 0.479\\[-4pt]
-        &1.000 & 0.898 & 0.535 & 0.486 & 0.481\\[-1pt]
-    \multirow{2}{3mm}{\textbf{\gls{wf}}} &0.988 & 0.878 & 0.517 & 0.469 & 0.464\\[-4pt]
-        &1.000 & 0.894 & 0.526 & 0.478 & 0.474\\[-1pt]
-    \multirow{2}{3mm}{\textbf{\gls{wm}}} &0.989 & 0.896 & 0.527 & 0.476 & 0.470\\[-4pt]
-        &1.000 & 0.910 & 0.535 & 0.483 & 0.478\\[-1pt]
-    \midrule
-    \multirow{2}{3mm}{\textbf{Avg.}} &0.992 & 0.881 & 0.526 & 0.478 & 0.474\\[-4pt]
-        &1.000 & 0.891 & 0.531 & 0.483 & 0.479\\[-10pt]
-\end{tabular}
-\end{center}
-\glsreset{tar}
-\glsreset{far}
-%  \vspace{-3mm}
-
-\end{table}
-
-The gender-based \gls{det} curve shows a difference in performances between \gls{m} and \gls{f} with a fixed threshold (dashed line).
-Similar effects exist for the other curves as well (lines omitted to declutter).
-For many \gls{fr} applications, systems are operated at the highest \gls{fpr} allowed, so the line of constant threshold indicates that a single threshold produces different operating points (\ie \gls{fpr}), which is undesirable.
-The difference in \gls{fpr} is approximately a factor of two-- quite large indeed.
-If this is the case in an industrial system, one would expect a difference in about double the false positives to be reported based on subgroup alone.
-The potential ramifications of such a bias should not be overlooked, which it has not as of lately-- gaining the interest in even main-stream media ~\cite{england2019,snow2018}.
-
-% \subsubsection{Other \gls{cnn} models.}~\label{app:sec:other:models}
-Variations in optimal threshold exist across models (Fig.~\ref{fig:sdm-appendix-a}).
-Like in Fig.~\ref{fig:detcurves}, the \gls{det} curves for three \gls{cnn}-based models, each trained on VGG2 with softmax but with different backbones.\footnote{\href{https://github.com/rcmalli/keras-vggface}{https://github.com/rcmalli/keras-vggface}} Notice similar trends across subgroups and models, which is consistent with  Sphereface as well (Fig.~\ref{fig:detcurves}).
-For example, the plots generated with Spherface and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst).
-Thus, the dditional \gls{cnn}-based models demonstrate the same phenomena: proportional to the overall model performance, exact in which the ordering subgroups insensitivity in scores space.
-
-\subsection{Verification threshold} \label{subsec:analysis:verification}
-We seek to reduce the bias between subgroups.
-Such that an operating point (\ie \gls{fpr}) is constant across subgroups.
-To accomplish that, we used a per subgroup threshold.
-In \gls{fv}, we consider one image as the query and all others as the test.
-For this, the ethnicity of the query image is assumed.
-We can then examine the \gls{det} curves and pick the best threshold per subgroup for a certain \gls{fpr}.
-
-We evaluated \gls{tar} for specific \gls{far} values.
-As described in Section~\ref{subsec:pf}, the verification experiments were 5-fold, with no overlap in subject ID between folds.
-Results reported are averaged across folds in all cases and are shown in Table~\ref{tab:ethnicy-far}.
-For each subgroup, the \gls{tar} of using a global threshold is reported (upper row), as well as using the optimal per subgroup threshold (lower row).
-
-Even for lower \gls{far}, there are notable improvements, often of the order of 1\%, which can be challenging to achieve when \gls{far} is near $\geq$90\%.
-More importantly, each subgroup has the desired \gls{fpr}, so that substantial differences in \gls{fpr} will remain unfounded.
-We experimented with ethnicity estimators on both the query and test image, which yielded similar results to those reported here.
+%  for it better adapts wh specific to subgroup) shift in sensitivity across subgroups when deep encodings are compared, the underlying mechanism
 
@@
 	\centering    
-	\includegraphics[width=.85\linewidth] {figures/human_eval.pdf}
-		\caption{\small{\textbf{Qualitative results of human assessment.} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}.
+	\includegraphics[width=.55\linewidth] {figures/human_eval.pdf}
+		\caption{\small{\textbf{Human assessment (qualitative).} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}.
 Accuracy scores shown as bar plots.
 Humans are most successful at recognizing their own subgroup, with a few exceptions (\eg bottom).}}
-		% (\eg bottom row). } Result summary in Table~\ref{tab:human-eval}.
 		\label{fig:human-eval} 
-% 		\vspace{-1mm}
 \end{figure}
-% \vspace{-4mm}
-% \vspace{-5pt}
-
-\subsection{Human evaluation}
- Quantitative and qualitative results are in Table~\ref{tab:humsn-eval-results} and Fig.~\ref{fig:human-eval}, respectively.
-Subjects of a subgroup likely have mostly been exposed to others of the same.
-Therefore, it is expected they would be best at labeling their own; similar for the same ethnicity, but another gender.
-Our findings concur.
-Each subgroup is best at labeling their type, and then second best at labeling the same ethnicity but opposite sex.
-Interestingly, each group of images is best tagged by the corresponding subgroup, with the second-best having the same ethnicity and opposite gender.
-On average, subgroups are comparable at labeling images.
-Analogous to the \gls{fr} system, performance ratings differ when analyzing within and across subgroups.
-In other words, performance on \gls{bfw} improved with subgroup-specific thresholds.
-Similarly, humans tend to better recognize individuals by facial cues of the same or similar demographics.
-Put differently, as the recognition performances drop with a global threshold optimized for one subgroup and deployed on another, human performance tends to drop when across subgroups (\ie most familiar with the same subgroup; performance seems to drop when faced with another, less familiar subgroup).
-
-%  for it better adapts wh specific to subgroup) shift in sensitivity across subgroups when deep encodings are compared, the underlying mechanism
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{table}[t!]
-\begin{center}
-    \caption{\small{\textbf{Quantitative of human assessment.} Different subgroups listed per row and column.
-Note that each subject is best within their subgroup, and second-best within the same subgroup but different gender.
-CF shows the least variation, but with the lowest accuracy.
-CM shows the best accuracy, but second to \gls{wm} in deviation from the mean.
-Thus, scores for males vary more than females.}}
-    \label{tab:humsn-eval-results} 
-    %  \vspace{-2mm}
-\footnotesize
-\scalebox{0.8}{
-\begin{tabular}{l c c c c c}
-        &  CF  &      CM     &   WF     &   WM     &  Avg\\\midrule
-        CF &  \textbf{\textcolor{blue}{52.9}}&   \textbf{\textcolor{red}{48.0}}&43.8&44.7 &47.4$\pm$4.1 \\
-        CM &  \textbf{\textcolor{red}{45.6}} & \textbf{\textcolor{blue}{0.504}}  & 44.4 &36.2 &44.1$\pm$5.9 \\
-        WF & 44.7 &43.8 & \textbf{\textcolor{blue}{0.573}}& \textbf{\textcolor{red}{48.0}} & 48.5$\pm$6.2 \\
-        WM & 30.1&\textbf{\textcolor{red}{47.4}} &  45.3 & \textbf{\textcolor{blue}{56.1}} & 44.7$\pm$10.8 \\\midrule
-        Avg &  43.3& 47.4&47.7 &46.3 &46.2$\pm$2.0\\
- \end{tabular}}
- \end{center}
-%  \vspace{-7mm}
-\end{table}
 
@@
 % \begin{bibunit}
-\input{supplemental.tex}
+% \input{supplemental.tex}
 % \end{bibunit}

### 2020-04-13 sections/analysis.tex
moved back into section (ie organized by file)

@@
-\begin{figure}[t!]
-	\centering    
-	\includegraphics[width=\linewidth]{figures/confusion.pdf}
-		\caption{\small{\textbf{Confusion matrix.} Percent error (Rank 1) for all faces of \gls{bfw} queried against all others.
-Notice errors concentrate within a subgroup, consistent with the \gls{sdm} in Fig.~\ref{fig:detection-model} (\ie \gls{af} show worst performance, and is mostly confused with faces of the same demographic).
-This plot is evidence that while race/ethnicity may be challenging to define, the subgroups are meaningful.}}
-		\label{fig:confusion} 
-\end{figure}
-
 \section{Results and Analysis}
@@
 For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https://github.com/clcarwin/sphereface\_pytorch}} As reported in~\cite{wang2019racial}, \gls{lfw} has about a 13\%, 14\%, 3\%, and 70\% ratio in Asians, Blacks, Indians, and Whites, respectively; furthermore, CASIA-Web is even more unbalanced (again, as reported in~\cite{wang2019racial}), with about  3\%, 11\%, 2\%, and 85\% for the same subgroups.
+\vspace{-4mm}
+\glsunset{det}
+% \subsection{\gls{det} analysis}
+\noindent\paragraph{\gls{det} analysis.}
+\gls{det} curves (5-fold, averaged) show per-subgroup trade-offs (Fig.~\ref{fig:detcurves}).
+Note that \gls{m} performs better than \gls{f}, precisely as one would expect from the tails of score-distributions for \emph{genuine} pairs (Fig.~\ref{fig:detection-model}).
+\Gls{af} and \gls{if} perform the worst.
+\glsunset{m}
+\glsunset{f}
 
-Additional \gls{cnn}-based models demonstrate the same phenomena: proportional to the overall model performance, exact in which the ordering subgroups in sensitivity in scores space (Appendix~\ref{app:sec:other:models}).
 % Specifically, we analyze three \glspl{cnn} variants: we encode faces using VGG-16~\cite{simonyan2014very},  50-layer residual network (\ie ResNet-50)~\cite{he2016deep}, and \gls{senet} (\ie SENet-50)~\cite{hu2018squeeze}, with results of only the latter (\ie the best performing \gls{senet}) model used throughout the main paper.
 Results for the other two, alongside \gls{senet} for comparison, are in the Supplemental Material.
+\vspace{-4mm}
 
 \glsunset{am}\glsunset{af}\glsunset{bm}\glsunset{bf}\glsunset{im}\glsunset{if}\glsunset{wm}\glsunset{wf}
-
-\subsection{Score analysis}
+\begin{figure*}[h!]
+    \centering
+    \begin{subfigure}[t]{.3\linewidth}
+    \includegraphics[width=.71\linewidth]{figures/curve_vgg16_subgroups-crop.pdf}
+    \caption{VGG16~\cite{simonyan2014very}}
+ \end{subfigure}
+    \begin{subfigure}[t]{.27\linewidth}
+    \includegraphics[width=.75\linewidth]{figures/curve_resnet50_subgroups-crop.pdf}
+    \caption{ResNet50~\cite{he2016deep}}
+   \end{subfigure}
+    \begin{subfigure}[t]{.27\linewidth}
+    \includegraphics[width=.75\linewidth]{figures/curve_senet50_subgroups-crop.pdf}
+    \caption{SENet~\cite{hu2018squeeze}}
+    \end{subfigure}
+    \caption{\small{\textbf{\gls{det} curves for different CNNs}.
+\gls{fnr} (\%) (vertical) vs \gls{fpr}  (horizontal, log-scale) for VGG2~\cite{Cao18} models with different backbones (VGG16, Resnet50, SEnet50).
+Lower is better.
+For each plot, \gls{wm} is the top-performer, while \gls{af} is the worst.
+The the ordering of the curves is roughly the same for each backbone.}}\label{fig:sdm-appendix-a}
+    \vspace{-5mm}
+\end{figure*}
+\noindent\paragraph{Score analysis.}
+% \subsection{Score analysis}
 Fig.~\ref{fig:detection-model} shows score distributions for faces of the same (\ie \emph{Genuine}) and different (\ie \emph{Imposter}) identity, with a subgroup per \gls{sdm} plot.
-Notice that score distributions for imposters tends to peak about zero for all subgroups, and with minimal deviation comparing modes of the different plots.
+Notice that score distributions for imposters tend to peak about zero for all subgroups, and with minimal deviation comparing modes of the different plots.
 On the other hand, the score distribution of the \emph{genuine} pairs varies across subgroups in location (\ie score value) and spread (\ie overall shape).
@@
 Fig.~\ref{fig:confusion} shows the confusion matrix of the subgroups.
-A vast majority of errors occurs in intra-subgroup.
+A vast majority of errors occur in the intra-subgroup.
 It is interesting to note that while the definition of  each group  based on ethnicity and race may not be crisply defined, the confusion matrix indicates that in practice the \gls{cnn} finds that the groups are effectively separate.
-The categories are, therefore, meaningful in the context of \gls{fr}.
+The categories are, therefore, meaningful in for \gls{fr}.
 
-\subsection{\gls{det} analysis}
-\glsunset{m}
-\glsunset{f}
-
-\gls{det} curves averaged across 5-folds show per-subgroup trade-offs (Fig.~\ref{fig:detcurves}).
-Note that \gls{m} perform better than \gls{f}, precisely as one would expect from the tails of score-distributions for \emph{genuine} pairs (Fig.~\ref{fig:detection-model}).
+% \gls{det} curves averaged across 5-folds show per-subgroup trade-offs (Fig.~\ref{fig:detcurves}).
+Note that \gls{m} performs better than \gls{f}, precisely as one would expect from the tails of score-distributions for \emph{genuine} pairs (Fig.~\ref{fig:detection-model}).
 \Gls{af} and \gls{if} perform the worst.
 
-\begin{table}[t!]
+\begin{table}[b!]
+    \vspace{-4mm}
 \glsunset{tar}
@@
 \caption{\small{\textbf{\gls{tar} at intervals of \gls{far}}.
-For each subgroup, and overall average \gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom).
-Higher is better.
-The proposed shows improvement in all cases.}}\label{tab:ethnicy-far}
-%  \vspace{-2mm}
+gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom).
+Higher is better.}}\label{tab:ethnicy-far}
+ \vspace{-2mm}
 \begin{center}
+\tiny
 \begin{tabular}{l c c c c c}
@@
 \end{table}
+\glsunset{fpr}
 
-The gender-based \gls{det} curve shows a difference in performances between \gls{m} and \gls{f} with a fixed threshold (dashed-line).
-Similar effects exist for the other curves as well (lines omitted to declutter).
-For many \gls{fr} applications, systems are operated at the highest \gls{fpr} allowed, so the line of constant threshold indicates that a single threshold produces different operating points (\ie \gls{fpr}), which is undesirable.
-The difference in \gls{fpr} is approximately a factor of two-- quite large indeed.
-If this is the case in an industrial system, one would expect a difference in about double the false positives to be reported on based on subgroup alone.
-The potential ramifications of such a bias should not be overlooked, which it has not as of lately-- gaining interest of even main-stream media ~\cite{england2019,snow2018}.
+The gender-based \gls{det} curves show performances with a fixed threshold (dashed line).
+Other curves relate similarity (lines omitted to declutter).
+For many \gls{fr} applications, systems operate at the highest \gls{fpr} allowed.
+The constant threshold shows that a single threshold produces different operating points (\ie \gls{fpr}) across subgroups, which is undesirable.
+% The difference in \gls{fpr} is approximately a factor of two.
+If this is the case in an industrial system, one would expect a difference in about double the FPs reported based on subgroup alone.
+The potential ramifications of such a bias should be considered, which it has not as of lately-- even noticed in main-stream media ~\cite{england2019,snow2018}.
 
-\subsection{Verification threshold} \label{subsec:analysis:verification}
+To further demonstrate the extend of the problem, we follow settings typical for systems in practice: we set the desired \gls{fpr}, and then determine the percent difference, \ie desired versus actual (Fig.~\ref{fig:percent:difference}, \emph{top}).
+Furthermore, we mitigate this highly skewed phenomena by applying subgroup-specific thresholds (\emph{bottom}).
+By this, minimal error from the desired \gls{fpr}.
+In addition, where there is small error, the offset is balanced across subgroups.
+
+% \subsubsection{Other \gls{cnn} models.}~\label{app:sec:other:models}
+\vspace{-5mm}
+\noindent\paragraph{Model analysis.}
+Variations in optimal threshold exist across models (Fig.~\ref{fig:sdm-appendix-a}).
+Like in Fig.~\ref{fig:detcurves}, the \gls{det} curves for three \gls{cnn}-based models, each trained on VGG2 with softmax but with different backbones.\footnote{\href{https://github.com/rcmalli/keras-vggface}{https://github.com/rcmalli/keras-vggface}} Notice similar trends across subgroups and models, which is consistent with  Sphereface as well (Fig.~\ref{fig:detcurves}).
+For example, the plots generated with Spherface and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst).
+Thus, the additional \gls{cnn}-based models demonstrate the same phenomena: proportional to the overall model performance, exact in the ordering of curves for each subgroup.
+
+\vspace{-5mm}
+\noindent\paragraph{Verification threshold analysis.}%\subsection{Verification threshold} \label{subsec:analysis:verification}
 We seek to reduce the bias between subgroups.
@@
 To accomplish that, we used a per subgroup threshold.
-In \gls{fv}, we consider one image as the query and all others as test.
+In \gls{fv}, we consider one image as the query and all others as the test.
 For this, the ethnicity of the query image is assumed.
@@
 Results reported are averaged across folds in all cases and are shown in Table~\ref{tab:ethnicy-far}.
-For each subgroup, the \gls{tar} of using a global-threshold is reported (upper row), as well as using the optimal per subgroup threshold (lower row).
+For each subgroup, the \gls{tar} of using a global threshold is reported (upper row), as well as using the optimal per subgroup threshold (lower row).
 
@@
 More importantly, each subgroup has the desired \gls{fpr}, so that substantial differences in \gls{fpr} will remain unfounded.
-We further experimented using ethnicity estimators and using the ethnicity of both query and test image, which yielded similar results to those reported here.
+We experimented with ethnicity estimators on both the query and test image, which yielded similar results to those reported here.
 
-\begin{figure}[t!] 
-	\centering    
-	\includegraphics[width=\linewidth] {figures/human_eval.pdf}
-		\caption{\small{\textbf{Qualitative results of human assessment.} $\checkmark$ for \emph{match}; $\times$ for \emph{non-match}.
-Accuracy scores shown as bar plots.
-Humans are most successful at recognizing their own subgroup, with a few exceptions (\eg bottom).}}
-		% (\eg bottom row). } Result summary in Table~\ref{tab:human-eval}.
-		\label{fig:human-eval} 
-% 		\vspace{-1mm}
-\end{figure} 
-% \vspace{-4mm}
-% \vspace{-5pt}
-\subsection{Human evaluation}
- Quantitative and qualitative results are in Table~\ref{tab:humsn-eval-results} and Fig.~\ref{fig:human-eval}, respectfully.
-Subjects of a subgroup likely have mostly been exposed to others of the same.
-Therefore, it is expected they would be best at labeling their own; similar for the same ethnicity, but other gender.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{table}[t!]
+\begin{center}
+    \caption{\small{\textbf{Human assessment (quantitative).} Subgroups listed per row (\ie human) and column (\ie image).
+Note, most do the best intra-subgroup (\textcolor{blue}{blue}), and second-best (\textcolor{red}{red}) intra-subgroup but inter-gender.
+WF performs the best; WF pairs are most correctly matched.}}
+    % CF shows the least variation, but with the lowest accuracy. WF shows the best accuracy, in human performance and on face pairs of that subgroup.}}
+    \label{tab:humsn-eval-results} 
+    % \vspace{-5mm}
+     \vspace{-2mm}
+\footnotesize
+\scalebox{0.94}{
+\begin{tabular}{c}
+\begin{tabular}{c l c c c cc}
+&&\multicolumn{4}{c}{\textbf{Image}}\\
+   \multicolumn{2}{c}{(Acc, \%)}& CF  & CM & WF &WM& Avg\\
+\end{tabular}\\
+\begin{tabular}{c l| r r r r| r}
+\cline{3-7}
+       &CF &\  \textbf{\textcolor{blue}{52.9}}&   \textbf{\textcolor{red}{48.0}}&43.8&44.7 &47.4 \\
+
+        \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{\textbf{Human}}}}  \hspace{-5mm}
+
+        %  \multirow{\items}{*}{\rotatebox{90}{\textbf{Human}}} \hspace{-5mm}
+        &CM &  \textbf{\textcolor{red}{45.6}} & \textbf{\textcolor{blue}{50.4}}  & 44.4 &36.2 &44.1 \\
+
+        &WF & 44.7 &43.8 & \textbf{\textcolor{blue}{57.3}}& \textbf{\textcolor{red}{48.0}} & \textbf{48.5} \\
+        &WM & 30.1&\textbf{\textcolor{red}{47.4}} &  45.3 & \textbf{\textcolor{blue}{56.1}} & 44.7\\\cline{3-7}
+        &Avg &  43.3& 47.4&\textbf{47.7} &46.3 &46.2\\
+ \end{tabular}
+ \end{tabular}
+ }
+ \end{center}
+ \vspace{-7mm}
+\end{table}
+
+\vspace{-5mm}
+\noindent\paragraph{Human evaluation analysis.}
+% \subsection{Human evaluation}
+Subjects of a subgroup likely have mostly been exposed to others of the same (Table~\ref{tab:humsn-eval-results} and Fig.~\ref{fig:human-eval}).
+Therefore, it is expected they would be best at labeling their own; similar for the same ethnicity, but another gender.
 Our findings concur.
 Each subgroup is best at labeling their type, and then second best at labeling the same ethnicity but opposite sex.
-Interestingly, each group of images is best tagged by the corresponding subgroup, with the second-to-best having the same ethnicity and opposite gender.
+Interestingly, each group of images is best tagged by the corresponding subgroup, with the second-best having the same ethnicity and opposite gender.
 On average, subgroups are comparable at labeling images.
@@
 Similarly, humans tend to better recognize individuals by facial cues of the same or similar demographics.
-Put differently, as the recognition performances drops with a global threshold optimized for one subgroup and deployed on another, human performance tends to drop when across subgroups (\ie most familiar with the same subgroup; performance seems to drop when faced with another, less familiar subgroup).
-
-%  for it better adapts wh specific to subgroup) shift in sensitivity across subgroups when deep encodings are compared, the underlying mechanism
+Put differently, as the recognition performances drop with a global threshold optimized for one subgroup and deployed on another, human performance tends to drop when across subgroups (\ie performances drop for less familiar subgroups).

### 2020-04-13 sections/introduction.tex
Version for arXiv
added cvpr template and started filling in
moved back into section (ie organized by file)

--- 2020-02-06 sections/introduction.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \glsresetall
 \section{Introduction}
-As more of society becomes integrated with \gls{ml}, bias, fairness, and the formalization of \gls{ml} standards are topics of high interest ~\cite{10.1007/978-3-030-13469-3_68, anne2018women, wang2018racial}.
-One effect of the growing dependency on technology is the increasing concern regarding biased and unfair algorithms.
-For instance, \gls{fr} systems can be untrustworthy and racist in some cases~\cite{england2019,snow2018}.
-.
-
-\begin{figure}[t!]
-\centering
-         \includegraphics[trim=0in 0.2in 0in 0in,clip,width=.9\linewidth]{images/montage.pdf}  
-\caption{\textbf{\gls{bfw}.} The average face of its different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender.
-The others represent the ethnicity and gender of the race and gender, respectfully.
-Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}
-\label{fig:avg-faces}
-% \vspace{-3pt}
-\end{figure}
+    As more of society becomes integrated with \gls{ml}, bias, fairness, and the formalization of \gls{ml} standards attract more attention~\cite{10.1007/978-3-030-13469-3_68, anne2018women, wang2018racial}.
+Thus, an effect of the growing dependency on \gls{ml} are growing concerns for biased and unfair algorithms, \eg untrustworthy and racist \gls{fr} systems~\cite{england2019,snow2018}.
 
     Typically, \glspl{cnn} are trained on faces identified by a detection system.
-Specifically, for \gls{fr}, the goal is to project faces to an N-dimensional space with minimal distances between samples of the same identity and maximize the gap separating those that are different.
-Thus, the overarching goal is to discriminate between subjects with face encodings.
-We can deploy such a \gls{cnn} to encode faces to compare via a similarity score: genuine pair if the score is high enough; and rejected as an imposter.
+Specifically, for \gls{fr}, the goal is to encode faces in an N-dimensional space such that samples of the same identity are neighbors and those of different identity of further apart.
+Thus, the goal is to identify subjects with face encodings.
+We can deploy a \gls{cnn} to encode faces, and then compare via a similarity score: assume pairs with a high enough score are \emph{genuine}; else, an \emph{imposter}.
 
-A fixed threshold acts as a decision boundary in similarity score space.
-Thus, the face features in the same class must satisfy a criterion in the form of a single value~\cite{deng2019arcface, liu2017sphereface, wang2018additive, wang2018cosface}.
-However, a single, global threshold is a crude measure that leads to \gls{fr} errors.
-Furthermore, the held-out set used to determine the global threshold tends to share the same distribution with test data, which makes the results skewed in favor of specific demographics that make up a majority in both sets.
-That skew, the difference in the performance of an algorithm of certain ethnic groups, is our definition of bias.
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{figures/fig1-crop.pdf}
+    \caption{\small{\textbf{Depiction of biometrics.} The \gls{sdm} (\emph{top-left}) shows the sensitivity related to a single threshold $t_g$.
+Translated to error rate (\emph{bottom-left}), a direct trade-off between FNP and FPR.
+In practice, this is entirely application dependent.
+Specifically, the specific chose in desired FPR (\%) (\emph{bottom-right}).}}
+    \label{fig:biometrics}
+\end{figure}
+    % \begin{figure}[t!]
+    %     \centering
+    %     \includegraphics[width=.8\linewidth]{figures/montage.pdf}
+    %     \caption{\small{\textbf{\gls{bfw}.} Average face of the different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender.
+The others represent the ethnicity and gender of the race and gender, respectively.
+Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}}
+    %     \label{fig:avg-faces}
+    % \end{figure}
+
+    When comparing scores, typically, a fixed threshold sets the decision boundary (Fig.~\ref{fig:biometrics}).
+Thus, features of the same identity must satisfy a criterion via a single value~\cite{deng2019arcface, liu2017sphereface, wang2018additive, wang2018cosface}.
+However, we found that a single (\ie global) threshold is a crude measure that leads to skewed errors - the held-out set used to determine the threshold tends to share the same distribution with the test data, which favors specific demographics that are a majority.
+That skew, the difference in the performance of an algorithm of certain demographics, is our definition of bias.
 A key question is: \emph{is \gls{fr} too biased, or not?}
-
-\begin{table*}[t!]
-\scriptsize
-    \centering
-       \caption{\textbf{The proposed \gls{bfw} compared to related datasets.} \gls{bfw} is exactly balanced across ID, gender, and ethnicity (Table~\ref{tab:ethnic-splits}).
-Compared with \gls{dp}, \gls{bfw} provides more samples per subject and subgroups per set.
-Also, \gls{bfw} uses a single resource, VGG2.
-\gls{rfw}, on the other hand, supports a different task (\ie domain adaptation).
-Furthermore, \gls{rfw} focuses on race-distribution across of pairs, while not considering the distribution of identities.
-}
-    \begin{tabular}{rccccccc}%x{10mm}x{3mm}x{10mm}x{8mm}}%\toprule
-
-    \multicolumn{2}{c}{Database} & \multicolumn{3}{c}{Number of}& \multicolumn{3}{c}{Balanced Labels}\\
-    \cmidrule(lr){1-2}	\cmidrule(lr){3-5} \cmidrule(lr){6-8}
-    Name & Source Data & Faces &  IDs & Subgroups & ID & Ethnicity & Gender\\\midrule
-    \gls{dp}~\cite{demogPairs}     & CASIA-W, VGG\&VGG2 & 10,800& 600 & 6 &\checkc& \checkc &\checkc \\
-    \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &$\approx$80,000&$\approx$12,000& 4 & \xmark & \checkc &\xmark \\
-    \gls{bfw} (ours) & VGG2 & 20,000 & 800 &8 & \checkc & \checkc &\checkc \\\bottomrule
-    \end{tabular}
-    \label{tab:compared}
-    \vspace{-5mm}
-\end{table*}
-
-The adverse effects of a global threshold are two-fold: \textbf{(1)} the mapping produced by a \gls{cnn} is nonuniform.
-Therefore, distances between pairs of faces in different demographics vary in distribution of similarity scores (Fig~\ref{fig:detection-model}); \textbf{(2)} the evaluation set is imbalanced as well.
-Particular demographics making up a majority of the population will carry most weight on reported performance ratings.
-Reported results skew away from common traits to the underrepresented subgroups.
-Alas, demographics like gender, ethnicity, race, and age are underrepresented in most public datasets.
-
-Making matters more challenging is that race and ethnicity are loosely defined.
-For example, the US Census Bureau allows an individual to self-identify race (\href{https://www.census.gov/mso/www/training/pdf/race-ethnicity-onepager.pdf}{https://www.census.gov}).
-We define it as a group of people having facial characteristics similar to those found in a region.
-The result is various types of biases in FR systems in favor of or against particular demographics remain a question.
-
-To address \textbf{(2)}, the lack of a balanced data, we introduce a new benchmark for \gls{fr}, called \gls{bfw} (Table~\ref{tab:compared}, ~\ref{tab:ethnic-splits}).
-\gls{bfw} serves as a platform to fairly evaluate \gls{fr} systems and enable demographic-specific ratings to be reported.
-We use \gls{bfw} to gain a deeper understanding of the extent of bias present in facial embeddings extracted from a \gls{soa} \gls{cnn} model.
-We then suggest a mechanism to counter the biased feature space to mitigate problems of bias with more balanced performance ratings for different demographics, while improving the overall accuracy.
-Specifically, we propose using an adaptive threshold that varies depending on the characteristics of detected facial attributes (\ie gender and ethnicity, Fig~\ref{fig:avg-faces}).
-We show an increase in accuracy with a balanced performance for different subgroups of people.
-Similarly, we show the positive effect of adjusting the similarity threshold based on the facial features of matched faces.
-Thus, selective use of similarity thresholds in current \gls{soa} \gls{fr} systems provides more intuition in \gls{fr}-based research, while providing a method easily adoptable in practice.
 
@@
     \caption{\small{\textbf{Database stats and nomenclature, optimal thresholds ($t_o$), and accuracy scores.} \textit{Header:} Subgroup definitions.
-\textit{Top-row:} Statistics of \gls{bfw}.
-\textit{Middle-row:} Number of pairs for each partition.
-\textit{Bottom-row:} Accuracy when applying a global threshold $t_g$, the optimal threshold $t_o$, and accuracy with $t_o$ per subgroup.
+\textit{Top:} Statistics of \gls{bfw}.
+\textit{Bottom::} Number of pairs for each partition.
 Columns grouped by race and then further split by gender.
 Out of millions of pairs, accuracy is inconsistent across subgroups.
-Furthermore, $F$ tend to perform inferior to that of $M$ (\ie up to 8\%).}}\label{tab:ethnic-splits}
-    \scalebox{1}{
+Furthermore, $F$ tends to perform inferior to that of $M$ for all but B.}}\label{tab:ethnic-splits}
+    \scalebox{.75}{
      \resizebox{\textwidth}{!}{%
@@
         % \midrule
-        \specialrule{.01em}{.05em}{.05em}
-        Acc$@t_g$ & 0.876 & 0.944  &0.934 &0.942 &0.922&0.949 &0.916  &0.918&0.925$\pm$0.022 \\
-        $t_o$ & 0.235 &  0.274 & 0.267&0.254  &0.299 & 0.295& 0.242 &0.222&0.261\\ % row 4
-        Acc$@t_o$ & 0.916 &0.964 &0.955 &0.971 &0.933 &0.958 & 0.969 &0.973 & 0.955 $\pm$ 0.018\\
+        % \specialrule{.01em}{.05em}{.05em}
+        % Acc$@t_g$ & 0.941 & 0.952  & 0.967 & 0.960 & 0.958 &0.962  & 0.973 & 0.981 &0.962 $\pm$0.012 \\
+        % $t_o$ & 0.255 &  0.246 & 0.277 &0.242  &0.297 & 0.264& 0.233 &0.239 &0.257$\pm$0.006\\ % row 4
+
+        % Acc$@t_o$ & 0.942 &0.952 &0.968 &0.961 &0.960 &0.962 & 0.974 &0.982 & 0.963 $\pm$ 0.012\\
         \bottomrule
     \end{tabular}}
-    \glsunset{af}
+    % \glsunset{af}
     }
-    \glsunset{af}
-    \glsunset{am}
-    \glsunset{bf}
-    \glsunset{bm}
-    \glsunset{if}
-    \glsunset{im}
-    \glsunset{wf}
-    \glsunset{wm}
-    \vspace{-12pt}
+    % % \glsunset{af}
+    % \glsunset{am}
+    % \glsunset{bf}
+    % \glsunset{bm}
+    % \glsunset{if}
+    % \glsunset{im}
+    % \glsunset{wf}
+    % \glsunset{wm}
+    % \vspace{-12pt}
 \end{table*}
 
-The contributions in this paper are as follows:
-\begin{enumerate}
-    \item We built a balanced dataset as a proxy to measure verification performance per subgroup.
-    \item We analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup significantly boost and balances performances.
-    \item We showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
-    \item We conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
-\end{enumerate}
+    Making matters more challenging is that race and ethnicity are loosely defined.
+For example, the US Census Bureau allows an individual to self-identify race.\footnote{\scriptsize\href{https://www.census.gov/mso/www/training/pdf/race-ethnicity-onepager.pdf}{www.census.gov/mso/www/training/pdf/race-ethnicity-onepager}} For this work, we define subgroups as specific sub-populations with face characteristics similar to those found in a region.
+
+    The adverse effects of a global threshold are two-fold: \textbf{(1)} the mapping produced by a \gls{cnn} is nonuniform.
+Therefore, distances between pairs of faces in different demographics vary in distribution of similarity scores (Fig~\ref{fig:detection-model}); \textbf{(2)} the evaluation set is imbalanced as well.
+Particular demographics making up a majority of the population will carry most weight on the reported performance ratings.
+Reported results favor the common traits over the underrepresented.
+Alas, demographics like gender, ethnicity, race, and age are underrepresented in most public datasets~\cite{merler2019diversity, wang2018racial}.
+
+    % The result is various types of biases in FR systems in favor of or against particular demographics remain a question.
+
+    For \textbf{(1)} we propose subgroup-specific (\ie optimal) thresholds; to address \textbf{(2)}, we introduce a new benchmark for measuring bias in \gls{fr}, \gls{bfw} (Table~\ref{tab:compared} \&~\ref{tab:ethnic-splits}).
+\gls{bfw} allows for a fair evaluation of \gls{fr} systems, while enabling demographic-specific ratings to be reported.
+We use \gls{bfw} to gain a deeper understanding of the extent of bias present in \gls{soa} \gls{cnn} used to encode faces.
+Then, we suggest a mechanism to mitigate problems of bias with more balanced performance ratings for different demographics.
+Specifically, we propose using an adaptive threshold that varies depending on the characteristics of detected facial attributes (\ie gender and ethnicity, Fig~\ref{fig:avg-faces}).
+We show an increase in accuracy with a balanced performance for different subgroups.
+Similarly, we show a positive effect of adjusting the similarity threshold based on the facial features of matched faces.
+Thus, selective use of similarity thresholds in current \gls{soa} \gls{fr} systems provides more intuition in \gls{fr} research with a method easy to adopt in practice.
+
+    The contributions of this work are 3-fold.
+(1) We built a balanced dataset as a proxy to measure verification performance per subgroup for studies of bias in \gls{fr}.
+(2) We revealed an unwanted bias in scores of face pairs - a bias that causes ratings to skew across demographics.
+For this, we showed that an adaptive threshold per subgroup balances performance (\ie the typical use of a global threshold unfavorable, which we address via optimal thresholds).
+(3) We surveyed humans to demonstrate bias in human perception (NIH-certified, \textit{Protect Humans in Research}).%\footnote{NIH-certified, \textit{Protect Humans in Research}.}%19-09-08}.}
+    % \begin{enumerate}
+    %     \item Built a balanced dataset as a proxy to measure verification performance per subgroup.
+    %     \item Analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup significantly boost and balance performances.
+    %     \item Showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
+    %     \item Conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
+    % \end{enumerate}

### 2020-04-13 sections/relatedworks.tex
moved back into section (ie organized by file)

--- 2020-02-06 sections/relatedworks.tex
@@
     \subsection{Bias in \gls{ml}}
-Although the community is excited about the progress and the commercial value of \gls{ml}, the trust between the society and techniques will take time to build due to the bias in algorithms.
-The exact definitions and implications of bias tend to vary between sources, as do its sources and types.
-One common theme is the tendency of bias to hinder performance ratings in ways that skew in favor of a particular subpopulation.
-At a high-level, bias can be introduced by humans~\cite{windmann1998subconscious}, data and label types~\cite{tommasi2017deeper}, \gls{ml} models~\cite{amini2019uncovering, kim2019learning}, and during evaluation~\cite{stock2018convnets}.
-For instance, a vehicle-detection model might miss cars if training set was mostly trucks.
-In practice, many \gls{ml} systems learn on biased datasets, which could harm our society.
+        The progress and the commercial value of \gls{ml} is exciting; however, the trust of society in \gls{ml} will take time to build due to inherent biases.
+The exact definitions and implications of bias vary between sources, as do its sources and types.
+A common theme is that bias hinders performance ratings in ways that skew to a particular sub-population-- the source varies, whether from humans~\cite{windmann1998subconscious}, data and label types~\cite{tommasi2017deeper}, \gls{ml} models~\cite{amini2019uncovering, kim2019learning}, or evaluation protocols~\cite{stock2018convnets}.
+For instance, a vehicle-detection model might miss cars if training data was mostly trucks.
+In practice, many \gls{ml} systems learn on biased data, which could be detrimental for society.
 
     \subsection{Bias in \gls{fr}}
-Problems of bias in \gls{fr} have been driven by different motivations and have therefore been solved by different means.
-To name a few: problems of data augmentation~\cite{yin2019feature}, one-shot learning~\cite{ding2018one}, demographic parity and fairness with priority on privacy~\cite{huang2018generative}, domain adaptation~\cite{wang2018racial}, differences in face-based attributes across demographics~\cite{wang2018they}, and even exploratory data analysis~\cite{muthukumar2019}.
-Yin et al.
-\cite{yin2019feature} proposed to augment the feature space of underrepresented classes using other classes with a diverse collection of samples to encourage distributions of underrepresented classes to more closely resemble that of the others.
-Similarly,~\cite{ding2018one} formulated the imbalanced class as one-shot learning and trained a \gls{gan} to generate face features as a means of augmenting class with as few as one sample.
-\cite{huang2018generative} proposed \gls{gapf} to create fair representations of the data in a quantifiable way, allowing for the finding of a decorrelation scheme from the data without access to its statistics.
-Wang et al.
-\cite{wang2018they} defined subgroups at a finer-level (\ie Chinese, Japanese, Korean) to determine the familiarity of faces across these subgroups.
-Genders have also been used as subgroups for work in bias, whether for efforts of analysis and understanding of gender-based face encodings~\cite{muthukumar2019}.
-Most recently,~\cite{wang2018racial} proposed adapting domains to bridge the gap between races by knowledge transfer, which was supported by a novel labeled data collection, \gls{rfw}.
-The release of \gls{rfw} came after the completion of \gls{bfw}.
-Although similar in terms of demographics (\ie Asian, Black, Indian, and White), \gls{rfw} uses faces from MSCeleb~\cite{guo2016ms} based on the assumption CASIA-Face~\cite{yi2014learning} and VGG2~\cite{Cao18} was used for training.
-In contrast, our \gls{bfw} samples from VGG2 and assumes MSCeleb was the training set.
-As shown in \cite{wang2018racial}, MSCeleb is highly imbalanced, primarily consisting of images of white individuals.
-For this, we expect bias from data, for this is the training set.
-Furthermore, our experiments allow for the parsing of results based on gender demographics (\ie although \gls{rfw} does separate males and females in each subgroup, no consideration for splits in just gender or just races are incorporated into the experimental design).
-Thus, \gls{rfw} and \gls{bfw} are complementary, with both adding metadata on demographics for subjects with faces in renowned, large-scale \gls{fr} datasets.
+        Problems of bias in \gls{fr} have been driven by different motivations; therefore, solutions have been posed for different matters.
+To name a few: data augmentation~\cite{yin2019feature}, one-shot learning~\cite{ding2018one}, demographic parity and fairness with priority on privacy~\cite{huang2018generative}, domain adaptation~\cite{wang2018racial}, differences in face-based attributes across demographics~\cite{wang2018they}, and even data exploration~\cite{muthukumar2019}.
+Yin~\etal proposed to augment the feature space of underrepresented classes using other classes with a diverse collection of samples to encourage distributions of underrepresented classes to more closely resemble that of the others~\cite{yin2019feature}.
+Similarly, others formulated the imbalanced class problem as one-shot learning, where a \gls{gan} was trained to generate face features to augment classes with fewer samples~\cite{ding2018one}.
+\gls{gapf} was proposed to create fair representations of the data in a quantifiable way, allowing for the finding of a de-correlation scheme from the data without access to its statistics~\cite{huang2018generative}.
+Wang~\etal defined subgroups at a finer-level (\ie Chinese, Japanese, Korean), and determined the familiarity of faces inter-subgroup~\cite{wang2018they}.
+Genders have also been used to make subgroups (\eg for analysis of gender-based face encodings~\cite{muthukumar2019}).
+Most recently,~\cite{wang2018racial} proposed adapting domains to bridge the bias gap by knowledge transfer, which was supported by a novel data collection, \gls{rfw}.
+The release of \gls{rfw} occurred after \gls{bfw} was built - although similar in terms of demographics, \gls{rfw} uses faces from MSCeleb~\cite{guo2016ms} for testing, and assumes CASIA-Face~\cite{yi2014learning} and VGG2~\cite{Cao18} were used to train.
+In contrast, our \gls{bfw} assumes VGG2 as the test set.
+        % As shown in \cite{wang2018racial}, MSCeleb is highly imbalanced, primarily consisting of images of white individuals. 
+        % For this, we expect bias from data, for this is the training set. 
+        Furthermore, \gls{bfw} balances gender and race: \gls{rfw} splits subgroups by gender and race, while \gls{bfw} has gender, race, or both). 
+        % Thus, \gls{rfw} and \gls{bfw} complement one another. % - both add demographic for subjects with faces in renowned, large-scale \gls{fr} datasets.
 
-Most similar to our work are the recent efforts in~\cite{das2018, demogPairs, lopez2019dataset, srinivas2019face}.
-A common factor of these efforts are claims of insufficient data supply to support studies on bias in \gls{fr}, and the introduction of new metadata on demographics for image sets parsed out of existing collections.
-Specifically, \cite{demogPairs} curated a set of faces based on racial demographics (\ie Asian, Black, and White) called \gls{dp}, while~\cite{srinivas2019face} honed in on adults versus children called gls{itwcc}.
+Most similar to us is~\cite{das2018, demogPairs, lopez2019dataset, srinivas2019face} - each were motivated by insufficient paired data for studying bias in \gls{fr}; then, problems were addressed using labeled data from existing image collections.
+Specifically, Hupont~\etal curated a set of faces based on racial demographics (\ie \gls{a}, \gls{b}, and \gls{w}) called \gls{dp}~\cite{demogPairs}, while~\cite{srinivas2019face} honed in on adults versus children called \gls{itwcc}.
 Like the proposed \gls{bfw}, both were built by sampling existing databases, but with the addition of tags for the respective subgroups of interest.
@@
 
+    \begin{table*}[t!]
+
+        \centering
+        \caption{\small{\textbf{\gls{bfw} compared to related datasets.} Our \gls{bfw} is balanced across ID, gender, and ethnicity (Table~\ref{tab:ethnic-splits}).
+Compared with \gls{dp}, \gls{bfw} provides more samples per subject and subgroups per set, while only using a single resource, VGG2.
+\gls{rfw}, on the other hand, supports a different task (\ie domain adaptation), and focuses on race-distribution, but not the distribution of identities.}}
+        \scriptsize
+        \begin{tabular}{rccccccc}%x{10mm}x{3mm}x{10mm}x{8mm}}%\toprule
+
+            \multicolumn{2}{c}{Database} & \multicolumn{3}{c}{Number of}& \multicolumn{3}{c}{Balanced Labels}\\
+            \cmidrule(lr){1-2}	\cmidrule(lr){3-5} \cmidrule(lr){6-8}
+            Name & Source Data & Faces &  IDs & Subgroups & ID & Ethnicity & Gender\\\midrule
+            \gls{dp}~\cite{demogPairs}     & CASIA-W~\cite{yi2014learning}, VGG~\cite{schroff2015facenet} \&VGG2~\cite{Cao18} & 10,800& 600 & 6 &\checkc& \checkc &\checkc \\
+            \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &$\approx$80,000&$\approx$12,000& 4 & \xmark & \checkc &\xmark \\
+            \gls{bfw} (ours) & VGG2 & 20,000 & 800 &8 & \checkc & \checkc &\checkc \\\bottomrule
+        \end{tabular}
+        \label{tab:compared}
+            \vspace{-5mm}
+    \end{table*}
+
 \subsection{Human bias in \gls{ml}}
-Bias is not unique to \gls{ml}, as humans also susceptible to a perceived bias across demographics: it exists across different races, genders, and even ages~\cite{10.1007/978-3-030-13469-3_68, bar2006, meissner2001, nicholls2018}.
-\cite{wang2018they} showed machines surpass human performance in classifying faces as Japanese, Chinese, or Korean by nearly 150\%.
-Precisely, humans barely pass random with
-38.89\% accuracy (\ie $\frac{1}{3}$ is random).
+Bias is not unique to \gls{ml} - humans are also susceptible to a perceived bias.
+In fact, biases exist across race, gender, and even age~\cite{10.1007/978-3-030-13469-3_68, bar2006, meissner2001, nicholls2018}.
+Wang~\etal showed machines surpass human performance in discriminating between Japanese, Chinese, or Korean faces by nearly 150\%~\cite{wang2018they}, as humans just pass random (\ie 38.89\% accuracy).
 
-We expect human bias to skew results in favor of their genders and races.
-For this, we measure the human perception of faces across demographics in a controlled experiment.
-In the end, the results concur~\cite{wang2018they}-- we too observed an overall average below random (\ie $<$50\%).
-Furthermore, we provide details on settings and counts on the number of submissions per demographics while analyzing per demographic.
+We expect human bias to skew to their own genders and races.
+For this, we measure the human perception with face pairs of different subgroups (Section~\ref{subsec:human-assessment}).
+The results concur with~\cite{wang2018they}, as we also recorded overall averages below random ($<$50\%).
+%The details on settings and number of submissions per demographics are discussed in Section~\ref{}.

### 2020-04-13 sections/section3.tex
moved back into section (ie organized by file)

@@
-\glsunset{if}\glsunset{im}\glsunset{af}\glsunset{am}\glsunset{bf}\glsunset{bm}\glsunset{wf}\glsunset{wm}
 \section{The BFW Benchmark and Dataset}
 
-We now discuss the \gls{bfw} dataset, and protocols to evaluate \gls{ml}-based \gls{fr}.
-We  conclude this section by reviewing the human evaluation conducted as part of this work, to detect bias in humans.
+We now discuss the \gls{bfw} dataset (\ie collection and preparation), along with the evaluation protocols. We conclude this section by reviewing the settings followed to survey bias in humans.
 
 \subsection{The data}
-Problems of bias in \gls{fr} motivated the design of \gls{bfw}.
-The data evenly represent various subgroups partitioned by demographics.
-Inspired by \gls{dp}~\cite{demogPairs}, the specification of \gls{bfw} follows in suit.
-\gls{bfw} includes additional subgroups (\ie \gls{if} and \gls{im}), an increase in the number subjects per subgroup, and with many more pairs (Table~\ref{tab:ethnic-splits}).
+Problems of bias in \gls{fr} motivated us to build \gls{bfw}.
+The data evenly represent various subgroups partitioned by demographic.
+Inspired by \gls{dp}~\cite{demogPairs}, the specification of \gls{bfw} follows in suit, but with additional subgroups (\ie \gls{if} and \gls{im}), an increase in the number of subjects per subgroup, and many more pairs (Table~\ref{tab:ethnic-splits}).
 
@@
 \noindent\textbf{Compiling subject list.} 
- All subjects were sampled from VGG2~\cite{Cao18}.
-Thus, unlike others that depend on multiple sources, \gls{bfw} has less potential conflicts in train/test overlap for existing models.
+Subjects were sampled from VGG2~\cite{Cao18} - unlike others built from multiple sources, \gls{bfw} has fewer potential conflicts in train and test overlap with existing models.
 We used pre-trained ethnicity~\cite{ambekar2009name} and gender~\cite{levi2015age} classifiers to find candidates for the different subgroups.
@@
 \vspace{1mm}
-\noindent\textbf{Detecting faces.} Faces were detected using \gls{mtcnn}~\cite{zhang2016joint}.\footnote{\href{https://github.com/polarisZhao/mtcnn-pytorch}{https://github.com/polarisZhao/mtcnn-pytorch}} Then, assigned into one of two sets.
+\noindent\textbf{Detecting faces.} Faces were detected using MTCNN~\cite{zhang2016joint}.\footnote{\href{https://github.com/polarisZhao/mtcnn-pytorch}{https://github.com/polarisZhao/mtcnn-pytorch}} Then, assigned into one of two sets.
 Faces within detected bounding box (BB) regions extended out 130\% in each direction, with zero-padding as the boundary condition made-up one set.
@@
 
-\begin{figure}[!t] 
-	\centering    
- \glsunset{fpr}
-  \glsunset{fnr}
-	\includegraphics[width=\linewidth]{figures/detcurve-improved.pdf}
-		\caption{\textbf{\gls{det} curves.} \emph{Top-left}: per gender.
-\emph{Top-right}: per ethnicity.
-\emph{Bottom}: per subgroup (\ie combined).
-A dashed lined shows a difference by about 2$\times$ in \gls{fpr} for the same threshold $\theta_{const}$.
-\gls{fnr} is the number of match errors, so closer a curve is to the bottom better.}
-		\glsreset{det}\glsreset{fpr}
-\label{fig:detcurves} 
-\end{figure}
-
 \vspace{1mm}
 \noindent\textbf{Validating labels.} 
-Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Sphereface~\cite{liu2017sphereface}: applying an affine transformation to align faces according to pre-defined eye locations, each face fed through twice (\ie the original and horizontally flipped), with the final encoding obtained by fusing the two features by average pooling (\ie 512 D).
-A matrix of cosine similarity scores was then generated for each subject, and removed samples (\ie rows) with median scores below threshold $\theta=0.2$ (set manually).
+Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Sphereface~\cite{liu2017sphereface}: applying an affine transformation to align faces according to predefined eye locations, each face fed through twice (\ie the original and horizontally flipped), with the final encoding obtained by fusing the two features by average pooling (\ie 512 D).
+A matrix of cosine similarity scores was then generated for each subject and removed samples (\ie rows) with median scores below threshold $\theta=0.2$ (set manually).
 Mathematically, the $n^{th}$ sample for the $j^{th}$ subject with $N_j$ faces was removed if the ordinal rank of its score $n = \frac{P\times N}{100}\geq\theta$, where $P=50$.
@@
 
+\begin{figure}[!t] 
+	\centering    
+ \glsunset{fpr}
+  \glsunset{fnr}
+	\includegraphics[width=.9\linewidth]{figures/detcurve-improved.pdf}
+		\caption{\small{\textbf{\gls{det} curves.} \emph{Top-left}: per gender.
+\emph{Top-right}: per ethnicity.
+\emph{Bottom}: per subgroup (\ie combined).
+Dashed line shows about 2$\times$ difference in \gls{fpr} for the same threshold $\theta_{const}$.
+\gls{fnr} is the match error count (closer to the bottom is better).}}
+		\glsreset{det}\glsreset{fpr}
+\label{fig:detcurves} 
+\end{figure} 
 \vspace{1mm}
 \noindent\textbf{Sampling faces and creating folds.} We created lists of pairs in five-folds with subjects split evenly per subject and without overlap across folds.
-Furthermore, balance in the number of faces per was obtained by sampling twenty-five faces at random from each.
-Next, we generated a list of all face pairs per subject, resulting in $\sum_{l=1}^{L}\sum_{k=1}^{K_d} {N_k \choose 2}$ positive pairs, where the number of faces of all $K_l$ subjects $N_k=25$  for each of the $L$ subgroups (Table~\ref{tab:ethnic-splits}).
+Furthermore, a balance in the number of faces per was obtained by sampling twenty-five faces at random from each.
+Next, we generated a list of all the face pairs per subject, resulting in $\sum_{l=1}^{L}\sum_{k=1}^{K_d} {N_k \choose 2}$ positive pairs, where the number of faces of all $K_l$ subjects $N_k=25$  for each of the $L$ subgroups (Table~\ref{tab:ethnic-splits}).
 Next, we assigned subjects to a fold.
@@
 Later, a negative set from samples within the same subgroup randomly matched until the count met that of the positive.
-Finally, we doubled the total count with negative pairs from across subgroups but in the same fold.
+Finally, we doubled the number with negative pairs from across subgroups but in the same fold.
+
 \begin{figure}[t!] 
 	\glsunset{fpr}
+% 	\glsunset{sdm}
 	\centering
@@
 	\includegraphics[width=1\linewidth]{figures/violinplots.pdf}
-		\caption{\textbf{\Gls{sdm} across subgroups.} The scores for \emph{imposters} have medians about 0.3 but with variation in upper percentiles, while \emph{genuine} pairs vary in both (\eg \gls{af} has more area of overlap).
-A varying threshold varying across different subgroups yields a constant \gls{fpr}.} \label{fig:detection-model}
-
+		\caption{\small{\textbf{\Gls{sdm} across subgroups.} Scores of \emph{imposters} have medians around 0.3 but with variations in upper percentiles; \emph{genuine} pairs vary in mean and spread (\eg \gls{af} has more area of overlap).
+A threshold varying across different subgroups yields a constant \gls{fpr}.}} \label{fig:detection-model}
+% 		\glsreset{sdm}
 \end{figure}
@@
 \Gls{fv} is the special case of the two-class (\ie boolean) classification.
-Hence, pairs are labeled as the ``same'' or ``different'' \textit{genuine} pairs (\ie \textit{match}) or \textit{imposter} (\ie \textit{mismatch}), respectfully.
+Hence, pairs are labeled as the ``same'' or ``different'' \textit{genuine} pairs (\ie \textit{match}) or \textit{imposter} (\ie \textit{mismatch}), respectively.
 This formulation (\ie \gls{fv}) is highly practical for applications like access control, re-identification, and surveillance.
-Typically, training a separate model for each unique subjects is infeasible.
+Typically, training a separate model for each unique subject is infeasible.
 Firstly, the computational costs compound as the number of subjects increase.
@@
 
+\begin{figure}[t!]
+	\centering    
+	\includegraphics[width=.75\linewidth]{figures/confusion.pdf}
+		\caption{\small{\textbf{Confusion matrix.} Error (Rank 1, \%) for all faces of \gls{bfw} vs.
+all others.
+Error concentrates intra-subgroup - consistent with the \gls{sdm} (Fig.~\ref{fig:detection-model}), as \gls{af} performs the worst, mostly confused by \gls{af}.
+Although race/ethnicity are challenging to define, this shows the subgroups are meaningful for \gls{fr}.}}
+		\label{fig:confusion} 
+		    \vspace{-6mm}
+\end{figure}
+
 \subsection{Human assessment}\label{subsec:human-assessment}
-We evaluated human on face pairs focusing on two racial groups: Chinese and Caucasians.
-To focus on the experiment, we honed-in on two groups, white Americans (W) and Chinese from China (C).
-The purpose was to the minimize variability by only analyzing the subsets of the broader groups of whites and Asians.
+To focus on the human evaluation experiment, we honed-in on pairs from two groups, white Americans (W) and Chinese from China (C).
+This minimized variability by only analyzing the subsets of the broader groups of whites and Asians.
+%We evaluated human on face pairs focusing on two subgroups: Chinese and Caucasians.
 
@@
 The other 10 pairs are of others (\eg Hispanic/ Latino, Japanese, African).
-Survey was created, distributed, and recorded via \href{https://paperform.co}{PaperForm}.
+The survey was created, distributed, and recorded via \href{https://paperform.co}{PaperForm}.

### 2020-04-20 sections/analysis.tex
updated bar plot (Fig 6)
cleanup; removed -vspaces
wip

@@
 A single \gls{cnn} was used as a means to control the experiments.
-For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https://github.com/clcarwin/sphereface\_pytorch}} As reported in~\cite{wang2019racial}, \gls{lfw} has about a 13\%, 14\%, 3\%, and 70\% ratio in Asians, Blacks, Indians, and Whites, respectively; furthermore, CASIA-Web is even more unbalanced (again, as reported in~\cite{wang2019racial}), with about  3\%, 11\%, 2\%, and 85\% for the same subgroups.
-\vspace{-4mm}
+For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https://github.com/clcarwin/sphereface\_pytorch}} As reported in~\cite{wang2019racial}, \gls{lfw} has about a 13\%, 14\%, 3\%, and 70\% ratio in Asian, Black, Indian, and White, respectively.
+Furthermore, CASIA-Web is even more unbalanced (again, as reported in~\cite{wang2019racial}), with about  3\%, 11\%, 2\%, and 85\% for the same subgroups.
+%\vspace{-4mm}
 \glsunset{det}
@@
 Results for the other two, alongside \gls{senet} for comparison, are in the Supplemental Material.
-\vspace{-4mm}
+%\vspace{-4mm}
 
@@
 For each plot, \gls{wm} is the top-performer, while \gls{af} is the worst.
-The the ordering of the curves is roughly the same for each backbone.}}\label{fig:sdm-appendix-a}
-    \vspace{-5mm}
+The ordering of the curves is roughly the same for each backbone.}}\label{fig:sdm-appendix-a}
+%    \vspace{-5mm}
 \end{figure*}
@@
 A vast majority of errors occur in the intra-subgroup.
-It is interesting to note that while the definition of  each group  based on ethnicity and race may not be crisply defined, the confusion matrix indicates that in practice the \gls{cnn} finds that the groups are effectively separate.
-The categories are, therefore, meaningful in for \gls{fr}.
+It is interesting to note that while the definition of  each group  based on ethnicity and race may not be crisply defined, the confusion matrix indicates that in practice, the \gls{cnn} finds that the groups are effectively separate.
+The categories are, therefore, meaningful for \gls{fr}.
 
@@
 \begin{table}[b!]
-    \vspace{-4mm}
+%    \vspace{-4mm}
 \glsunset{tar}
@@
 \caption{\small{\textbf{\gls{tar} at intervals of \gls{far}}.
-gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom).
+\gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom).
 Higher is better.}}\label{tab:ethnicy-far}
- \vspace{-2mm}
+% \vspace{-2mm}
 \begin{center}
-\tiny
+\scriptsize
 \begin{tabular}{l c c c c c}
@@
 
-To further demonstrate the extend of the problem, we follow settings typical for systems in practice: we set the desired \gls{fpr}, and then determine the percent difference, \ie desired versus actual (Fig.~\ref{fig:percent:difference}, \emph{top}).
+To further demonstrate the extent of the problem, we follow settings typical for systems in practice.
+We set the desired \gls{fpr}, and then determine the percent difference, \ie desired versus actual (Fig.~\ref{fig:percent:difference}, \emph{top}).
 Furthermore, we mitigate this highly skewed phenomena by applying subgroup-specific thresholds (\emph{bottom}).
@@
 % \subsubsection{Other \gls{cnn} models.}~\label{app:sec:other:models}
-\vspace{-5mm}
+%\vspace{-5mm}
 \noindent\paragraph{Model analysis.}
@@
 Like in Fig.~\ref{fig:detcurves}, the \gls{det} curves for three \gls{cnn}-based models, each trained on VGG2 with softmax but with different backbones.\footnote{\href{https://github.com/rcmalli/keras-vggface}{https://github.com/rcmalli/keras-vggface}} Notice similar trends across subgroups and models, which is consistent with  Sphereface as well (Fig.~\ref{fig:detcurves}).
-For example, the plots generated with Spherface and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst).
+For example, the plots generated with Sphereface and VggFace2 all have the \gls{wm} curve at the bottom (\ie best) and \gls{af} on top (\ie worst).
 Thus, the additional \gls{cnn}-based models demonstrate the same phenomena: proportional to the overall model performance, exact in the ordering of curves for each subgroup.
 
-\vspace{-5mm}
+%\vspace{-5mm}
 \noindent\paragraph{Verification threshold analysis.}%\subsection{Verification threshold} \label{subsec:analysis:verification}
@@
 To accomplish that, we used a per subgroup threshold.
-In \gls{fv}, we consider one image as the query and all others as the test.
+In \gls{fv}, we consider one image as the query, and all others as the test.
 For this, the ethnicity of the query image is assumed.
@@
  \end{center}
- \vspace{-7mm}
+% \vspace{-7mm}
 \end{table}
 
-\vspace{-5mm}
+%\vspace{-5mm}
 \noindent\paragraph{Human evaluation analysis.}

### 2020-04-20 sections/introduction.tex
wip

@@
 \section{Introduction}
-    As more of society becomes integrated with \gls{ml}, bias, fairness, and the formalization of \gls{ml} standards attract more attention~\cite{10.1007/978-3-030-13469-3_68, anne2018women, wang2018racial}.
-Thus, an effect of the growing dependency on \gls{ml} are growing concerns for biased and unfair algorithms, \eg untrustworthy and racist \gls{fr} systems~\cite{england2019,snow2018}.
+    As more of society becomes integrated with \gls{ml}, topics related to bias, fairness, and the formalization of \gls{ml} standards attract more attention~\cite{10.1007/978-3-030-13469-3_68, anne2018women, wang2018racial}.
+Thus, an effect of the growing dependency on \gls{ml} is a growing concern for biased and unfair algorithms, \eg untrustworthy and racist \gls{fr} systems~\cite{england2019, nagpal2019deep,snow2018}.
 
-    Typically, \glspl{cnn} are trained on faces identified by a detection system.
-Specifically, for \gls{fr}, the goal is to encode faces in an N-dimensional space such that samples of the same identity are neighbors and those of different identity of further apart.
-Thus, the goal is to identify subjects with face encodings.
-We can deploy a \gls{cnn} to encode faces, and then compare via a similarity score: assume pairs with a high enough score are \emph{genuine}; else, an \emph{imposter}.
+    Nowadays, \glspl{cnn} are trained on faces identified by a detection system.
+Specifically, for \gls{fr}, the goal is to encode faces in an N-dimensional space such that samples of the same identity are neighbors, while different identities are farther apart.
+Thus, we can deploy a \gls{cnn} to encode faces, and then compare via a similarity score (\ie pairs with a high enough score classified as \emph{genuine}; else, \emph{imposter}).
 
@@
     \includegraphics[width=\linewidth]{figures/fig1-crop.pdf}
-    \caption{\small{\textbf{Depiction of biometrics.} The \gls{sdm} (\emph{top-left}) shows the sensitivity related to a single threshold $t_g$.
+    \caption{\small{\textbf{Depiction of bio-metrics.} The \gls{sdm} (\emph{top-left}) shows the sensitivity related to a single threshold $t_g$.
 Translated to error rate (\emph{bottom-left}), a direct trade-off between FNP and FPR.
@@
 \end{figure}
-    % \begin{figure}[t!]
-    %     \centering
-    %     \includegraphics[width=.8\linewidth]{figures/montage.pdf}
-    %     \caption{\small{\textbf{\gls{bfw}.} Average face of the different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender.
-The others represent the ethnicity and gender of the race and gender, respectively.
-Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}}
-    %     \label{fig:avg-faces}
-    % \end{figure}
 
-    When comparing scores, typically, a fixed threshold sets the decision boundary (Fig.~\ref{fig:biometrics}).
-Thus, features of the same identity must satisfy a criterion via a single value~\cite{deng2019arcface, liu2017sphereface, wang2018additive, wang2018cosface}.
-However, we found that a single (\ie global) threshold is a crude measure that leads to skewed errors - the held-out set used to determine the threshold tends to share the same distribution with the test data, which favors specific demographics that are a majority.
-That skew, the difference in the performance of an algorithm of certain demographics, is our definition of bias.
+Typically, a fixed threshold sets the decision boundary by which to compare scores (Fig.~\ref{fig:bio-metrics}).
+Thus, features of the same identity must satisfy a criterion based on a single value~\cite{deng2019arcface, liu2017sphereface, wang2018additive, wang2018cosface}.
+However, we found that a single (\ie global) threshold is a crude measure that leads to skewed errors.
+Furthermore, the held-out set used to determine the threshold tends to share the same distribution with the test set, favoring specific demographics that are the majority.
+That skew (\ie the difference in the performance of an algorithm of certain demographics) is our definition of bias.
 A key question is: \emph{is \gls{fr} too biased, or not?}
@@
     \end{tabular}}
-    % \glsunset{af}
     }
-    % % \glsunset{af}
-    % \glsunset{am}
-    % \glsunset{bf}
-    % \glsunset{bm}
-    % \glsunset{if}
-    % \glsunset{im}
-    % \glsunset{wf}
-    % \glsunset{wm}
     % \vspace{-12pt}
@@
 
-    % The result is various types of biases in FR systems in favor of or against particular demographics remain a question.
-
-    For \textbf{(1)} we propose subgroup-specific (\ie optimal) thresholds; to address \textbf{(2)}, we introduce a new benchmark for measuring bias in \gls{fr}, \gls{bfw} (Table~\ref{tab:compared} \&~\ref{tab:ethnic-splits}).
+    For \textbf{(1)} we propose subgroup-specific (\ie optimal) thresholds, while we address \textbf{(2)} with a new benchmark dataset for measuring bias in \gls{fr}, \gls{bfw} (Table~\ref{tab:ethnic-splits} and~\ref{tab:compared}).
 \gls{bfw} allows for a fair evaluation of \gls{fr} systems, while enabling demographic-specific ratings to be reported.
-We use \gls{bfw} to gain a deeper understanding of the extent of bias present in \gls{soa} \gls{cnn} used to encode faces.
+We use \gls{bfw} to gain a deeper understanding of the extent to which bias presents in \gls{soa} \gls{cnn} used to encode faces.
 Then, we suggest a mechanism to mitigate problems of bias with more balanced performance ratings for different demographics.
@@
 (3) We surveyed humans to demonstrate bias in human perception (NIH-certified, \textit{Protect Humans in Research}).%\footnote{NIH-certified, \textit{Protect Humans in Research}.}%19-09-08}.}
-    % \begin{enumerate}
-    %     \item Built a balanced dataset as a proxy to measure verification performance per subgroup.
-    %     \item Analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup significantly boost and balance performances.
-    %     \item Showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
-    %     \item Conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
-    % \end{enumerate}

### 2020-04-20 sections/section3.tex
updated bar plot (Fig 6)
wip

@@
 
-We now discuss the \gls{bfw} dataset (\ie collection and preparation), along with the evaluation protocols. We conclude this section by reviewing the settings followed to survey bias in humans.
+We now discuss the \gls{bfw} dataset (\ie collection and preparation), along with the evaluation protocols. We conclude this section by reviewing the settings used to survey bias in humans.
 
@@
 Problems of bias in \gls{fr} motivated us to build \gls{bfw}.
-The data evenly represent various subgroups partitioned by demographic.
-Inspired by \gls{dp}~\cite{demogPairs}, the specification of \gls{bfw} follows in suit, but with additional subgroups (\ie \gls{if} and \gls{im}), an increase in the number of subjects per subgroup, and many more pairs (Table~\ref{tab:ethnic-splits}).
+Inspired by \gls{dp}~\cite{demogPairs}, the data is made up of evenly split subgroups, but with an increase in subgroups (\ie \gls{if} and \gls{im}), subjects per subgroup, and face pairs (Table~\ref{tab:ethnic-splits}).
 
@@
 Subjects were sampled from VGG2~\cite{Cao18} - unlike others built from multiple sources, \gls{bfw} has fewer potential conflicts in train and test overlap with existing models.
-We used pre-trained ethnicity~\cite{ambekar2009name} and gender~\cite{levi2015age} classifiers to find candidates for the different subgroups.
+To find candidates for the different subgroups, we first parsed the list of names using a pre-trained ethnicity model~\cite{ambekar2009name}.
+This was then further refined by processing faces using ethnicity~\cite{fu2014learning} and gender~\cite{levi2015age} classifiers.
+This resulted in hundreds of candidates per subgroup, which allowed us to manually filter 100 subjects per the 8 subgroups.
 
 \vspace{1mm}
-\noindent\textbf{Detecting faces.} Faces were detected using MTCNN~\cite{zhang2016joint}.\footnote{\href{https://github.com/polarisZhao/mtcnn-pytorch}{https://github.com/polarisZhao/mtcnn-pytorch}} Then, assigned into one of two sets.
+\noindent\textbf{Detecting faces.} Faces were detected using MTCNN~\cite{zhang2016joint}.\footnote{\href{https://github.com/polarisZhao/mtcnn-pytorch}{https://github.com/polarisZhao/mtcnn-pytorch}} Then, faces were assigned to one of two sets.
 Faces within detected bounding box (BB) regions extended out 130\% in each direction, with zero-padding as the boundary condition made-up one set.
@@
 \noindent\textbf{Validating labels.} 
-Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Sphereface~\cite{liu2017sphereface}: applying an affine transformation to align faces according to predefined eye locations, each face fed through twice (\ie the original and horizontally flipped), with the final encoding obtained by fusing the two features by average pooling (\ie 512 D).
+Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Sphereface~\cite{liu2017sphereface}.
+For this, each face was aligned to predefined eye locations via an affine transformation.
+Then, faces were fed through the \gls{cnn} twice (\ie the original and horizontally flipped), with two features fused by average pooling (\ie 512 D).
 A matrix of cosine similarity scores was then generated for each subject and removed samples (\ie rows) with median scores below threshold $\theta=0.2$ (set manually).
 Mathematically, the $n^{th}$ sample for the $j^{th}$ subject with $N_j$ faces was removed if the ordinal rank of its score $n = \frac{P\times N}{100}\geq\theta$, where $P=50$.
-In other words, the median (\ie 50 percentile) of all scores for a faces with respect to all of faces for the respective subject must pass a threshold of $\theta=0.2$; else, the face is dropped.
+In other words, the median (\ie 50 percentile) of all scores for a face with respect to all of faces for the respective subject must pass a threshold of $\theta=0.2$; else, the face is dropped.
 This allowed us to quickly prune \gls{fp} face detections.
@@
 \end{figure} 
-\vspace{1mm}
+%\vspace{1mm}
 \noindent\textbf{Sampling faces and creating folds.} We created lists of pairs in five-folds with subjects split evenly per subject and without overlap across folds.
-Furthermore, a balance in the number of faces per was obtained by sampling twenty-five faces at random from each.
+Furthermore, a balance in the number of faces per subgroup was obtained by sampling twenty-five faces at random from each.
 Next, we generated a list of all the face pairs per subject, resulting in $\sum_{l=1}^{L}\sum_{k=1}^{K_d} {N_k \choose 2}$ positive pairs, where the number of faces of all $K_l$ subjects $N_k=25$  for each of the $L$ subgroups (Table~\ref{tab:ethnic-splits}).
@@
 		\label{fig:confusion} 
-		    \vspace{-6mm}
+%		    \vspace{-6mm}
 \end{figure}
@@
 \subsection{Human assessment}\label{subsec:human-assessment}
-To focus on the human evaluation experiment, we honed-in on pairs from two groups, white Americans (W) and Chinese from China (C).
-This minimized variability by only analyzing the subsets of the broader groups of whites and Asians.
+To focus on the human evaluation experiment, we honed-in on pairs from two groups, White Americans (W) and Chinese from China (C).
+This minimized variability compared to the broader groups of Whites and Asians, which was thought to be best, provided only a small subset of the data was used on fewer humans than subjects in \gls{bfw}.
+
 %We evaluated human on face pairs focusing on two subgroups: Chinese and Caucasians.
 
-Samples were collected by recruiting subjects from multiple sources (\eg social media, email lists, and family/friends)-- a total of 120 participants were sampled at random from all the submissions that were (1) complete and (2) from a W or C participant.
+Samples were collected from individuals recruited from multiple sources (\eg social media, email lists, and family/friends) - a total of 120 participants were sampled at random from all the submissions that were (1) complete and (2) done by a W or C participant.
 Specifically, there were 60 W and 60 C, both with \gls{m} and \gls{f} split evenly.
-A total of 50 face pairs of non-famous ``look-alikes'' were collected from the internet, with 20 ({\emph WA}) and 20 ({\emph C}) pairs (male and female split evenly).
-The other 10 pairs are of others (\eg Hispanic/ Latino, Japanese, African).
+A total of 50 face pairs of non-famous ``look-alikes'' were collected from the internet, with 20 ({\emph WA}) and 20 ({\emph C}) pairs with, again, \gls{m} and \gls{f} split evenly.
+The other 10 pairs were of a different demographic (\eg Hispanic/ Latino, Japanese, African).
 The survey was created, distributed, and recorded via \href{https://paperform.co}{PaperForm}.
+It is important to note that participants were only aware of the task (\ie to verify whether or not a face-pair was a \emph{match} or \emph{non-match}, but with no knowledge of it being a study on bias).

### 2020-04-20 cvprworkshop.tex

@@
 
-% \author{First Author\\
-% Institution1\\
-% Institution1 address\\
-% {\tt\small firstauthor@i1.org}
-% % For a paper whose authors are all at the same institution,
-% % omit the following lines up until the closing ``}''.
-% % Additional authors and addresses can be added with ``\and'',
-% % just like the second author.
-% % To save space, use either the email address or home page, not both
-% \and
-% Second Author\\
-% Institution2\\
-% First line of institution2 address\\
-% {\tt\small secondauthor@i2.org}
-% }
-
 \maketitle
-%\thispagestyle{empty}
 
@@
 We show variations in the optimal scoring threshold for face-pairs across different subgroups.
-Thus, the conventional approach of learning a  global threshold for all pairs resulting in performance gaps among subgroups.
-By learning subgroup-specific thresholds, we not only mitigate problems in performance gaps but also show a notable boost in the overall performance.
-Furthermore, we do a human evaluation to measure the bias in humans, which supports the hypothesis that such a bias exists in human perception.
+Thus, the conventional approach of learning a  global threshold for all pairs results in performance gaps between subgroups.
+By learning subgroup-specific thresholds, we reduce performance gaps, and also show a notable boost in overall performance.
+Furthermore, we do a human evaluation to measure bias in humans, which supports the hypothesis that an analogous bias exists in human perception.
 For the \gls{bfw} database, source code, and more, visit \href{https://github.com/visionjo/facerec-bias-bfw}{https://github.com/visionjo/facerec-bias-bfw}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \begin{bibunit}
 \input{sections/introduction}
@@
        \centering
-    \includegraphics[width=.95\linewidth]{figures/adaptive_fpr_percent_diff_complete.pdf}
+    \includegraphics[width=.95\linewidth]{figures/fpr_percent_diff-crop.pdf}
     \caption{\small{\textbf{Percent difference from intended \gls{fpr}.} \emph{Top:} $t_g$ yields \gls{fpr} the span as large as 2$\times$ (\ie 200\%) that intended (\ie \gls{wm} for 1e-4).
@@
 Once established that the results do suffer from problems of bias, we observed that the same threshold across ethnic and gender subgroups leads to differences in the \gls{fpr} up to a factor of two.
-Furthermore, the percent difference from in which is seemingly the cause of the frenzy about bias in FR in main-stream media.
-Furthermore, we ameliorate these differences with a per-subgroup threshold, leveling out FPR, and achieving a higher TPR.
-We hypothesized that most humans grew amongst more than their own demographic and, therefore, effectively learn from imbalanced datasets-- a human evaluation validated that humans are biased, as most recognized their personal demographic best.
+Also, the percent difference that seemingly the cause of the frenzy about bias in FR in main-stream media.
+Furthermore, we ameliorate these differences with a per-subgroup threshold, leveling out \gls{fpr}, and achieving a higher \gls{tpr}.
+We hypothesized that most humans grew among more than their own demographic and, therefore, effectively learn from imbalanced datasets-- a human evaluation validated that humans are biased, as most recognized their personal demographic best.
 The focused research findings presented here, along with the public database included, are extendable in vast ways.
@@
 
-{\small
+{\scriptsize
 \bibliographystyle{ieee_fullname}

### 2020-04-20 sections/relatedworks.tex
cleanup; removed -vspaces

@@
     \subsection{Bias in \gls{ml}}
-        The progress and the commercial value of \gls{ml} is exciting; however, the trust of society in \gls{ml} will take time to build due to inherent biases.
+        The progress and commercial value of \gls{ml} is exciting; however, the trust of society in \gls{ml} will take time to build due to inherent biases.
 The exact definitions and implications of bias vary between sources, as do its sources and types.
@@
         Problems of bias in \gls{fr} have been driven by different motivations; therefore, solutions have been posed for different matters.
-To name a few: data augmentation~\cite{yin2019feature}, one-shot learning~\cite{ding2018one}, demographic parity and fairness with priority on privacy~\cite{huang2018generative}, domain adaptation~\cite{wang2018racial}, differences in face-based attributes across demographics~\cite{wang2018they}, and even data exploration~\cite{muthukumar2019}.
+Instances can be in problems of data augmentation~\cite{yin2019feature}, one-shot learning~\cite{ding2018one}, demographic parity and fairness with priority on privacy~\cite{huang2018generative}, domain adaptation~\cite{wang2018racial}, differences in face-based attributes across demographics~\cite{wang2018they}, and even data exploration~\cite{muthukumar2019}.
 Yin~\etal proposed to augment the feature space of underrepresented classes using other classes with a diverse collection of samples to encourage distributions of underrepresented classes to more closely resemble that of the others~\cite{yin2019feature}.
@@
 Genders have also been used to make subgroups (\eg for analysis of gender-based face encodings~\cite{muthukumar2019}).
-Most recently,~\cite{wang2018racial} proposed adapting domains to bridge the bias gap by knowledge transfer, which was supported by a novel data collection, \gls{rfw}.
+Most recently,~\cite{wang2018racial} proposed to adapt domains to bridge the bias gap by knowledge transfer, which was supported by a novel data collection, \gls{rfw}.
 The release of \gls{rfw} occurred after \gls{bfw} was built - although similar in terms of demographics, \gls{rfw} uses faces from MSCeleb~\cite{guo2016ms} for testing, and assumes CASIA-Face~\cite{yi2014learning} and VGG2~\cite{Cao18} were used to train.
@@
 
-Most similar to us is~\cite{das2018, demogPairs, lopez2019dataset, srinivas2019face} - each were motivated by insufficient paired data for studying bias in \gls{fr}; then, problems were addressed using labeled data from existing image collections.
+Most similar to us is~\cite{das2018, demogPairs, lopez2019dataset, srinivas2019face} - each were motivated by insufficient paired data for studying bias in \gls{fr}.
+Then, problems were addressed using labeled data from existing image collections.
 Specifically, Hupont~\etal curated a set of faces based on racial demographics (\ie \gls{a}, \gls{b}, and \gls{w}) called \gls{dp}~\cite{demogPairs}, while~\cite{srinivas2019face} honed in on adults versus children called \gls{itwcc}.
 Like the proposed \gls{bfw}, both were built by sampling existing databases, but with the addition of tags for the respective subgroups of interest.
-Besides, the additional data of \gls{bfw} (\ie added an additional subgroup \gls{i}, along with additional subjects with more faces for all subgroups), we also further split subgroups by gender.
+Aside from the additional data of \gls{bfw} (\ie added an additional subgroup \gls{i}, along with additional subjects with more faces for all subgroups), we also further split subgroups by gender.
 Furthermore, we focus on the problem of facial verification and the different levels of sensitivity in cosine similarity scores per subgroup.
@@
         \label{tab:compared}
-            \vspace{-5mm}
+%            \vspace{-5mm}
     \end{table*}

### 2020-07-09 cvprworkshop.tex
camera-ready

@@
 \usepackage{pifont}% http://ctan.org/pkg/pifont
+\usepackage{balance}
 
@@
     \includegraphics[width=.8\linewidth]{figures/montage.pdf}
-    \caption{\small{\textbf{\gls{bfw}.} Average face of the different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender.
-The others represent the ethnicity and gender of the race and gender, respectively.
+    \caption{\small{\textbf{\gls{bfw}.} Average face per subgroup: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per ethnicity;  \emph{left-column}: per gender.
+The others represent the ethnicity and gender, respectively.
 Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}}
@@
 \emph{Bottom:} Subgroup-specific thresholds reduces this difference to near zero, where there are small differences, the percent difference across different subgroups is fair (\ie FPR=1e-4).}}\label{fig:percent:difference}
+    \vspace{-4mm}
 \end{figure}
@@
 		\label{fig:human-eval} 
+		 \vspace{-5mm}
 \end{figure}
 
-\glsresetall
+\glsreset{bfw}
 \section{Conclusion}
-We introduce a new data set \gls{bfw} with eight subgroups balanced across gender and ethnicity.
+We introduce the \gls{bfw} dataset with eight subgroups balanced across gender and ethnicity.
 With this, and upon highlighting the challenges and shortcomings of grouping subjects as a single subset, we provide evidence that forming subgroups is meaningful, as the FR algorithm rarely makes mistakes across subgroups.
@@
 Once established that the results do suffer from problems of bias, we observed that the same threshold across ethnic and gender subgroups leads to differences in the \gls{fpr} up to a factor of two.
-Also, the percent difference that seemingly the cause of the frenzy about bias in FR in main-stream media.
+Also, we clearly showed notable percent differences in ratings across subgroups.
 Furthermore, we ameliorate these differences with a per-subgroup threshold, leveling out \gls{fpr}, and achieving a higher \gls{tpr}.
-We hypothesized that most humans grew among more than their own demographic and, therefore, effectively learn from imbalanced datasets-- a human evaluation validated that humans are biased, as most recognized their personal demographic best.
-The focused research findings presented here, along with the public database included, are extendable in vast ways.
-Thus, we see this as the slither to a much larger problem of bias in ML.
+We hypothesized that most humans grew among more than their own demographic and, therefore, effectively learn from imbalanced datasets.
+In essence, a human evaluation validated that humans are biased, as most recognized their personal demographic best.
+This research, along with the data and resources, are extendable in vast ways.
+Thus, this is just a sliver of the larger problem of bias in ML.
 
-{\scriptsize
+{\small
+\balance
 \bibliographystyle{ieee_fullname}

### 2020-07-09 sections/analysis.tex
camera-ready

@@
 A single \gls{cnn} was used as a means to control the experiments.
-For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https://github.com/clcarwin/sphereface\_pytorch}} As reported in~\cite{wang2019racial}, \gls{lfw} has about a 13\%, 14\%, 3\%, and 70\% ratio in Asian, Black, Indian, and White, respectively.
-Furthermore, CASIA-Web is even more unbalanced (again, as reported in~\cite{wang2019racial}), with about  3\%, 11\%, 2\%, and 85\% for the same subgroups.
+For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https://github.com/clcarwin/sphereface\_pytorch}} As reported in~\cite{wang2019racial}, \gls{lfw} has about 13\%, 14\%, 3\%, and 70\% ratio in Asian, Black, Indian, and White, respectively.
+Furthermore, CASIA-Web is even more unbalanced (again, as reported in~\cite{wang2018racial}), with about  3\%, 11\%, 2\%, and 85\% for the same subgroups.
 %\vspace{-4mm}
@@
 % \subsection{\gls{det} analysis}
+\vspace{-5mm}
 \noindent\paragraph{\gls{det} analysis.}
@@
 The ordering of the curves is roughly the same for each backbone.}}\label{fig:sdm-appendix-a}
-%    \vspace{-5mm}
+    \vspace{-5mm}
 \end{figure*}
+
+\vspace{-5mm}
 \noindent\paragraph{Score analysis.}
 % \subsection{Score analysis}
-Fig.~\ref{fig:detection-model} shows score distributions for faces of the same (\ie \emph{Genuine}) and different (\ie \emph{Imposter}) identity, with a subgroup per \gls{sdm} plot.
+Fig.~\ref{fig:detection-model} shows score distributions for faces of the same (\ie \emph{genuine}) and different (\ie \emph{imposter}) identity, with a subgroup per \gls{sdm} graph.
 Notice that score distributions for imposters tend to peak about zero for all subgroups, and with minimal deviation comparing modes of the different plots.
 On the other hand, the score distribution of the \emph{genuine} pairs varies across subgroups in location (\ie score value) and spread (\ie overall shape).
-Asian Fem
 Fig.~\ref{fig:confusion} shows the confusion matrix of the subgroups.
@@
 \begin{table}[b!]
-%    \vspace{-4mm}
+    \vspace{-4mm}
 \glsunset{tar}
@@
 Higher is better.}}\label{tab:ethnicy-far}
-% \vspace{-2mm}
+\vspace{-2mm}
 \begin{center}
@@
 \glsreset{far}
-%  \vspace{-3mm}
+ \vspace{-3mm}
 
@@
 We set the desired \gls{fpr}, and then determine the percent difference, \ie desired versus actual (Fig.~\ref{fig:percent:difference}, \emph{top}).
-Furthermore, we mitigate this highly skewed phenomena by applying subgroup-specific thresholds (\emph{bottom}).
-By this, minimal error from the desired \gls{fpr}.
-In addition, where there is small error, the offset is balanced across subgroups.
+Furthermore, we mitigate this highly skewed phenomenon by applying subgroup-specific thresholds (\emph{bottom}) - by this, minimal error from the desired \gls{fpr}.
+Besides where there is a small error, the offset is balanced across subgroups.
 
 % \subsubsection{Other \gls{cnn} models.}~\label{app:sec:other:models}
-%\vspace{-5mm}
+\vspace{-5mm}
 \noindent\paragraph{Model analysis.}
@@
 
-%\vspace{-5mm}
+\vspace{-5mm}
 \noindent\paragraph{Verification threshold analysis.}%\subsection{Verification threshold} \label{subsec:analysis:verification}
@@
  \end{center}
-% \vspace{-7mm}
+\vspace{-5mm}
 \end{table}
 
-%\vspace{-5mm}
+\vspace{-5mm}
 \noindent\paragraph{Human evaluation analysis.}
@@
 Subjects of a subgroup likely have mostly been exposed to others of the same (Table~\ref{tab:humsn-eval-results} and Fig.~\ref{fig:human-eval}).
-Therefore, it is expected they would be best at labeling their own; similar for the same ethnicity, but another gender.
+Therefore, it is expected they would be best at labeling their own, similar to the same ethnicity, but another gender.
 Our findings concur.
 Each subgroup is best at labeling their type, and then second best at labeling the same ethnicity but opposite sex.
+
 Interestingly, each group of images is best tagged by the corresponding subgroup, with the second-best having the same ethnicity and opposite gender.
@@
 Similarly, humans tend to better recognize individuals by facial cues of the same or similar demographics.
-Put differently, as the recognition performances drop with a global threshold optimized for one subgroup and deployed on another, human performance tends to drop when across subgroups (\ie performances drop for less familiar subgroups).
+Put differently, as the recognition performances drop with a global threshold optimized for one subgroup and deployed on another, human performance tends to fall when across subgroups (\ie performances drop for less familiar subgroups).

### 2020-07-09 sections/introduction.tex
camera-ready

@@
     As more of society becomes integrated with \gls{ml}, topics related to bias, fairness, and the formalization of \gls{ml} standards attract more attention~\cite{10.1007/978-3-030-13469-3_68, anne2018women, wang2018racial}.
-Thus, an effect of the growing dependency on \gls{ml} is a growing concern for biased and unfair algorithms, \eg untrustworthy and racist \gls{fr} systems~\cite{england2019, nagpal2019deep,snow2018}.
+Thus, an effect of the growing dependency on \gls{ml} is an ever-increasing concern about the biased and unfair algorithms, \eg untrustworthy and prejudiced \gls{fr}~\cite{nagpal2019deep,snow2018}.
 
     Nowadays, \glspl{cnn} are trained on faces identified by a detection system.
-Specifically, for \gls{fr}, the goal is to encode faces in an N-dimensional space such that samples of the same identity are neighbors, while different identities are farther apart.
+Specifically, for \gls{fr}, the goal is to encode faces in an N-dimensional space such that samples of the same identity are neighbors, while different persons are farther apart.
 Thus, we can deploy a \gls{cnn} to encode faces, and then compare via a similarity score (\ie pairs with a high enough score classified as \emph{genuine}; else, \emph{imposter}).
@@
 
-Typically, a fixed threshold sets the decision boundary by which to compare scores (Fig.~\ref{fig:bio-metrics}).
-Thus, features of the same identity must satisfy a criterion based on a single value~\cite{deng2019arcface, liu2017sphereface, wang2018additive, wang2018cosface}.
-However, we found that a single (\ie global) threshold is a crude measure that leads to skewed errors.
+Typically, a fixed threshold sets the decision boundary by which to compare scores (Fig.~\ref{fig:biometrics}).
+As such, features of the same identity must satisfy a criterion based on a single value~\cite{deng2019arcface, liu2017sphereface, wang2018additive, wang2018cosface}.
+However, we found that an individual (\ie global) threshold is a crude measure that leads to skewed errors.
 Furthermore, the held-out set used to determine the threshold tends to share the same distribution with the test set, favoring specific demographics that are the majority.
-That skew (\ie the difference in the performance of an algorithm of certain demographics) is our definition of bias.
+That skew (\ie the difference in the performance of an algorithm of particular demographics) is our definition of bias.
 A key question is: \emph{is \gls{fr} too biased, or not?}
@@
     \centering
-    \caption{\small{\textbf{Database stats and nomenclature, optimal thresholds ($t_o$), and accuracy scores.} \textit{Header:} Subgroup definitions.
+    \caption{\small{\textbf{Database stats and nomenclature.} \textit{Header:} Subgroup definitions.
 \textit{Top:} Statistics of \gls{bfw}.
-\textit{Bottom::} Number of pairs for each partition.
-Columns grouped by race and then further split by gender.
-Out of millions of pairs, accuracy is inconsistent across subgroups.
-Furthermore, $F$ tends to perform inferior to that of $M$ for all but B.}}\label{tab:ethnic-splits}
+\textit{Bottom:} Number of pairs for each partition.
+Columns grouped by ethnicity and then further split by gender.}}\label{tab:ethnic-splits}
     \scalebox{.75}{
@@
     }
-    % \vspace{-12pt}
+    \vspace{-12pt}
 \end{table*}
 
-    Making matters more challenging is that race and ethnicity are loosely defined.
-For example, the US Census Bureau allows an individual to self-identify race.\footnote{\scriptsize\href{https://www.census.gov/mso/www/training/pdf/race-ethnicity-onepager.pdf}{www.census.gov/mso/www/training/pdf/race-ethnicity-onepager}} For this work, we define subgroups as specific sub-populations with face characteristics similar to those found in a region.
+    Making matters more challenging, precise definition of race and ethnicity vary from source-to-source.
+For example, the US Census Bureau allows an individual to self-identify race.\footnote{\scriptsize\href{https://www.census.gov/mso/www/training/pdf/race-ethnicity-onepager.pdf}{www.census.gov/mso/www/training/pdf/race-ethnicity-onepager}} Even gender, our attempt to encapsulate the complexities of the sex of a human as one of two labels.
+Others have addressed the oversimplified class labels by representing gender as a continuous value between 0 and 1 - rarely is a person entirely \emph{M} or \emph{F}, but most are somewhere in between~\cite{merler2019diversity}.
+For this work, we define subgroups as specific sub-populations with face characteristics similar to others in a region.
+Specifically, we focus on 8 subgroups (Fig.~\ref{fig:avg-faces}).
 
-    The adverse effects of a global threshold are two-fold: \textbf{(1)} the mapping produced by a \gls{cnn} is nonuniform.
-Therefore, distances between pairs of faces in different demographics vary in distribution of similarity scores (Fig~\ref{fig:detection-model}); \textbf{(2)} the evaluation set is imbalanced as well.
-Particular demographics making up a majority of the population will carry most weight on the reported performance ratings.
+    The adverse effects of a global threshold are two-fold: \textbf{(1)} mappings produced by \glspl{cnn} are nonuniform.
+Therefore, distances between pairs of faces in different demographics vary in distribution of similarity scores (Fig~\ref{fig:detection-model}); \textbf{(2)} evaluation set is imbalanced.
+Subgroups that make up a majority of the population will carry most weight on the reported performance ratings.
 Reported results favor the common traits over the underrepresented.
-Alas, demographics like gender, ethnicity, race, and age are underrepresented in most public datasets~\cite{merler2019diversity, wang2018racial}.
+Demographics like gender, ethnicity, race, and age are underrepresented in most public datasets~\cite{merler2019diversity, wang2018racial}.
 
-    For \textbf{(1)} we propose subgroup-specific (\ie optimal) thresholds, while we address \textbf{(2)} with a new benchmark dataset for measuring bias in \gls{fr}, \gls{bfw} (Table~\ref{tab:ethnic-splits} and~\ref{tab:compared}).
-\gls{bfw} allows for a fair evaluation of \gls{fr} systems, while enabling demographic-specific ratings to be reported.
-We use \gls{bfw} to gain a deeper understanding of the extent to which bias presents in \gls{soa} \gls{cnn} used to encode faces.
+    For \textbf{(1)}, we propose subgroup-specific (\ie optimal) thresholds while addressing \textbf{(2)} with a new benchmark dataset to measure bias in \gls{fr}, \gls{bfw} (Table~\ref{tab:ethnic-splits} and~\ref{tab:compared}).
+\gls{bfw} serves as a proxy for fair evaluations for \gls{fr} while enabling per subgroup ratings to be reported.
+We use \gls{bfw} to gain an understanding of the extent to which bias is present in \gls{soa} \gls{cnn}s used \gls{fr}.
 Then, we suggest a mechanism to mitigate problems of bias with more balanced performance ratings for different demographics.
-Specifically, we propose using an adaptive threshold that varies depending on the characteristics of detected facial attributes (\ie gender and ethnicity, Fig~\ref{fig:avg-faces}).
+Specifically, we propose using an adaptive threshold that varies depending on the characteristics of detected facial attributes (\ie gender and ethnicity).
 We show an increase in accuracy with a balanced performance for different subgroups.

### 2020-07-09 sections/relatedworks.tex
camera-ready

@@
     \subsection{Bias in \gls{ml}}
-        The progress and commercial value of \gls{ml} is exciting; however, the trust of society in \gls{ml} will take time to build due to inherent biases.
-The exact definitions and implications of bias vary between sources, as do its sources and types.
-A common theme is that bias hinders performance ratings in ways that skew to a particular sub-population-- the source varies, whether from humans~\cite{windmann1998subconscious}, data and label types~\cite{tommasi2017deeper}, \gls{ml} models~\cite{amini2019uncovering, kim2019learning}, or evaluation protocols~\cite{stock2018convnets}.
-For instance, a vehicle-detection model might miss cars if training data was mostly trucks.
-In practice, many \gls{ml} systems learn on biased data, which could be detrimental for society.
+        The progress and commercial value of \gls{ml} are exciting.
+However, due to inherent biases in ML, society is not readily able to trust completely in its widespread use.
+The exact definitions and implications of bias vary between sources, as do its causes and types.
+A common theme is that bias hinders performance ratings in ways that skew to a particular sub-population.
+In essence, the source varies, whether from humans~\cite{windmann1998subconscious}, data or label types~\cite{tommasi2017deeper}, \gls{ml} models~\cite{amini2019uncovering, kim2019learning}, or evaluation protocols~\cite{stock2018convnets}.
+For instance, a vehicle-detection model might miss cars if training data were mostly trucks.
+In practice, many \gls{ml} systems learn biased data, which could be detrimental to society.
 
     \subsection{Bias in \gls{fr}}
-        Problems of bias in \gls{fr} have been driven by different motivations; therefore, solutions have been posed for different matters.
-Instances can be in problems of data augmentation~\cite{yin2019feature}, one-shot learning~\cite{ding2018one}, demographic parity and fairness with priority on privacy~\cite{huang2018generative}, domain adaptation~\cite{wang2018racial}, differences in face-based attributes across demographics~\cite{wang2018they}, and even data exploration~\cite{muthukumar2019}.
-Yin~\etal proposed to augment the feature space of underrepresented classes using other classes with a diverse collection of samples to encourage distributions of underrepresented classes to more closely resemble that of the others~\cite{yin2019feature}.
+        Different motivations have driven problems of bias in \gls{fr}.
+Instances can be in issues of data augmentation~\cite{yin2019feature}, one-shot learning~\cite{ding2018one}, demographic parity and fairness with priority on privacy~\cite{huang2018generative}, domain adaptation~\cite{wang2018racial}, differences in face-based attributes across demographics~\cite{wang2018they}, data exploration~\cite{muthukumar2019}, and even characterizing different commercial systems~\cite{buolamwini2018gender}.
+
+        Yin~\etal proposed to augment the feature space of underrepresented classes using different classes with a diverse collection of samples~\cite{yin2019feature}.
+This was to encourage distributions of underrepresented classes to resemble the others more closely.
 Similarly, others formulated the imbalanced class problem as one-shot learning, where a \gls{gan} was trained to generate face features to augment classes with fewer samples~\cite{ding2018one}.
 \gls{gapf} was proposed to create fair representations of the data in a quantifiable way, allowing for the finding of a de-correlation scheme from the data without access to its statistics~\cite{huang2018generative}.
-Wang~\etal defined subgroups at a finer-level (\ie Chinese, Japanese, Korean), and determined the familiarity of faces inter-subgroup~\cite{wang2018they}.
+Wang~\etal defined subgroups at a finer level (\ie Chinese, Japanese, Korean), and determined the familiarity of faces inter-subgroup~\cite{wang2018they}.
 Genders have also been used to make subgroups (\eg for analysis of gender-based face encodings~\cite{muthukumar2019}).
@@
         % For this, we expect bias from data, for this is the training set. 
-        Furthermore, \gls{bfw} balances gender and race: \gls{rfw} splits subgroups by gender and race, while \gls{bfw} has gender, race, or both). 
+        Furthermore, \gls{bfw} balance subgroups: \gls{rfw} splits subgroups by gender and race, while \gls{bfw} has gender, race, or both). 
         % Thus, \gls{rfw} and \gls{bfw} complement one another. % - both add demographic for subjects with faces in renowned, large-scale \gls{fr} datasets.
 
-Most similar to us is~\cite{das2018, demogPairs, lopez2019dataset, srinivas2019face} - each were motivated by insufficient paired data for studying bias in \gls{fr}.
+Most similar to us is~\cite{das2018, demogPairs, lopez2019dataset, srinivas2019face} - each was motivated by insufficient paired data for studying bias in \gls{fr}.
 Then, problems were addressed using labeled data from existing image collections.
-Specifically, Hupont~\etal curated a set of faces based on racial demographics (\ie \gls{a}, \gls{b}, and \gls{w}) called \gls{dp}~\cite{demogPairs}, while~\cite{srinivas2019face} honed in on adults versus children called \gls{itwcc}.
+Uniquely, Hupont~\etal curated a set of faces based on racial demographics (\ie \gls{a}, \gls{b}, and \gls{w}) called \gls{dp}~\cite{demogPairs}.
+In contrast,~\cite{srinivas2019face} honed in on adults versus children called \gls{itwcc}.
 Like the proposed \gls{bfw}, both were built by sampling existing databases, but with the addition of tags for the respective subgroups of interest.
-Aside from the additional data of \gls{bfw} (\ie added an additional subgroup \gls{i}, along with additional subjects with more faces for all subgroups), we also further split subgroups by gender.
+Aside from the additional data of \gls{bfw} (\ie added subgroup \gls{i}, along with other subjects with more faces for all subgroups), we also further split subgroups by gender.
 Furthermore, we focus on the problem of facial verification and the different levels of sensitivity in cosine similarity scores per subgroup.
@@
         \centering
-        \caption{\small{\textbf{\gls{bfw} compared to related datasets.} Our \gls{bfw} is balanced across ID, gender, and ethnicity (Table~\ref{tab:ethnic-splits}).
-Compared with \gls{dp}, \gls{bfw} provides more samples per subject and subgroups per set, while only using a single resource, VGG2.
-\gls{rfw}, on the other hand, supports a different task (\ie domain adaptation), and focuses on race-distribution, but not the distribution of identities.}}
+        \caption{\small{\textbf{\gls{bfw} and related datasets.} \gls{bfw} is balanced across ID, gender, and ethnicity (Table~\ref{tab:ethnic-splits}).
+Compared with \gls{dp}, \gls{bfw} provides more samples per subject and subgroups per set, while using a single resource, VGG2.
+\gls{rfw}, on the other hand, supports domain adaptation, and focuses on race-distribution - not the distribution of identities.}}
         \scriptsize
@@
         \label{tab:compared}
-%            \vspace{-5mm}
+            \vspace{-4mm}
     \end{table*}
@@
 Bias is not unique to \gls{ml} - humans are also susceptible to a perceived bias.
-In fact, biases exist across race, gender, and even age~\cite{10.1007/978-3-030-13469-3_68, bar2006, meissner2001, nicholls2018}.
+Biases exist across race, gender, and even age~\cite{10.1007/978-3-030-13469-3_68, bar2006, meissner2001, nicholls2018}.
 Wang~\etal showed machines surpass human performance in discriminating between Japanese, Chinese, or Korean faces by nearly 150\%~\cite{wang2018they}, as humans just pass random (\ie 38.89\% accuracy).
 
-We expect human bias to skew to their own genders and races.
-For this, we measure the human perception with face pairs of different subgroups (Section~\ref{subsec:human-assessment}).
+We expect the human bias to skew to their genders and races.
+For this, we measure human perception with face pairs of different subgroups (Section~\ref{subsec:human-assessment}).
 The results concur with~\cite{wang2018they}, as we also recorded overall averages below random ($<$50\%).

### 2020-07-09 sections/section3.tex
camera-ready

@@
 \section{The BFW Benchmark and Dataset}
-
-We now discuss the \gls{bfw} dataset (\ie collection and preparation), along with the evaluation protocols. We conclude this section by reviewing the settings used to survey bias in humans.
+We now discuss the acquisition of \gls{bfw}, protocols, and settings used to survey bias in humans.
 
@@
 
-\vspace{1mm}
+\vspace{.5mm}
 \noindent\textbf{Compiling subject list.} 
@@
 
-\vspace{1mm}
+\vspace{.5mm}
 \noindent\textbf{Detecting faces.} Faces were detected using MTCNN~\cite{zhang2016joint}.\footnote{\href{https://github.com/polarisZhao/mtcnn-pytorch}{https://github.com/polarisZhao/mtcnn-pytorch}} Then, faces were assigned to one of two sets.
@@
 
-\vspace{1mm}
+\vspace{.5mm}
 \noindent\textbf{Validating labels.} 
@@
   \glsunset{fnr}
-	\includegraphics[width=.9\linewidth]{figures/detcurve-improved.pdf}
+	\includegraphics[width=.83\linewidth]{figures/detcurve-improved.pdf}
 		\caption{\small{\textbf{\gls{det} curves.} \emph{Top-left}: per gender.
@@
 \label{fig:detcurves} 
+ \vspace{-4mm}
 \end{figure} 
 %\vspace{1mm}
-\noindent\textbf{Sampling faces and creating folds.} We created lists of pairs in five-folds with subjects split evenly per subject and without overlap across folds.
+\noindent\textbf{Sampling faces and creating folds.} We created lists of pairs in five-folds with subjects split evenly per person and without overlap across folds.
 Furthermore, a balance in the number of faces per subgroup was obtained by sampling twenty-five faces at random from each.
@@
 % 		\glsreset{sdm}
+	    \vspace{-4mm}
 \end{figure}
@@
 This formulation (\ie \gls{fv}) is highly practical for applications like access control, re-identification, and surveillance.
-Typically, training a separate model for each unique subject is infeasible.
+Typically, training a separate model for each unique subject is unfeasible.
 Firstly, the computational costs compound as the number of subjects increase.
@@
 	\centering    
-	\includegraphics[width=.75\linewidth]{figures/confusion.pdf}
-		\caption{\small{\textbf{Confusion matrix.} Error (Rank 1, \%) for all faces of \gls{bfw} vs.
-all others.
-Error concentrates intra-subgroup - consistent with the \gls{sdm} (Fig.~\ref{fig:detection-model}), as \gls{af} performs the worst, mostly confused by \gls{af}.
-Although race/ethnicity are challenging to define, this shows the subgroups are meaningful for \gls{fr}.}}
+	\includegraphics[width=.65\linewidth]{figures/confusion.pdf}
+		\caption{\small{\textbf{Confusion matrix.} Error (Rank 1, \%) for all \gls{bfw} faces versus all others.
+Errors concentrate intra-subgroup - consistent with the \gls{sdm} (Fig.~\ref{fig:detection-model}).
+Although subgroups are challenging to define, this shows the ones chosen are meaningful for \gls{fr}.}}
 		\label{fig:confusion} 
-%		    \vspace{-6mm}
+		  %  \vspace{-3mm}
 \end{figure}
@@
 
-Samples were collected from individuals recruited from multiple sources (\eg social media, email lists, and family/friends) - a total of 120 participants were sampled at random from all the submissions that were (1) complete and (2) done by a W or C participant.
+Samples were collected from individuals recruited from multiple sources (\eg social media, email lists, family, and friends) - a total of 120 participants were sampled at random from all submissions that were completed and done by a W or C participant.
 Specifically, there were 60 W and 60 C, both with \gls{m} and \gls{f} split evenly.
 A total of 50 face pairs of non-famous ``look-alikes'' were collected from the internet, with 20 ({\emph WA}) and 20 ({\emph C}) pairs with, again, \gls{m} and \gls{f} split evenly.
-The other 10 pairs were of a different demographic (\eg Hispanic/ Latino, Japanese, African).
-The survey was created, distributed, and recorded via \href{https://paperform.co}{PaperForm}.
-It is important to note that participants were only aware of the task (\ie to verify whether or not a face-pair was a \emph{match} or \emph{non-match}, but with no knowledge of it being a study on bias).
+The other 10 were of a different demographic (\eg Hispanic/ Latino, Japanese, African).
+The survey was created, distributed, and recorded via \emph{\href{https://paperform.co}{PaperForm}}.
+It is important to note that participants were only aware of the task (\ie to verify whether or not a face-pair was a \emph{match} or \emph{non-match}, but with no knowledge of it being a study on the bias).
